### 2022-10-01 drafts/dcr_white_identity_relative_deprivation_fa_2022.qmd
another draft

---
title: |
  Economic concerns appear to be weak causal predictors of White political identity
short-title: Economic concerns
published: "Manuscript draft: please do not share without author permission."
code-repo: "For replication, go to: <https:://github.com/DamonCharlesRoberts/white-identity-sources>."
author:
  - name: Damon C. Roberts
    email: damon.roberts-1@colorado.edu
    orcid: 0000-0002-4360-3675
    title: PhD Candidate
    affiliations:
      - id: CU
        name: University of Colorado Boulder
        department: Political Science
        address: 333 UCB
        city: Boulder
        region: CO 
        postal-code: 80309-0333
abstract: |
    Does economic or social and political threat explain white identity? Two leading arguments suggest that White identity results from a disruption of the racial status quo. However, they disagree on which disrupted element drives it. While one suggests that White identity comes from discriminatory attitudes, the other suggests that it is the result of concern with increased economic competition between Whites and other racial groups. Using data from the American National Election Study that measure egotropic and sociotropic economic attitudes and leveraging questions gauging attitudes about social and political changes of influence from the 2012, 2016, and 2020 American National Election Studies, this manuscript takes advantage of penalized regression for variable selection. On balance, the models suggest that the hypothesis of economic competition is weaker than some suggest. I take this as evidence that White political identity is largely the result of Whites' concerns with protecting the social and political status quo - and not so much economic.
thanks: |
    I would like to thank Jennifer Wolak for her feedback on the many drafts of this manuscript. I would also like to thank the Keller Center for the First Amendment for the research appointment that gave me the space to work on the early stages of this project. I would also like to thank Anand Sokhey and John Griffin for their advice and suggestions in the early drafts of this paper. I would also like to thank Tabitha Bonilla, Nicole Yadon, and participants and audience members of a Midwestern Political Science Association panel for which an earlier draft this idea was presented. I would also like to thank Tamar Malloy, Andrew Philips, Kathryn Schauer, Rachel O'Neal, Alexander Jensen, and Komal Kauer for their encouragement and feedback at various stages of this project's development. All errors contained in this manuscript are ultimately mine.
additional-info: |
  I have no known conflict of interest to disclose.
keywords:
  - white identity
  - economic threat
  - cultural threat
date: today
bibliography: "../assets/wid.bib"
format:
  hikmah-pdf: 
    biblatex-chicago: true
    biblio-style: authordate
    biblatexoptions:
      - backend=biber
      - autolang=hyphen
      - isbn=false
      - uniquename=false
linestretch: 2
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: block-setup

# set seed
set.seed(90210)

# source cleaning scripts
source('../code/dcr_anes_2012_cleaning.R')
source('../code/dcr_anes_2016_cleaning.R')
source('../code/dcr_anes_2020_cleaning.R')

# modularly load functions
box::use(
    dplyr = dplyr[select, filter, mutate, case_when, rename],
    tidyr = tidyr[drop_na],
    mice = mice[mice, pool],
    miceadds = miceadds[mids2datlist],
    brms = brms[brm, brm_multiple, set_prior, lasso, cumulative, prior],
    tibble = tibble[tribble],
    modelsummary = modelsummary[datasummary_skim, modelsummary],
    parameters = parameters[parameters, standardize_names],
    ggplot2 = ggplot2[ggplot, aes, position_dodge, theme_minimal, geom_vline, labs, theme, element_text, ggsave],
    tidybayes = tidybayes[geom_pointinterval]
)

# create list object for datasets
data = list()

# load the cleaned data
    #* 2012
data[['anes_2012']] = read.csv('../data/anes-2012/anes-2012-updated.csv') |>
    select(wid, female, rboff, epast, unpast, pid, edu, losejob, immtakejobs, raceResent, ecfamily, tradbreak, govbias, inflwhite, whitediscrim)
    #* 2016
data[['anes_2016']] = read.csv('../data/anes-2016/anes-2016-updated.csv') |>
    select(wid, female, rboff, epast, unpast, pid, edu, losejob, immtakejobs, raceResent, ecfamily, tradbreak, govbias, inflwhite, whitediscrim)
    #* 2020
data[['anes_2020']] = read.csv('../data/anes-2020/anes-2020-updated.csv') |>
    select(wid, female, rboff, epast, unpast, pid, edu, losejob, immtakejobs, raceResent, ecfamily, govbias, inflwhite, whitediscrim)
```

# Introduction

Those conducting post-mortem analyses of the 2016 Presidential election of Donald J. Trump, a presidential candidate that personifies White grievance politics, contend that the stronger than expected presence of non-college educated Whites explains his electoral success [@griffin_et-al_2017]. Analyses by political scientists largely contribute to Trump's success in attracting voters who see their Whiteness as a politically relevant identity where they are distinct from Americans of other racial groups [@jardina_2020_pp]. For a long time, political scientists saw White political behavior and attitudes through the lens of prejudice. These analyses exclude another dimension of Whites' racial attitudes: their affinity for, feelings of linked fate with other Whites and an expression that their race is a politically-relevant identity. More recently, scholars using the @tajfel_turner_1986 social psychological framework of identity challenge this tendency to focus on Whites' prejudice by arguing that some Whites may see their race as an identity essential and relevant to politics. As a result, while some may express out-group attitudes, they also express an in-group preference for other Whites [@jardina_2019_cup]. Despite the importance of these attitudes, the source of these in-group preferences remains somewhat unclear.

There appears to be a mismatch between those who study sources of a variety of social identities and those that study White political identity. Scholars studying identity more broadly often focus on economic threats as a key motivator for an expressed connection to a social in-group [see @huddy_2001], however some scholars studying White political attitudes and racial identity tend to focus on the role of cultural and political threats [see @bartels_2020]. This mismatch leads to some to conceptualizing White identity as originating from economic views, in the tradition of the identity literature, while others conceptualize white identity as coming from political and cultural threat.

I agree with those who focus on the connection between White political identity and cultural and political threats. I argue that White identifiers are concerned about their status in politics and society. Though, I acknowledge that economic threat is a potent activation of identity and encourages engagement of group-based behavior, White identity holds unique cultural and institutional significance in the U.S. [@rile-hayward_2013]. Concerns about economic standing is not entirely new to Whites (e.g., the Great Depression, the 2008 Housing Market Crash) and is likely more nebulous as to whether it disproportionately effects Whites. As political and societal power are shifting away from Whites, I suspect that this reflects a change to the racial hierarchy that is new to Whites and this may be driving the high levels of White political identification that many scholars report.

I examine the correlates of white identity. Using data from the 2012, 2016, and 2020 American National Election Study, I utilize penalized regression to test the claim that a handful of predictors that reflect attitudes about cultural and political power are more influential than general (sociotropic) and pocketbook (egotropic) attitudes about the economy. In each year, I find that cultural and political predictors are consistently stronger than economic attitudes. I close by arguing that scholars should continue to examine the mechanisms that explains this connection.

<!-- 
TODO: 
  - I've set up the puzzle, now explain where the literature is at
    - DONE
  - explain how the relative deprivation argument is better covered in the identity literature
    - DONE
  - make a strong assertion that I am testing against the relative deprivation argument and that I am in support of cultural explanations
  - explain what the next few sections of the paper do.
    - DONE
  - listen to the advice of Stimson and the Offspring
    - DONE
-->

<!--
The leading argument conceptualizes those expressing a positive White political identity as those motivated by group delineation and difference [@jardina2019] through cultural or other non-materialistic evaluations. Another interpretation of in-group affinity in the debate surrounding social identity theory sees in-group affinity as the result of relative deprivation or group competition [see @Huddy2012]; that is, a reaction to threats to material status. Those who argue for the latter illustrate this idea of group competition because non-college-educated Whites in 2016, who may be more susceptible to economic hardship and job market competition, were more supportive of Trump due to their concern about more staunch group competition. Many others disagree and argue that the 2016 electoral base for Trump was about prejudice and discrimination among Whites and that economic correlates are weak causal features [@fahey_et-al_2020].

To adjudicate between these two explanations of where White identity comes from in the post-Obama or "post-racial" era of American politics, I rely on the literature on social identity theory. This manuscript examines whether group conflict and feelings of relative material deprivation among Whites increase White identity.

Given the growing body of literature on the implications of White political identity, I contend that the evidence points to measures that capture group differences and concern with social positioning as opposed to materially-based group competition or relative deprivation. There is ample evidence in the literature that many of the precursors for a politically relevant social identity among Whites are present. While group competition and relative deprivation may explain the motivation of some Whites to express in-group affinity, it is much farther from a complete explanation than proponents of it admit.

Using data from 2012, 2016, and 2020 American National Election Studies (ANES), I use common sociotropic (generalized attitudes of the economy, overall) and egotropic (more specific attitudes related to their personal financial situation) measures of an individual's attitudes towards the economy. I then put them in a penalized regression model along with several questions gauging White respondents' feelings of a loss in cultural and political influence. The model provides evidence supporting my expectations that non-material concerns are stronger causal predictors of White identity than economic indicators.
-->

# Identity and White political identity

<!--
TODO:
  - Start with evidence that demonstrates how economic threats activate identity politics, then discuss how people suggest that reactions to Trump are economic rather than racialized, then go into how others suggest that it is actually about cultural and political threat
-->

Social identity theory suggests that individuals form attitudes based on groups that they associate themselves with, even if those groups are subjectively defined [see @tajfel_turner_1986]. In the famous Robber's Cave experiment, @sherif_1988 divided a group of adolescent boys into two groups arbitrarily. After a week of spending time with others in the group, the boys had already provided labels that defined the two groups and distinguished themselves; one group called themselves the Eagles and the other group referred to themselves as the Rattlers. With the introduction of competition, these boys reported fierce levels of hostility toward the members of the other group and expressed positive attitudes toward those within the group.

This example illustrates a number of features of social identities. The first feature of social identity theory is that these groups that we may associate ourselves with are often subjectively defined [see @huddy_2001]. The second feature is that these subjective definitions of the group make distinctions between the two groups that come to matter to members of each group - there are those part of our group and there are those that are part of that other group. Many cognitive theories suggest that this occurs as a result of the tendency for humans to reduce cognitive load by grouping things together [@huddy_2001]. The third feature is that the introduction of competition between these groups shapes attitudes toward those within and without your group. The motivator is that for groups that have higher status often want to keep that status and individuals in the high status group are much more likely to support inequality to satisfy that motivation [@burns_kinder_2012;@morrison_et-al_2009]. Fourth, these attitudes are long lasting and can be strongly held. An additional feature of theories of social identitiy is that the degree to which one might identify with one group or another is fluid and that whether one identity that one might hold is active or if it is another often depends on the context and the identity's relevance to that context [@huddy_2012].

Racial groups in the US contain a number of these features. As a result,a burgeoning literature in political science that has made this connection takes the racial identity of Whites more seriously as one that informs their political attitudes [see @jardina_2019_cup]. Race in America is subjectively defined [see @rile-hayward_2013] and these definitions are deeply embeded and reinforced by our institutions [@lopez_2006]. Whiteness and one's willingness to identify with that group is important [@wong_cho_2005] as it reflects a desire for an association with the history and affordances that comes with being White [@mcdermott_samson_2005].

In many eras of American history, there are a number of instances where we have sought to reclassify who is White. During the antebellum period, Whiteness was no longer simply one's economic class but now it had become a reflection of a racial hierarchy that afforded some various rights and excluded others from enjoying the same rights (e.g., literacy tests). During this era, rather than conceptualizing Whites as land-owners from Caucasian descent, the definition of what was White shifted to include those who had long considered to be non-White Europeans such as the Irish. Changes to definitions of who is White placed race and not just class as a primary source of who is afforded some rights and who are not [@rile-hayward_2013].

Whites hold a number of strong political attitudes that are racialized by politicians. Whites' attitudes toward welfare programs are not all too distinct from their attitudes toward Black Americans [@gilens_1999]. A provided explanation is that social welfare policies are more unpopular among Whites when the media frames these policies as benefiting non-Whites [@winter_2008;@van-doorn_2015]. When attitudes toward non-Whites are explicitly cued by messages that remind individuals of the racial hierarchy, these attitudes toward Blacks are more entrenched [@lyle_2014; see also @wellman_et-al_2016]. Attitudes about healthcare were shaped by prominent conservatives who pitched that Barack Obama, the first Black president, supported the Affordable Care Act as a way to disproportionately help Black Americans [@tesler_2012;@tesler_2016], that there was a rise of survey reports expressing old fashioned racism among Whites during his administration [@tesler_2013]. While Whites tend to report negative views about government aid for those in need when they are non-White recipients, evidence suggests that Whites who live in communities with a large porportion of the poor being White report much more positive views of the poor than those that view the poor as largely represented by non-Whites [@hopkins_2009] and that telling Whites that other Whites are going to be the beneficiaries of an entitlement program increases their support for such a program [@wetts_willer_2018].

As an identity, a large portion of the literature views these attitudes as coming from the presence or focus on economic threats. Some studying White political identity find economic predictors as substantively and statistically significant [see @perez_et-al_2022_pp]. From non-experimental evidence, others make similar arguments about what predicts white political identity. @bunyasi_2019 provides evidence which suggests that Trump's messages highlighting increased job market competition that was harming Whites because of affirmative action was a successful strategy and helped build his base of support. After all, the identity literature, more broadly, often focuses on economic threats. Evolutionary psychologists often take the position that identity and the resulting attitudes and behaviors directed toward out-groups is the result of competition over material resources [see @huddy_2001]. Among political scientists, many use this as an explanation for negative attitudes directed toward immigrants among those who feel insecure about their job prospects [@pardos-prado_xena_2019].

Not all agree, however. Some suggest that the negative attitudes that Whites express toward immigrants are a result of concerns about cultural changes to those that live in border towns [@hainmueller_hopkins_2014]. Others suggest that simply reminding Whites of these demographic shifts lead to an activation of these racial identity-based attitudes [@craig_richeson_2014]. Some other evidence suggests that these identity-based attitudes are better predicted by ethnic antagonism [@bartels_2020] and by racial animus [@fahey_et-al_2020] and explicitly take the position that these are not driven by a reaction to economic threats. However, these attitudes tend to focus on prejudiced views rather than a reflection of a connection to other whites.

# Is white identity non-material?

As much of the literature on Whites' racial attitudes rely heavily on social identity theory as a paradigm, a significant portion of it presents the source of these group-based views as responses to economic threat. However, some disagree with such a position and view these attitudes as the result of changes to the racial hierarchy. Other perspectives provide explanations for Whites' attitudes and conceptualize the definition of Whiteness and White identity as a desire for retaining political power.

Historical accounts that contextualize attitudes stemming from White racial identity seem to suggest that, though economics is very embedded with racial attitudes, what it meant to be White was no longer about class but it became about what it meant for access to and privileges afforded by institutions [@rile-hayward_2013]. In the 19th century privilege came from one's freedom and from their wealth. For those who were not enslaved but were poor, they were not considered White and were than Whites who owned land [@rile-hayward_2013]. With the abolition of slavery, the concern became not about wealth but now about race. This motivated a shift in how the institutions and how Whites defined themselves. It became the focus for Whites to be a higher status group, regardless of whether or not they were poorer than formerly enslaved Black Americans [@rile-hayward_2013].

Many conceptualizations of White political identity align with its primary source as concern for the group's political power. While a key component of identity is a feeling of your fate as linked with others in your group [@leach_et-al_2008], this linked fate is conceptualized as the connection between attitudes and material forms of threat but that identity, in and of itself, is better conceptualized as a result of cultural or ideational threat [@donnelly_2020]. Aligning with the contextual account of racial identity for Whites, many conceptualize White identity as an expressed desire for a recognition of privilege and status by insitutions [@lopez_2006] and a connection with the history that connects Whites to their status [@mcdermott_samson_2005]. The prominent account of White political identity and its implications in contemporary American politics argues for a strong connection between White political identity and a desire to protect this racial hiearchy and status which reflects the tendency to report in-group oriented, and not necessarily out-group oriented, attitudes [@jardina_2019_cup].

As this is an open question and is rather a complex one, I attempt to first wade into this by parsing through which of these two perspectives are more plausible. As I discuss in the concluding section, I argue that scholars should continue to take the conceptualization and measurement of White political identification seriously and examine the particular mechanisms that connect economic and political threat to White political identity. In the next section, I examine whether attitudes expressing concern with economic threats or political, or broader societal, threats are stronger predictors of the degree to which someone feels that being White is important to their identity. The goal is not to test particular causal mechanisms but to examine whether my belief and other's that White identity reflects reactions to political threat moreso than it does economic.

<!-- 
TODO:
    - Then move into making my argument. We've got two competing views - the one that is popular in the identity lit, and then the one that suggests that it could be cultural or political threat. Then argue that what it really is that whites have not experienced cultural/political threats to the racial hierarchy and explain how historically that has been responded to.
      - DONE
-->

# Data

To examine whether Whites' attitudes expressing concern about political threats are stronger than their perceptions of the economy, I use 2012, 2016, and 2020 American National Election Studies (ANES). The studies are nationally representtaive questions that capture several questions that lend themselves to testing the relationship between economic and political concerns while also asking White respondents to report how important being White is to their identity.

Each study asks White respondents "How important is being White to your Identity" and offers them a five-point Likert scale with the options reflecting level of agreement with the question ^[Details about question wording, original ANES coding, and my coding for all variables I include in my models are included in @tbl-anes-measures]. They do not ask White respondents to report on measure items that are included in the White political identity measures of @jardina_2019_cup or the more complex battery of items that @helms_carter_1990 discuss. However, this question does reflect the conceptualization of White identity that reflects the degree to which one sees their Whiteness as part of who they are and reflects the conceptualization that others have used [see @schildkraut_2017;@perez_et-al_2022_pp].

The ANES does not ask questions that explicitly capture the attitudes of respondents about perceptions of the economy for different groups. Those that have lower incomes are more likely to feel insecurity about their financial status and are more likely to use this economic threat as a motivation for their negative attitudes toward immigrants [@pardos-prado_xena_2019]. Therefore, one measure of economic threat that I use is the respondent's reported level of family income. Those who hold lower levels of education are also likely to feel similar levels of insecurity about their economic status. Whites who report that they feel that their family is doing worse or will be doing worse also reflects a degree of insecurity about their economic status and may reflect perceptions of economic threat [@jardina_2019_cup]. I additionally include an item reflecting respondent's perceptions of economic mobility today relative to in the past. I additionally include questions about retrospective sociotropic evaluations of the economy. In 2012, Whites harboring negative attitudes of Obama rooted in racial animous were more likely to express more negative views of the economy relative to how it had been in the past [@sides_et-al_2018].

The ANES includes a number of questions that reflect perceptions of Whites losing political power. First I include a question that asks respondents to report whether or not they felt that Whites currently have too little or too much influence on politics today. I also want to examine the degree to which respondents feel that the institutions benefit non-Whites rather than whites by including a question that Whites face discrimmination and that the government is biased against Whites. I additionally include a question asking respondents about their belief that traditions in this country are breaking^[The ANES only include this question in the 2012 and 2016 questionaire].

Furthermore, the ANES includes the racial resentment battery which asks a series of questions that reflect Whites' beliefs that Blacks are individually responsible for their circumstances,  and that they are undeserving of any assistance. I control for racial resentment as the individual items likely confound the relationship between both economic and political threat measures on expressed White identity.

The summary of the literature points to a large number of potential predictors that influence the degree to which one takes on White political identity.  To adjudicate between hypotheses, there is value in a comprehensive model that considers the plausibility of each. Most models consider those predictors they believe are most plausible and do not include others that the researcher believes to be less plausible. They understandably do this out of concern with increases in the inefficiency of one's regression estimates the more predictors you include and therefore increases the risk of overfitting your model to your particular sample, thereby erroneously generalizing conclusions drawn from the model. Though this concern is reasonable, such a model necessarily neglects to examine the effects of the predictors a researcher believes to be less plausible.

Therefore, I elect to use a penalized regression model to balance between these concerns.  The literature surveyed above suggests that there are a small number of variables that do predict White political identity, and there are many others that have no effect. This implies that the L2 regularization parameter equals 0. This is because the L2 regularization parameter shrinks coefficients asymptotically toward zero. The L1 regularization parameter, however, does allow for a shrinkage parameter to pull coefficients to 0. Given this, it suggests that the more appropriate model to capture this debate would be the LASSO penalized regression.

One option for implementation is to use the ``classical" LASSO, which relies on an optimization procedure through cross-validation. It is a frequentist approach to LASSO penalized regression. However, there are several positive benefits to the second implementation of the LASSO penalized regression in a Bayesian framework [see @van-erp_et-al_2019_jmp]. From my review of the literature, I have \textit{apriori} expectations for the distribution of effects on my predictors in the form of a Laplace distribution which I explicitly define in my prior density function. I also have the additional benefit of having the ability to define its hyperparameters. Meaning I suspect that there is a peak around zero and that there are relatively large tails. This characterization of my prior fits with that of the Lasso - there are a large number of included predictors that have zero effect with a relatively smaller proportion of included predictors that do have an effect but are relatively unsure as to how large those coefficients might be. I can specify the scale of that uncertainty by specifying a hyperparameter for this LASSO prior. Without having to perform cross-validation as I would with the classical LASSO, my posterior distribution naturally reflects the uncertainty around the size of my L2 shrinkage parameter and can estimate models with different L2 shrinkage parameters simultaneously. As \verb|stan| includes implementations of this model, I rely on Hamiltonian Monte Carlo, a type of Markov Chain Monte Carlo sampling which allows for non-convexity in the L2 shrinkage parameters. As classical LASSO models are challenging to interpret due to cross-validation, using credible intervals offers a much less error-prone interpretation.

My measure of White Identity is ordinal. Though White Identity as a concept is likely continuous, through the specification of a cumulative density function with a logit link, my model bins this latent continuous space based on the number of response options. Though it creates a more complicated computational task, it is a more appropriate model specification given the available measure and my assumptions about the concept [@liddell_kruschke_2018_jesp].

To account for potential problems stemming from missing data, I impute my data by using multiple imputation with chained equations where the particular model I use is a random forest model to allow for more flexibility and to reduce model dependence [@buuren_goothuis-oudshoorn_2011, @marbach_2021]^[I impute 10 datasets and simultaneously fit my Bayesian LASSO on each dataset and pool the results. Uncertainty reflected in my reported credible intervals not only reflect the uncertainty from one model but my reported uncertainty also reflects the uncertainty generated from the imputation procedure. One drawback with this particular approach is that my normalized split-$\hat{R}$ will likely be greater than the widely recognized 1.01 cutoff that indicatess model convergence as a result of the pooling of my multiply imputed datasets.]. I report the results of the model fitted using listwise deletion in @tbl-2012-lasso-simple, @2016-lasso-simple, and @tbl-2020-lasso-simple.

@fig-2012-2016-2020-ordinal-lasso presents the estimates of my fitted ordinal models with the LASSO shrinkage prior^[@tbl-2012-2016-2020-ordinal-lasso presents the table of results.]. These models are fit on 6 chains with 2000 iterations each. To make the conclusion of whether a particular predictor "matters" for predicting White political identity, I examine the estimated credible intervals, which report the probability that the true estimate falls within the estimated range. I construct these credible intervals by reporting estimates between the 2.5% and 97.5% quantiles. This means that for a given credible interval, there is a 95% probability that the true value falls within that range. For values outside the range of the credible interval, it is relativelyu implausible that the given value is the true value.

<!--
TODO:
  - Discussion of why these economic indicators versus others. Set up expectations about how/why these are different and what their effects on WID may be
  - Discussion of why these cultural/political indicators versus others. Set up expectations about how/why these are different and what their effects on WID may be 
  - Shorten discussion of penalized regression. Add details in footnotes and appendix
  - Footnote details about multiple imputation
  - Why include white influence and white discrimmination measures? Are they endogeneous?
  - Why include PID, what are my expectations and motivation for this and other control variables?
  - What does tradition breaking item represent?
-->

## Evidence from a penalized regression

## Results

<!--
TODO:
  - Group rows of results for figures and tables by economic and cultural/political attitudes
  - Cut discussion of priors about year to year differences
    - just do a post-hoc "isn't this interesting" interpretation and a discussion of what might be going on
  - Discuss a footnote on multicollinearity to make sure that economic nulls aren't due to being highly intercorrelated
-->

# Conclusion

<!--
TODO:
  - Make a strong assertion about my results
  - Discuss why I think economic factors don't matter when they matter in most identity papers
  - What does this mean for how we understand white identity? Elaborate on the takeaway of the paper and what I added but also what questions are still there.
-->

# Appendix

+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
|                                          | Abbreviated question                                       | ANES Coding                                                                                                                                                      | Author coding                                                                                                                                                             | Years            |
+==========================================+============================================================+==================================================================================================================================================================+===========================================================================================================================================================================+==================+
| White identity                           | How important is being White to your identity              | 1 = Extremely important; 2 = Very important; 3 = Moderately important; 4 = A little important; 5 = Not at all important; \<1 Missing, not asked, etc             | 1 = Not at all important; 2 = A little important; 3 = Moderately important; 4 = Very important; 5 = Extremely important; NA = Missing, not asked, etc                     | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| **Economics**                            |                                                            |                                                                                                                                                                  |                                                                                                                                                                           |                  |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Retrospective worse off                  | How much better worse off than 1 year ago                  | 1 = Much better; 2 = Somewhat better; 3 = The same; 4 = Somewhat worse; 5 = Much worse, \< 1 Missing, not asked, etc                                             | -2 = Much worse; -1 = Somewhat worse; 0 = The same; 1 = Somewhat better; 2 = Much better; NA = Missing, not asked, etc                                                    | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Economy worse                            | U.S. economy better worse off than 1 year ago              | 1 = Much better; 2 = somewhat better; 3 = Stayed about the same, 4 = Somewhat worse; 5 = Much worse; \< 1 Missing, not asked, etc                                | -2 = Much worse; -1 = Somewhat worse; 0 = The same; 1 = Somewhat better; 2 = Much better; NA = Missing, not asked, etc                                                    | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Unemployment worse                       | Unemployment better or worse than 1 year ago               | 1 = Much better; 2 = somewhat better; 3 = Stayed about the same, 4 = Somewhat worse; 5 = Much worse; \< 1 Missing, not asked, etc                                | -2 = Much worse; -1 = Somewhat worse; 0 = The same; 1 = Somewhat better; 2 = Much better; NA = Missing, not asked, etc                                                    | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Education                                | Educational attainment                                     | 1 = \< than High school; 2 = High school; 3 = Some post-High school; 4 = Bachelor's degree; \< 1 Missing, not asked, etc                                         | 1 = \< than High school; 2 = High school; 3 = Some post-High school; 4 = Bachelor's degree; NA = Missing, not asked, etc                                                  | 2012, 2020       |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Education (2016)                         | Educational attainment                                     | See ANES codebook                                                                                                                                                | See ANES codebook; NA = MIssing, not asked, etc                                                                                                                           | 2016             |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Worried about losing job                 | How worried about losing job in near future                | 1 = Not at all; 2 = A little; 3 = Moderately; 4 = Very; 5 = Extremely; \< 1 Missing, not asked, etc                                                              | -2 = Not at all; -1 = A little; 0 = Moderately; 1 = Very; 2 = Extremely                                                                                                   | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Immigrants take jobs                     | How likely immigration will take away jobs                 | 1 = Extremely; 2 = Very; 3 = Somewhat; 4 = Not at all; \< 1 Missing, not asked, etc                                                                              | 1 = Not at all; 2 = Somewhat; 3 = Very; 4 = Extremely; NA = Missing, not asked, etc                                                                                       | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Worried about family finances            | Worry about family financial situation                     | 1 = Extremely; 2 = Very; 3 = Moderately; 4 = A little; 5 = Not at all; \< 1 Missing, not asked, etc                                                              | 1 = Not at all, 2 = A little; 3 = Moderately; 4 = Very; 5 = Extremely; NA = Missing, not asked, etc                                                                       | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Know someone lost job                    | Know someone who lost job                                  | 1 = Someone lost job; 2 = No one lost job; \< 1 Missing, not asked, etc                                                                                          | 0 = No one lost job; 1 = Someone lost job; NA = Missing, not asked, etc                                                                                                   | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| **Political**                            |                                                            |                                                                                                                                                                  |                                                                                                                                                                           |                  |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| New lifestyles breaking society          | Newer lifestyles breaking down society                     | 1 = Agree strongly; 2 = Agree somewhat; 3 = Neither; 4 = Disagree somewhat; 5 = Disagree strongly; \< 1 Missing, not asked, etc                                  | -2 = Disagree strongly; -1 = Disagree somewhat; 0 = Neither, 1 = Agree somewhat; 2 = Agree strongly                                                                       | 2012, 2016       |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Government biased against Whites         | Does the Administration favor Blacks or Whites (2012 only) | 1 = Favors Whites; 2 = Favors Blacks; 3 = Treats both the same; \< 1 Missing, not asked, etc                                                                     | -1 = Favors Whites; 0 = Treats both the same; 1 = Favors Blacks; NA = Missing, not asked, etc                                                                             | 2012, 2016, 2020 |
|                                          |                                                            |                                                                                                                                                                  |                                                                                                                                                                           |                  |
|                                          | Does the Federal Gov treat Blacks or Whites Better         |                                                                                                                                                                  |                                                                                                                                                                           |                  |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Whites influence politics                | How much influence do Whites have on U.S. politics         | 1 = Too much; 2 = Just about right; 3 = Too little; \< 1 Missing, not asked, etc                                                                                 | -1 = Too little; 0 = Just about right; 1 = Too much; NA = Missing, not asked, etc                                                                                         | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Whites are discriminated against         | Discrimination in U.S. against Whites                      | 1 = A great deal; 2 = A lot; 3 = A moderate amount; 4 = A little; 5 = None at all; \< 1 Missing, not asked, etc                                                  | 1 = None at all; 2 = A little; 3 = A moderate amount; 4 = A lot; 5 = A great deal; NA = Missing, not asked, etc                                                           | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| **Controls**                             |                                                            |                                                                                                                                                                  |                                                                                                                                                                           |                  |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Racial Resentment                        | Standard, 4-item, battery                                  | 5-point Likert scale                                                                                                                                             | Avg of four questions; NA = Missing, not asked, etc                                                                                                                       | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Party ID                                 | Standard battery                                           | 1 = Strong Democrat; 2 = Democrat; 3 = Leans Democrat; 4 = Independent; 5 = Leans Republican; 6 = Republican; 7 = Strong Republican; \< 1 Missing, not asked, etc| 1 = Strong Democrat; 2 = Democrat; 3 = Leans Democrat; 4 = Independent/Neither; 5 = Leans Republican; 6 = Republican; 7 = Strong Republican; \< 1 Missing, not asked, etc | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Female                                   | Sex of respondent                                          | 1 = Male, 2 = Female; \< 1 Missing, not asked, etc                                                                                                               | 0 = Male, 1 = Female; NA = Missing, not asked, etc                                                                                                                        | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
: ANES Measures {#tbl-anes-measures}

### 2022-10-01 code/dcr_anes_2012_cleaning.R

@@ -2,7 +2,7 @@
 
 # Notes: ----
     #* Description: Cleaning script for the 2012 ANES ----
-    #* Updated: 2022 - 02 - 07 ----
+    #* Updated: 2022-10-01 ----
     #* Updated by: dcr -----
 
 # Setup: -----
@@ -14,216 +14,149 @@ box::use(
     #* create list object
 anes_2012 = list()
     #* Load original ANES file ----
-anes_2012[['original']] = read_dta('data/anes-2012/anes-2012.dta')
+anes_2012[['original']] = read_dta('../data/anes-2012/anes-2012.dta')
 # Cleaning ----
 anes_2012[['clean']] = anes_2012[['original']] |>
     #* Filter dataset to include only white respondents ----
     filter(dem_raceeth_x == 1) |>
     mutate(
-    #* Female: ----
-        #** Coded as: gender_respondent_x - 1 male, 2 female
-        #** Recoded to: female - 0 male, 1 female
-        female = case_when(gender_respondent_x == 1 ~ 0,
-                            gender_respondent_x == 2 ~ 1),
-    #* Retrospective better off: ----
-        #** Coded as: finance_finpast_x - 1 much better, 5 much worse
-        #** Recoded to: rboff - -2 much worse, 2 much better
+    #* wid - White Identity
+        #** Coded as: ident_whiteid - 1 = Extremely Important - 5 Not at all important, < 1 Missing, not asked, etc
+        #** Recoded to: wid - 1 = Not at all important - 5 = Extremely important, NA = Missing, not asked, etc
+        wid = case_when(ident_whiteid == 1 ~ 5,
+                        ident_whiteid == 2 ~ 4,
+                        ident_whiteid == 3 ~ 3,
+                        ident_whiteid == 4 ~ 2,
+                        ident_whiteid == 5 ~ 1),
+    #* rboff - Retrospective better off: ----
+        #** Coded as: finance_finpast_x - 1 much better, 5 much worse, < 1 Missing, not asked, etc
+        #** Recoded to: rboff - -2 much worse, 2 much better, NA = missing, not asked, etc.
         rboff = ifelse(finance_finpast_x == 1, 2,
                         ifelse(finance_finpast_x == 2, 1,
                             ifelse(finance_finpast_x == 3, 0,
                                 ifelse(finance_finpast_x == 4, -1,
                                     ifelse(finance_finpast_x == 5, -2, NA))))),
[…]

### 2022-10-01 code/dcr_anes_2016_cleaning.R

@@ -2,259 +2,166 @@
 
 # Notes:
     #* Description: R script for cleaning the 2016 ANES 
-    #* Updated: 2022-07-19
+    #* Updated: 2022-10-01
     #* Updated by: dcr 
 
 # Setup
-    #* modularly load functions
-box::use(
-    haven = haven[read_dta],
-    dplyr = dplyr[filter, mutate, case_when, rename]
-)
     #* create list object
 anes_2016 = list()
     #* load dataset
-anes_2016[['original']] = read_dta('data/anes-2016/anes-2016.dta')
+anes_2016[['original']] = read_dta('../data/anes-2016/anes-2016.dta')
 
 # Cleaning
 anes_2016[['clean']] = anes_2016[['original']] |>
     #* Filter only white respondents
     filter(V161310x == 1) |>
     mutate(
-    #* rboff - Retrospective better off
-        #** 1 = Much Better - 5 Much Worse
-        #** RECODE TO: -2 = Much worse - 2 Much better
+    #* wid - White Identity
+        #** Coded as: V162327 - 1 = Extremely Important - 5 Not at all important, < 1 Missing, not asked, etc
+        #** Recoded to: wid - 1 = Not at all important - 5 = Extremely important, NA = Missing, not asked, etc
+    wid = ifelse(V162327 == 1, 5,
+            ifelse(V162327 == 2, 4,
+                ifelse(V162327 == 3, 3,
+                    ifelse(V162327 == 4, 2,
+                        ifelse(V162327 == 5, 1, NA))))),
+    #* rboff - Retrospective better off: ----
+        #** Coded as: V161110 - 1 much better, 5 much worse, < 1 Missing, not asked, etc
+        #** Recoded to: rboff - -2 much worse, 2 much better, NA = missing, not asked, etc.
     rboff = ifelse(V161110 == 1, 2,
                         ifelse(V161110 == 2, 1, 
                                ifelse(V161110 == 3, 0,
                                       ifelse(V161110 == 4, -1,
                                              ifelse(V161110 == 5, -2, NA))))),
-    #* pboff - Prospective better off
-        #** 1 = Much better - 5 = Much worse
-        #** RECODE TO: -2 = Much worse - 2 Much better
-    pboff = ifelse(V161111 == 1, 2,
-                        ifelse(V161111 == 2,1,
-                               ifelse(V161111 == 3,0,
-                                      ifelse(V161111 == 4, -1, 
[…]

### 2022-10-01 code/dcr_anes_2020_cleaning.r

@@ -2,121 +2,149 @@
 
 # Notes:
     #* Description: R Script to do the cleaning for the 2020 ANES 
-    #* Updated: 2022-08-12
+    #* Updated: 2022-10-01
     #* Updated by: dcr
 
 # Setup
-    #* Modularly load functions
-box::use(
-    haven = haven[read_dta],
-    dplyr = dplyr[filter, mutate,]
-)
     #* Create list object
 anes_2020 = list()
     #* Load dataset
-anes_2020[['original']] = read_dta('data/anes-2020/anes_2020_original.dta')
+anes_2020[['original']] = read_dta('../data/anes-2020/anes_2020_original.dta')
 
 # Clean the data
 anes_2020[['cleaned']] = anes_2020[['original']] |>
     #* Filter only white respondents
     filter(V201549x == 1) |>
     mutate(
-        #* rboff - Retrospective better off
-            #** 1 = Much better - 5 = Much worse
-            #** RECODE TO: -2 = Much worse - 2 = Much better
-        rboff = ifelse(V201502 == 1, 2,
-                    ifelse(V201502 == 2, 1,
-                        ifelse(V201502 == 3, 0,
-                            ifelse(V201502 == 4, -1,
-                                ifelse(V201502 == 5, -2, NA))))),
-        #* pboff - Prospective better off
-            #** 1 = Much better - 5 = Much worse
-            #** RECODE TO: -2 = Much worse - 2 = Much better
-        pboff = ifelse(V201503 == 1, 2,
-                    ifelse(V201503 == 2, 1,
-                        ifelse(V201503 == 3, 0,
-                            ifelse(V201503 == 4, -1,
-                                ifelse(V201503 == 5, -2, NA))))),
-        #* unpast - Unemployment better/worse than 1 year ago
-            #** 1 = Much better - 5 = Much worse
-            #** RECODE TO: -2 = Much worse - 2 = Much better
-        unpast = ifelse(V201333x == 1, 2,
-                    ifelse(V201333x == 2, 1,
-                        ifelse(V201333x == 3, 0,
-                            ifelse(V201333x == 4, -1,
-                                ifelse(V201333x == 5, -2, NA))))),
-        #* ecfamily - Worried about family finances
-            #** 1 = Extremely worried - 5 = Not at all worried
-            #** RECODE TO: -2 = Not at all - 2 = Extremely worried
-        ecfamily = ifelse(V201594 == 1, 2,
-                    ifelse(V201594 == 2, 1, 
[…]

### 2022-10-11 drafts/fa_2022/dcr_white_identity_relative_deprivation_fa_2022.qmd
pre-submission update... yay!

---
title: |
  Economic concerns appear to be weak predictors of white political identity
short-title: Economic concerns
published: "Manuscript draft: please do not share without author permission."
code-repo: "For replication, go to: <https:://github.com/DamonCharlesRoberts/white-identity-sources>."
author:
  - name: Damon C. Roberts
    email: damon.roberts-1@colorado.edu
    orcid: 0000-0002-4360-3675
    title: PhD Candidate
    affiliations:
      - id: CU
        name: University of Colorado Boulder
        department: Political Science
        address: 333 UCB
        city: Boulder
        region: CO 
        postal-code: 80309-0333
abstract: |
    Does economic or social and political threat explain white identity? Two leading arguments suggest that white identity results from a disruption of the racial status quo. However, they disagree on which disrupted element drives it. While one suggests that white identity comes from concern about whites' political influence, the other suggests that it is the result of concern with increased economic competition between whites and other racial groups. Considering the place that whiteness has in historical and contemporary institutions, I argue that white identity likely reflects concern about changes to their influence in the country's institutions. I use data from the American National Election Study that measure egotropic and sociotropic economic attitudes and leveraging questions gauging attitudes about social and political changes. Using a penalized regression, I examine whether proxies for economic threat appear to be plausible correlates of white identity. On balance, the models suggest that the hypothesis of economic threat is weaker than some suggest. I take this as evidence that white political identity, as it works today, is largely the result of whites' concerns with protecting the social and political status quo - and not so much economic.
thanks: |
    I would like to thank Jennifer Wolak for her feedback on the many drafts of this manuscript. I would also like to thank the Keller Center for the First Amendment for the research appointment that gave me the space to work on the early stages of this project. I would also like to thank Anand Sokhey and John Griffin for their advice and suggestions in the early drafts of this paper. I would also like to thank Tabitha Bonilla, Nicole Yadon, and participants and audience members of a Midwestern Political Science Association panel for which an earlier draft this idea was presented. I would also like to thank Tamar Malloy, Andrew Philips, Kathryn Schauer, Rachel O'Neal, Alexander Jensen, Komal Kaur, Andrew Engelhardt, and Madeline Mader for their encouragement and feedback at various stages of this project's development. All errors contained in this manuscript are ultimately mine.
additional-info: |
  I have no known conflict of interest to disclose.
keywords:
  - white identity
  - economic threat
  - cultural threat
date: today
bibliography: "../../assets/wid.bib"
format:
  hikmah-pdf: 
    biblatex-chicago: true
    biblio-style: authordate
    biblatexoptions:
      - backend=biber
      - autolang=hyphen
      - isbn=false
      - uniquename=false
linestretch: 2
execute:
  echo: false
  warning: false
  message: false
header-includes:  
- \usepackage{lscape}
---

```{r}
#| label: block-setup

# set seed
set.seed(90210)
# modularly load functions
box::use(
    dplyr = dplyr[select, filter, mutate, case_when, rename],
    tidyr = tidyr[drop_na],
    mice = mice[mice, pool],
    miceadds = miceadds[mids2datlist],
    brms = brms[brm, brm_multiple, set_prior, lasso, cumulative, prior],
    tibble = tibble[tribble],
    modelsummary = modelsummary[datasummary_skim, datasummary_correlation, modelsummary],
    parameters = parameters[parameters, standardize_names],
    ggplot2 = ggplot2[ggplot, aes, position_dodge, theme_minimal, geom_vline, labs, theme, element_text, scale_y_discrete],
    tidybayes = tidybayes[geom_pointinterval],
    performance = performance[check_model]
)
```

```{r}
#| label: cleaning-block

# source cleaning scripts
source('../../code/dcr_anes_2012_cleaning.R')
source('../../code/dcr_anes_2016_cleaning.R')
source('../../code/dcr_anes_2020_cleaning.R')

```

```{r}
#| label: load-data-block
# create list object for datasets
data = list()
# Create empty lasso list to store fitted models
lasso = list()
# Specify number of iterations for imputation
data_sets = 10
# load the cleaned data
    #* 2012
data[['anes_2012']] <- read.csv('../../data/anes-2012/anes-2012-updated.csv') |>
    select(wid, rboff, epast, unpast, edu, income, losejob, immtakejobs, ecfamily, ecjob, tradbreak, govbias, inflwhite, whitediscrim, raceResent, pid, female)
    #* 2016
data[['anes_2016']] <- read.csv('../../data/anes-2016/anes-2016-updated.csv') |>
    select(wid, rboff, epast, unpast, edu, income, losejob, immtakejobs, ecfamily, ecjob, tradbreak, govbias, inflwhite, whitediscrim, raceResent, pid, female)
    #* 2020
data[['anes_2020']] <- read.csv('../../data/anes-2020/anes-2020-updated.csv') |>
    select(wid, rboff, epast, unpast, edu, income, losejob, immtakejobs, ecfamily, ecjob, govbias, inflwhite, whitediscrim, raceResent, pid, female)
```

# Introduction

Those conducting post-mortem analyses of the 2016 Presidential election of Donald J. Trump, a presidential candidate that personifies white grievance politics, contend that the substantial presence of non-college-educated white supporters explains his electoral success [@griffin_et-al_2017]. Analyses by political scientists largely attribute Trump's success in attracting voters to those who see their whiteness as a politically relevant identity where they are distinct from Americans of other racial groups [@jardina_2020_pp]. For a long time, political scientists saw white political behavior and attitudes through the lens of prejudice. These analyses exclude another dimension of whites' racial attitudes: their affinity for, feelings of linked fate with other whites, and an expression that their race is a politically-relevant identity. More recently, scholars using the @tajfel_turner_1986 social psychological framework of identity challenge this tendency to focus on whites' prejudice by arguing that some whites may see their race as an identity essential and relevant to politics. As a result, while some may express out-group attitudes, they also express an in-group preference for other whites [@jardina_2019_cup]. Despite the importance of these attitudes, the source of these in-group preferences remains somewhat unclear.

There appears to be a mismatch between those who study sources of a variety of social identities and those that study white political identity. Scholars studying identity, more broadly, often focus on economic threats as a critical motivator for an expressed connection to a social in-group [see @huddy_2001]. However, some scholars studying white political attitudes and racial identity tend to focus on the role of political threats [see @bartels_2020]. This mismatch leads to some conceptualizing white identity as originating from economic views in the tradition of the identity literature. In contrast, others conceptualize white identity as coming from political and cultural threats.

I agree with those who focus on the connection between white political identity and political threats. Considering identity's historical and contextual roots [@trawalter_et-al_2022_cdps], I argue that white identifiers are concerned about their status in politics and society. Though I acknowledge that economic threat is a potent activation of identity and encourages engagement in group-based behavior, white identity holds decisive cultural and institutional significance in the U.S. [@rile-hayward_2013]. One's whiteness is not a descriptive characteristic, but it reflects a history of allowances by institutions and advantages in one's ability to influence politics. Though race and economics are undoubtedly correlated, we must be careful in specifying the data-generating process.

Along with providing a narrative that scholars might use to identify potential mechanisms connecting political threats to white identity, I examine the correlates of white identity. Using data from the 2012, 2016, and 2020 American National Election Study, I utilize penalized regression to test the claim that a handful of predictors that reflect attitudes about political power are more influential than general (sociotropic) and pocketbook (egotropic) attitudes about the economy. I find that proxies of political threat appear stronger than economic attitudes in the post-Trump era. As this particular analysis is not a causal test of the mechanism connecting threat to white identity, I close by arguing that scholars should continue to examine the mechanisms that explain this connection.

# Identity and white political identity

Social identity theory suggests that individuals form attitudes based on groups they associate with, even if those groups are subjectively defined [see @tajfel_turner_1986]. In the famous Robber's Cave experiment, @sherif_1988 divided a group of adolescent boys into two groups arbitrarily. After a week of spending time with others in the group, the boys came up with group names that defined the two groups and distinguished themselves; one group called themselves the Eagles, and the other group referred to themselves as the Rattlers. With the introduction of competition, these boys reported fierce hostility toward the members of the other group and expressed positive attitudes toward those within the group.

This example illustrates several features of social identities. The first feature of social identity theory is that these groups we may associate ourselves with are often subjectively defined [see @huddy_2001]. The second feature is that these subjective definitions of the group make distinctions between the two groups that come to matter to members of each group - some are part of our group, and some are part of that other group. Many cognitive theories suggest that this occurs as a result of the tendency for humans to reduce cognitive load by grouping things together [@huddy_2001]. The third feature is that introducing competition between these groups shapes attitudes toward those within and without your group. The motivator is that groups with higher status often want to keep that status, and individuals in the high-status group are much more likely to support inequality to satisfy that motivation [@burns_kinder_2012;@morrison_et-al_2009]. Fourth, individuals firmly hold these attitudes, which tend to be long-lasting. An additional feature of theories of social identity is that the degree to which one might identify with one group or another is fluid and that whether one identity that one might hold is active or if it is another often depends on the context and the identity's relevance to that context [@huddy_2012].

Racial groups in the U.S. contain a number of these features. As a result, a burgeoning literature in political science that has made this connection takes the racial identity of whites more seriously as one that informs their political attitudes [see @jardina_2019_cup]. Race in America is subjectively defined [see @rile-hayward_2013], and these definitions are deeply embedded and reinforced by our institutions [@lopez_2006]. Whiteness and one's willingness to identify with that group is important [@wong_cho_2005] as it reflects a desire for an association with the history and affordances of being white [@mcdermott_samson_2005].

In many eras of American history, there are several instances where we have sought to reclassify who is white. During the antebellum period, whiteness was no longer simply one's economic class. However, now it had become a reflection of a racial hierarchy that afforded some various rights and excluded others from enjoying the same rights (e.g., literacy tests). During this era, rather than conceptualizing whites as landowners of Caucasian descent, the definition of what was white shifted to include those previously labeled non-white Europeans, such as the Irish. Changes to definitions of who is white illustrate that the primary motivator was race and not class as to who holds some rights and who does not [@rile-hayward_2013].

More contemporaneously, whites hold several strong political attitudes racialized by politicians. Whites' attitudes toward welfare programs are not too distinct from those toward Black Americans [@gilens_1999]. A provided explanation is that social welfare policies are more unpopular among whites when the media frames these policies as benefiting non-whites [@winter_2008;@van-doorn_2015]. Explicit cues reminding individuals of the racial hierarchy entrench the recipients' attitudes toward Black Americans [@lyle_2014; see also @wellman_et-al_2016]. Conservatives argued that Barack Obama, the first Black president, expressed support for the Affordable Care Act as a way to disproportionately help Black Americans [@tesler_2012;@tesler_2016]. These messages corresponded with a rise of survey reports expressing old-fashioned racism among whites during his administration [@tesler_2013]. While whites tend to report negative views about government aid for those in need when they are non-white recipients, evidence suggests that whites who live in communities with a higher proportion of poor whites report much more positive views of the poor than those that perceive the poor as mostly non-white [@hopkins_2009]. Additionally, telling whites that other whites will be the beneficiaries of an entitlement program increases their support for such a program [@wetts_willer_2018].

As an identity, much of the literature view these attitudes as coming from the presence or focus on economic threats. Some studying white political identity find economic predictors as substantively and statistically significant [see @perez_et-al_2022_pp]. From non-experimental evidence, others make similar arguments about what predicts white political identity. @bunyasi_2019 provides evidence suggesting that Trump's messages highlighting increased job market competition harming whites because of affirmative action was a successful strategy and helped build his base of support. After all, the identity literature, more broadly, often focuses on economic threats. Evolutionary psychologists often take the position that identity and the resulting attitudes and behaviors directed toward out-groups are the result of competition over material resources [see @huddy_2001]. Many political scientists use this as an explanation for negative attitudes directed toward immigrants among those who feel insecure about their job prospects [@pardos-prado_xena_2019]. Others argue that deindustrialization and job loss activate whites' concerns about the racial hierarchy and predict their support for Republican challengers that express support for keeping the racial status quo [@baccini_weymouth_2021_apsr].

Not all agree, however. Some suggest that the negative attitudes that whites express toward immigrants result from concerns about cultural changes in those living in border towns [@hainmueller_hopkins_2014]. Others suggest that simply reminding whites of these demographic shifts activates these racial identity-based attitudes [@craig_richeson_2014]. Some other evidence suggests that these identity-based attitudes are better predicted by ethnic antagonism [@bartels_2020] and by racial animus [@fahey_et-al_2020] and explicitly take the position that a reaction to economic threats does not drive these. However, these attitudes tend to focus on prejudiced views rather than a reflection of a connection to other whites.

# Is white identity non-material?

As much of the literature on whites' racial attitudes rely heavily on social identity theory as a paradigm, a significant portion of it presents the source of these group-based views as responses to economic threat. However, some disagree with this position [see @jardina_2019_cup] and view these attitudes as the result of changes to the racial hierarchy. These other perspectives provide explanations for whites' attitudes and conceptualize the definition of whiteness and white identity as a desire to retain political power.

Historical accounts that contextualize attitudes stemming from white racial identity seem to suggest that, though economics correlates with racial attitudes, what it meant to be white was no longer about class. However, it became about what it meant for access and privileges afforded by institutions [@rile-hayward_2013]. In the $19^{th}$ century, privilege came from one's freedom and wealth. Those who were not enslaved but were poor were not considered white and were of a lower strata than whites who owned land [@rile-hayward_2013]. With the abolition of slavery, the expressed (explicit) concern became not about wealth but now about race. This motivated a shift in how the institutions and how whites defined themselves. It became the focus for whites to be a higher status group, regardless of whether or not they were poorer than formerly enslaved Black Americans [@rile-hayward_2013].

Many conceptualizations of white political identity align with its primary source as concern for the group's political power. While a critical component of identity is a feeling of your fate as linked with others in your group [@leach_et-al_2008], this linked fate is the connection between attitudes and material forms of threat but that identity, in and of itself, is better conceptualized as a result of cultural or ideational threat [@donnelly_2020]. Aligning with the contextual account of racial identity for whites, many conceptualize white identity as an expressed desire for a recognition of privilege and status by institutions [@lopez_2006] and a connection with the history that connects whites to their status [@mcdermott_samson_2005]. The prominent account of white political identity and its implications in contemporary American politics argues for a strong connection between white political identity and a desire to protect this racial hierarchy and status. This conceptualization reflects the tendency to report in-group-oriented, not necessarily out-group-oriented, attitudes [@jardina_2019_cup].

As this is an open question and is rather a complex one, I attempt to wade into this by parsing through which of these two perspectives are more plausible. As I discuss in the concluding section, I argue that scholars should continue taking the conceptualization and measurement of white political identification seriously and examine the mechanisms that connect economic and political threats to white political identity. In the next section, I examine whether attitudes expressing concern with economic or political threats are more robust correlates of the degree to which someone feels that being white is important to their identity. The goal is not to test particular causal mechanisms but to examine whether my belief and others' that white identity reflects reactions to political threats more than it does economic.

# Data

To examine whether whites' attitudes expressing concern about political threats are more robust correlates than their perceptions of the economy, I use the 2012, 2016, and 2020 American National Election Studies (ANES). The studies are nationally representative questions that capture several questions that lend themselves to testing the relationship between economic and political concerns while also asking white respondents to report how important being white is to their identity.

Of course, I cannot *prove* that these questions reflect the two types of threats I purport to represent. This leaves any evidence of the particular mechanism connecting these predictors to reported levels of white identity -- among other reasons -- suspect. However, the goal of this paper is not necessarily to test a particular causal mechanism -- this is a much larger project that we should continue to work on collectively. Nevertheless, the paper aims to offer an argument that contextualizes white identity in history and how that leads to different predictions about the correlates of white identity than many might assume from understanding the literature on the origins of social identity.

Each study asks white respondents, "How important is being white to your Identity" and offers them a five-point Likert scale with the options reflecting their level of agreement with the question ^[Details about question-wording, original ANES coding, and my coding for all variables I include in my models are included in @tbl-anes-measures located in the Appendix. I include descriptive statistics for the measures in these samples in @tbl-2012-descriptive-stats, @tbl-2016-descriptive-stats, and @tbl-2020-descriptive-stats in the Appendix]. They do not ask white respondents to report on items from the white political identity measures of @jardina_2019_cup or the more complex battery of items that @helms_carter_1990 discusses. However, this question does capture the degree to which one sees their whiteness as part of who they are and reflects the conceptualization that others have used [see @schildkraut_2017;@jardina_2019_cup;@perez_et-al_2022_pp].

The ANES does not ask questions that explicitly capture the attitudes of respondents about perceptions of the economy for different groups. That is, there are no questions asking respondents about their perceptions of the economy for Black Americans relative to white Americans. The literature suggests, however, that those who have lower incomes are more likely to feel insecure about their financial status and are more likely to use this economic threat as a motivation for their negative attitudes toward immigrants [@pardos-prado_xena_2019]. Therefore, one measure of economic threat that I use is the respondent's reported level of total family income. Those with lower levels of education are also likely to feel similar levels of insecurity about their economic status, as their ability to seek higher-paying jobs is likely more limited. Whites who report that they feel that their family is doing worse relative to the previous year and knowing others who lost a job also reflect a degree of insecurity about their economic status and may reflect perceptions of economic threat [@jardina_2019_cup]. I include an item reflecting respondents' perceptions of economic mobility today relative to the past. In 2012, whites harboring negative attitudes toward Obama rooted in racial animous were more likely to express more negative views of the economy relative to how it had been in the past [@sides_et-al_2018]. Others demonstrate that those who are reacting to economic threats and are concerned about their and other whites' status evaluate the economy and their economic position as worse [@baccini_weymouth_2021_apsr]. Therefore, I additionally include questions about retrospective sociotropic evaluations of the economy.

The ANES includes several questions that reflect perceptions of whites losing political power. First, I include a question that asks respondents to report whether or not they felt that whites currently have too little or too much influence on politics today. I also want to examine the degree to which respondents feel that the institutions benefit non-whites rather than whites by including the question that whites face discrimination and that the government is biased against whites. I include a question asking respondents about their belief that traditions in this country are breaking^[The ANES only included this question in the 2012 and 2016 questionnaires]. While these, like the economic attitudes, are just proxies of threat, they reflect the extent to which respondents may feel insecure about the political influence of whites.

Furthermore, the ANES includes the racial resentment battery, which asks questions reflecting whites' beliefs that Blacks are individually responsible for their circumstances and undeserving of assistance. I control for racial resentment as the individual items likely confound the relationship between both economic and political threat measures on expressed white identity. As economic evaluations often reflect partisans reporting a better economy when the government is run by co-partisans and a worse economy when the other party runs the government [@achen_bartels_2017], I additionally control for a 7-point item measure of partisanship. Finally, I control for whether or not the respondent is female, as Sex affects one's socioeconomic status.

## Evidence from a penalized regression

The summary of the literature points to a large number of potential predictors that influence the degree to which one takes on white political identity. To adjudicate between hypotheses, there is value in a comprehensive model that considers the plausibility of each. Most models consider those predictors they believe are most plausible and do not include others that the researcher believes to be less plausible [see @clarke_primo_2012]. They understandably do this out of concern with increases in the inefficiency of one's regression estimates: the more predictors you include, the risk of overfitting your model to your particular sample and from multicollinearity increases, thereby erroneously generalizing conclusions drawn from the fitted statistical model. Though this concern is reasonable, such a model neglects to examine the effects of the predictors a researcher believes to be less plausible.

Therefore, I use a penalized regression model to balance these concerns. The literature surveyed above suggests that there are a small number of variables that do predict white political identity, and there are many others that have no effect. Imagining an assumed hypothesis about the distribution of coefficients, this implies that the L2 regularization parameter equals 0. This is because the L2 regularization parameter shrinks coefficients asymptotically toward zero. The L1 regularization parameter, however, does allow for a shrinkage parameter to pull coefficients to 0. Given this, it suggests that the more appropriate model to capture this debate would be the LASSO penalized regression.

One option for implementation is to use the "classical" LASSO, which relies on an optimization procedure through cross-validation. It is a frequentist approach to LASSO penalized regression. However, there are several positive benefits to the second implementation of the LASSO penalized regression in a Bayesian framework [see @van-erp_et-al_2019_jmp]. From my review of the literature, I have \textit{apriori} expectations for the distribution of effects on my predictors in the form of a Laplace distribution which I explicitly define in my prior density function. I also have the additional benefit of having the ability to define its hyperparameters. Meaning I suspect that there is a peak around zero and that there are relatively large tails. This characterization of my prior fits with that of the LASSO - a large number of included predictors have zero effect, with a relatively smaller proportion of included predictors that do have an effect but are relatively unsure as to how large those coefficients might be. I can specify the scale of that uncertainty by specifying a hyperparameter for this LASSO prior. Without having to perform cross-validation as I would with the classical LASSO, my posterior distribution naturally reflects the uncertainty around the size of my L2 shrinkage parameter and can estimate models with different L2 shrinkage parameters simultaneously. As $\verb|stan|$ includes implementations of this model, I rely on Hamiltonian Monte Carlo, a type of Markov Chain Monte Carlo sampling which allows for non-convexity in the L2 shrinkage parameters. As classical LASSO models are challenging to interpret due to cross-validation, using credible intervals offers a much less error-prone interpretation.

My measure of white identity is ordinal. Though white identity as a concept is likely continuous, through the specification of a cumulative density function with a logit link, my model bins this latent continuous space based on the number of response options. Though it creates a more complicated computational task, it is a more appropriate model specification given the available measure and my assumptions about the concept [@liddell_kruschke_2018_jesp].

To account for potential problems stemming from missing data, I impute my data by using multiple imputation with chained equations (MICE) where the particular model I use is a random forest model to allow for more flexibility and to reduce model dependence [@buuren_goothuis-oudshoorn_2011; @marbach_2022_pa]^[I impute ```r sprintf('%.0f', data_sets)``` datasets and simultaneously fit my Bayesian LASSO on each dataset and pool the results. Uncertainty reflected in my reported credible intervals not only reflects the uncertainty from one model, but my reported uncertainty also reflects the uncertainty generated from the imputation procedure. One drawback with this particular approach is that my normalized split-$\hat{R}$ will likely be greater than the widely recognized 1.01 cutoff that indicates model convergence due to the pooling of my multiply imputed datasets.]. I report the results of the model fitted using listwise deletion in @tbl-2012-lasso-simple, @tbl-2016-lasso-simple, and @tbl-2020-lasso-simple.

```{r}
#| label: imputation-block
#| eval: false
# Imputation
    #* 2012
data[['anes_2012_imputed']] <- mice(data[['anes_2012']], m = data_sets, method = 'rf')
saveRDS(data[['anes_2012_imputed']], file = '../data/anes-2012/anes-2012-imputed.rds')
    #* 2016
data[['anes_2016_imputed']] <- mice(data[['anes_2016']], m = data_sets, method = 'rf')
saveRDS(data[['anes_2016_imputed']], file = '../data/anes-2016/anes-2016-imputed.rds')
    #* 2020
data[['anes_2020_imputed']] <- mice(data[['anes_2020']], m = data_sets, method = 'rf')
saveRDS(data[['anes_2020_imputed']], file = '../data/anes-2020/anes-2020-imputed.rds')
```

```{r}
#| label: imputation-loading-block

data[['anes_2012_imputed']] <- readRDS('../..//data/anes-2012/anes-2012-imputed.rds')

data[['anes_2016_imputed']] <- readRDS('../../data/anes-2016/anes-2016-imputed.rds')

data[['anes_2020_imputed']] <- readRDS('../../data/anes-2020/anes-2020-imputed.rds')
```

## Results

@fig-2012-2016-2020-ordinal-lasso presents the estimates of my fitted ordinal models with the LASSO shrinkage prior^[@tbl-2012-2016-2020-ordinal-lasso presents the table of results.]. These models are fit on 6 chains with 2000 iterations each. To examine the evidence of whether a particular predictor "matters" for predicting white political identity, I examine the estimated credible intervals, which report the probability that the true estimate falls within the estimated range. I construct these credible intervals by reporting estimates between the 2.5% and 97.5% quantiles. This means that for a given credible interval, there is a two-tailed 95% probability that the true value falls within that range. For values outside the range of the credible interval, it is relatively implausible that the given value is the true value.

```{r}
#| label: models-block
#| eval: false

# Models
    #* 2012
lasso[['2012_simple']] <- brm(wid ~ rboff + epast + unpast + edu + income + losejob + ecfamily + ecjob + tradbreak + govbias + whitediscrim + raceResent + pid + female, data = data[['anes_2012']], prior = set_prior(lasso(1)), chains = 6, iter = 2000, family = cumulative(link = 'logit'))
saveRDS(lasso[['2012_simple']], file = '../data/lasso_2012_simple.rds')

lasso[['2012_ordinal']] <- brm_multiple(wid ~ rboff + epast + unpast + edu + income + losejob +  ecfamily + ecjob + tradbreak + govbias + whitediscrim + raceResent + pid + female, data = data[['anes_2012_imputed']], prior = set_prior(lasso(1)), chains = 6, iter = 2000, family = cumulative(link = 'logit'))
saveRDS(lasso[['2012_ordinal']], file = '../data/lasso_2012_ordinal.rds')
    #* 2016
lasso[['2016_simple']] <- brm(wid ~ rboff + epast + unpast + edu + income + losejob + ecfamily + ecjob + tradbreak + govbias + whitediscrim + raceResent + pid + female, data = data[['anes_2016']], prior = set_prior(lasso(1)), chains = 6, iter = 2000, family = cumulative(link = 'logit'))
saveRDS(lasso[['2016_simple']], file = '../data/lasso_2016_simple.rds')

lasso[['2016_ordinal']] <- brm_multiple(wid ~ rboff + epast + unpast + edu + income + losejob + ecfamily + ecjob + tradbreak + govbias + whitediscrim + raceResent + pid + female, data = data[['anes_2016_imputed']], prior = set_prior(lasso(1)), chains = 6, iter = 2000, family = cumulative(link = 'logit'))
saveRDS(lasso[['2016_ordinal']], file = '../data/lasso_2016_ordinal.rds')
    #* 2020
lasso[['2020_simple']] <- brm(wid ~ rboff + epast + unpast + edu + income + losejob + ecfamily + ecjob + govbias + whitediscrim + raceResent + pid + female, data = data[['anes_2020']], prior = set_prior(lasso(1)), chains = 6, iter = 2000, family = cumulative(link = 'logit'))
saveRDS(lasso[['2020_simple']], file = '../data/lasso_2020_simple.rds')

lasso[['2020_ordinal']] <- brm_multiple(wid ~ rboff + epast + unpast + edu + income + losejob + ecfamily + ecjob + govbias + whitediscrim + raceResent + pid + female, data = data[['anes_2020_imputed']], prior = set_prior(lasso(1)), chains = 6, iter = 2000, family = cumulative(link = 'logit'))
saveRDS(lasso[['2020_ordinal']], file = '../data/lasso_2020_ordinal.rds')
```

```{r}
#| label: loading-fitted-models-block

# 2012
lasso[['2012 LWD']] <- readRDS('../../data/lasso_2012_simple.rds') 
lasso[['2012']] <- readRDS('../../data/lasso_2012_ordinal.rds') 
# 2016
lasso[['2016 LWD']] <- readRDS('../../data/lasso_2016_simple.rds') 
lasso[['2016']] <- readRDS('../../data/lasso_2016_ordinal.rds')
# 2020
lasso[['2020 LWD']] <- readRDS('../../data/lasso_2020_simple.rds')
lasso[['2020']] <- readRDS('../../data/lasso_2020_ordinal.rds')
```

```{r}
#| label: fig-2012-2016-2020-ordinal-lasso
#| fig-cap: Economic factors are weak predictors of white identity
#| fig-height: 9
#| fig-width: 9
# Graphs
    #* Posterior draws
lasso[['2012_ordinal_tidy']] <- parameters(lasso[['2012']], conf.int = 0.95, verbose = FALSE) |>
    standardize_names(style = 'broom') |>
    data.frame() |>
    mutate(model = '2012') |>
    filter(!is.na(term))

lasso[['2016_ordinal_tidy']] <- parameters(lasso[['2016']], conf.int = 0.95, verbose = FALSE) |>
    standardize_names(style = 'broom') |>
    data.frame() |>
    mutate(model = '2016') |>
    filter(!is.na(term))

lasso[['2020_ordinal_tidy']] <- parameters(lasso[['2020']], conf.int = 0.95, verbose = FALSE) |>
    standardize_names(style = 'broom') |>
    data.frame() |>
    mutate(model = '2020') |>
    filter(!is.na(term))

lasso[['ordinal_tidy']] <- rbind(lasso[['2012_ordinal_tidy']], lasso[['2016_ordinal_tidy']], lasso[['2020_ordinal_tidy']]) |>
  mutate(
    term = case_when(term == 'b_rboff' ~ 'Retrospective better off', term == 'b_epast' ~ 'Economy better', term == 'b_unpast' ~ 'Unemployment better', term == 'b_edu' ~ 'Education', term == 'b_income' ~ 'Income', term == 'b_losejob' ~ 'Worried about losing job', term == 'b_ecfamily' ~ 'Worried about family finances', term == 'b_ecjob' ~ 'Know someone lost job', term == 'b_tradbreak' ~ 'New lifestyles breaking society', term == 'b_govbias' ~ 'Gov. biased against whites', term == 'b_inflwhite' ~ 'whites influence politics', term == 'b_whitediscrim' ~ 'whites are discriminated against', term == 'b_raceResent' ~ 'Racial resentment', term == 'b_pid' ~ 'Party ID', term == 'b_female' ~ 'Female'),
    term = factor(term, levels = c('Retrospective better off', 'Economy better', 'Unemployment better', 'Education', 'Income', 'Worried about losing job', 'Worried about family finances', 'Know someone lost job', 'New lifestyles breaking society', 'Gov. biased against whites', 'whites influence politics', 'whites are discriminated against', 'Racial resentment', 'Party ID', 'Female'))
  ) |>
  filter(!is.na(term))

ggplot(data = lasso[['ordinal_tidy']], aes(y = term, x = estimate, xmin = conf.low, xmax = conf.high, linetype = factor(model), shape = factor(model))) +
    geom_pointinterval(position = position_dodge(width = 0.4), size = 3) +
    geom_vline(xintercept = 0, linetype = 2) +
    scale_y_discrete(limits = rev) +
    theme_minimal() +
    labs(linetype = 'Model', shape = 'Model', x = 'Estimate', y = '', caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.') + 
    theme(text=element_text(size=12, family = 'sans'))
```

Examining the figure, it is readily apparent that in the face of indicators of attitudes about political status, economic attitudes inconsistently associate with white political identity. In all three years, reported attitudes about the nation's economy being better relative to the previous year, reported feelings about the nation's unemployment rate, income, concern about one's family finances, and knowing someone who lost a job are not meaningfully correlated with white political identity. For other indicators of economic threat, none appear to be consistently meaningful. Retrospective evaluations of one's financial situation and education matter in 2012, but these effects were not present in 2016 and 2020.

Turning to indicators of political threat, we still see significant inconsistency. In both 2012 and 2016, whether someone feels that new lifestyles are breaking society positively affects reported levels of white political identity. This relationship is meaningful in 2012 and 2016 (this question is not present in the 2020 questionnaire). In 2020, those who reported that the government favored Black Americans also reported a higher white political identity, and those who feel that whites are discriminated against; however, neither of these effects are meaningful in 2012 or 2016.

## Discussion

Overall, these results suggest that the effects of economic threat on white political identity are more robust in the pre-Trump era. Again, this is not confirmatory as there is no test of a causal mechanism for either economic or political threats. The goal of this analysis is to examine whether, in an omnibus test, either of these threats appears to be a more plausible predictor of white political identity. Though the relationship between political threats is inconsistent, their relationships appear more likely; particularly when we consider how many of these attitudes are explained by a rather complex data-generating process.

Economic threats tend to play a more significant part in explaining white identification in 2012. Political threats, however, appear to be a stronger predictor of white identification in the post-Trump era. In 2012, the U.S. economy was recovering from a massive recession and housing market collapse. However, as Obama's administration continued, conservatives became more outspoken about Obama's race [@tesler_2016]. This shift to focus on Obama's race continued through the 2016 Presidential election and may reflect a much more explicit message to whites about political threats [@sides_et-al_2018]. Though in 2020, the global economy faced significant uncertainty due to the COVID-19 pandemic, economic factors appear to be relatively less important. These results suggest that white political identity results from a political threat in the post-Trump era.

# Conclusion

The literature on social identity suggests that the centrality of one's identity and group-based attitudes and behaviors result from economic threat. However, white political identity has a historical and institutional significance that, while tied up in economics, reacts to changes in whites' influence over these institutions.

As pitting two theories against one another is a methodological challenge, I use data from the 2012, 2016, and 2020 ANES to demonstrate whether the relationships between attitudes proxying political and economic threats are more plausible. While this exercise is not and should not be interpreted as causal, I do demonstrate that economic factors appear to have a weaker relationship with white identity than political attitudes - particularly in the post-Trump era.

The narrative I advocate for suggests that white political identity is most responsive to threats of whites' institutional influence. The descriptive analysis of the relationships between political and economic threat fall short of causal identification. With the results of these analyses, scholars have evidence suggesting that economic threats appear to weakly associate with one's expression of a white political identity when including political threats in the same model.

The task for scholars is to examine whether political threats have any causal bearing on explaining the source of white political identity. This undoubtedly is a difficult task as threats are hard to treat experimentally and are reasonably suspect with observational data.

{{< pagebreak >}}
<!--
TODO:
  - Make a strong assertion about my results
  - Discuss why I think economic factors don't matter when they matter in most identity papers
  - What does this mean for how we understand white identity? Elaborate on the takeaway of the paper and what I added but also what questions are still there.
-->

# Appendix{#sec-appendix}

## Measures{#sec-measures}

+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
|                                          | Abbreviated question                                       | ANES Coding                                                                                                                                                      | Author coding                                                                                                                                                             | Years            |
+==========================================+============================================================+==================================================================================================================================================================+===========================================================================================================================================================================+==================+
| white identity                           | How important is being white to your identity              | 1 = Extremely important; 2 = Very important; 3 = Moderately important; 4 = A little important; 5 = Not at all important; \<1 Missing, not asked, etc             | 1 = Not at all important; 2 = A little important; 3 = Moderately important; 4 = Very important; 5 = Extremely important; NA = Missing, not asked, etc                     | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| **Economics**                            |                                                            |                                                                                                                                                                  |                                                                                                                                                                           |                  |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Retrospective better off                 | How much better worse off than 1 year ago                  | 1 = Much better; 2 = Somewhat better; 3 = The same; 4 = Somewhat worse; 5 = Much worse, \< 1 Missing, not asked, etc                                             | -2 = Much worse; -1 = Somewhat worse; 0 = The same; 1 = Somewhat better; 2 = Much better; NA = Missing, not asked, etc                                                    | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Economy better                           | U.S. economy better worse off than 1 year ago              | 1 = Much better; 2 = somewhat better; 3 = Stayed about the same, 4 = Somewhat worse; 5 = Much worse; \< 1 Missing, not asked, etc                                | -2 = Much worse; -1 = Somewhat worse; 0 = The same; 1 = Somewhat better; 2 = Much better; NA = Missing, not asked, etc                                                    | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Unemployment better                      | Unemployment better or worse than 1 year ago               | 1 = Much better; 2 = somewhat better; 3 = Stayed about the same, 4 = Somewhat worse; 5 = Much worse; \< 1 Missing, not asked, etc                                | -2 = Much worse; -1 = Somewhat worse; 0 = The same; 1 = Somewhat better; 2 = Much better; NA = Missing, not asked, etc                                                    | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Education                                | Educational attainment                                     | 1 = \< than High school; 2 = High school; 3 = Some post-High school; 4 = Bachelor's degree; \< 1 Missing, not asked, etc                                         | 1 = \< than High school; 2 = High school; 3 = Some post-High school; 4 = Bachelor's degree; NA = Missing, not asked, etc                                                  | 2012, 2020       |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Education (2016)                         | Educational attainment                                     | See ANES codebook                                                                                                                                                | See ANES codebook; NA = MIssing, not asked, etc                                                                                                                           | 2016             |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Income                                   | Total family income                                        | See ANES codebook; \< Missing, not asked, etc                                                                                                                    | See ANES codebook; NA = MIssing, not asked, etc                                                                                                                           | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Worried about losing job                 | How worried about losing job in near future                | 1 = Not at all; 2 = A little; 3 = Moderately; 4 = Very; 5 = Extremely; \< 1 Missing, not asked, etc                                                              | -2 = Not at all; -1 = A little; 0 = Moderately; 1 = Very; 2 = Extremely                                                                                                   | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Immigrants take jobs                     | How likely immigration will take away jobs                 | 1 = Extremely; 2 = Very; 3 = Somewhat; 4 = Not at all; \< 1 Missing, not asked, etc                                                                              | 1 = Not at all; 2 = Somewhat; 3 = Very; 4 = Extremely; NA = Missing, not asked, etc                                                                                       | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Worried about family finances            | Worry about family financial situation                     | 1 = Extremely; 2 = Very; 3 = Moderately; 4 = A little; 5 = Not at all; \< 1 Missing, not asked, etc                                                              | 1 = Not at all, 2 = A little; 3 = Moderately; 4 = Very; 5 = Extremely; NA = Missing, not asked, etc                                                                       | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Know someone lost job                    | Know someone who lost job                                  | 1 = Someone lost job; 2 = No one lost job; \< 1 Missing, not asked, etc                                                                                          | 0 = No one lost job; 1 = Someone lost job; NA = Missing, not asked, etc                                                                                                   | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| **Political**                            |                                                            |                                                                                                                                                                  |                                                                                                                                                                           |                  |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| New lifestyles breaking society          | Newer lifestyles breaking down society                     | 1 = Agree strongly; 2 = Agree somewhat; 3 = Neither; 4 = Disagree somewhat; 5 = Disagree strongly; \< 1 Missing, not asked, etc                                  | -2 = Disagree strongly; -1 = Disagree somewhat; 0 = Neither, 1 = Agree somewhat; 2 = Agree strongly                                                                       | 2012, 2016       |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Government biased against whites         | Does the Administration favor Blacks or whites (2012 only) | 1 = Favors whites; 2 = Favors Blacks; 3 = Treats both the same; \< 1 Missing, not asked, etc                                                                     | -1 = Favors whites; 0 = Treats both the same; 1 = Favors Blacks; NA = Missing, not asked, etc                                                                             | 2012, 2016, 2020 |
|                                          |                                                            |                                                                                                                                                                  |                                                                                                                                                                           |                  |
|                                          | Does the Federal Gov treat Blacks or whites Better         |                                                                                                                                                                  |                                                                                                                                                                           |                  |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| whites influence politics                | How much influence do whites have on U.S. politics         | 1 = Too much; 2 = Just about right; 3 = Too little; \< 1 Missing, not asked, etc                                                                                 | -1 = Too little; 0 = Just about right; 1 = Too much; NA = Missing, not asked, etc                                                                                         | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| whites are discriminated against         | Discrimination in U.S. against whites                      | 1 = A great deal; 2 = A lot; 3 = A moderate amount; 4 = A little; 5 = None at all; \< 1 Missing, not asked, etc                                                  | 1 = None at all; 2 = A little; 3 = A moderate amount; 4 = A lot; 5 = A great deal; NA = Missing, not asked, etc                                                           | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| **Controls**                             |                                                            |                                                                                                                                                                  |                                                                                                                                                                           |                  |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Racial Resentment                        | Standard, 4-item, battery                                  | 5-point Likert scale                                                                                                                                             | Avg of four questions; NA = Missing, not asked, etc                                                                                                                       | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Party ID                                 | Standard battery                                           | 1 = Strong Democrat; 2 = Democrat; 3 = Leans Democrat; 4 = Independent; 5 = Leans Republican; 6 = Republican; 7 = Strong Republican; \< 1 Missing, not asked, etc| 1 = Strong Democrat; 2 = Democrat; 3 = Leans Democrat; 4 = Independent/Neither; 5 = Leans Republican; 6 = Republican; 7 = Strong Republican; \< 1 Missing, not asked, etc | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Female                                   | Sex of respondent                                          | 1 = Male, 2 = Female; \< 1 Missing, not asked, etc                                                                                                               | 0 = Male, 1 = Female; NA = Missing, not asked, etc                                                                                                                        | 2012, 2016, 2020 |
+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
: ANES Measures {#tbl-anes-measures}

{{< pagebreak >}}

## Descriptive statistics {#sec-desc-stats}

```{r}
#| label: tbl-2012-descriptive-stats
#| tbl-cap: 2012 ANES Descriptive Statistics

data[['anes_2012']] |>
  rename(`white identity` = wid, `Retrospective better off` = rboff, `Economy better` = epast,    `Unemployment better` = unpast, `Education` = edu, `Income` = income, `Worried about losing job` = losejob, `Immigrants take jobs` = immtakejobs, `Worried about family finances` = ecfamily, `Know someone lost job` = ecjob, `New lifestyles breaking society` = tradbreak, `Government biased against whites` = govbias, `whites influence politics` = inflwhite, `whites are discriminated against` = whitediscrim, `Racial resentment` = raceResent, `Party ID` = pid, `Female` = female) |>
  datasummary_skim(notes = c('NAs are included in Unique column.', 'Data source: 2012 American National Election Study, unweighted.'))
```

```{r}
#| label: tbl-2016-descriptive-stats
#| tbl-cap: 2016 ANES Descriptive Statistics

data[['anes_2016']] |>
  rename(`white identity` = wid, `Retrospective better off` = rboff, `Economy better` = epast,    `Unemployment better` = unpast, `Education` = edu, `Income` = income, `Worried about losing job` = losejob, `Immigrants take jobs` = immtakejobs, `Worried about family finances` = ecfamily, `Know someone lost job` = ecjob, `New lifestyles breaking society` = tradbreak, `Government biased against whites` = govbias, `whites influence politics` = inflwhite, `whites are discriminated against` = whitediscrim, `Racial resentment` = raceResent, `Party ID` = pid, `Female` = female) |>
  datasummary_skim(notes = c('NAs are included in Unique column.', 'Data source: 2016 American National Election Study, unweighted.'))
```

```{r}
#| label: tbl-2020-descriptive-stats
#| tbl-cap: 2020 ANES Descriptive Statistics

data[['anes_2020']] |>
  rename(`white identity` = wid, `Retrospective better off` = rboff, `Economy better` = epast,    `Unemployment better` = unpast, `Education` = edu, `Income` = income, `Worried about losing job` = losejob, `Immigrants take jobs` = immtakejobs, `Worried about family finances` = ecfamily, `Know someone lost job` = ecjob, `Government biased against whites` = govbias, `whites influence politics` = inflwhite, `whites are discriminated against` = whitediscrim, `Racial resentment` = raceResent, `Party ID` = pid, `Female` = female) |>
  datasummary_skim(notes = c('NAs are included in Unique column.', 'Data source: 2020 American National Election Study, unweighted.'))
```

{{< pagebreak >}}

## Model results
```{r}
#| label: model-table-setup-block
    #* Set up non-coefficient related information from models
gm <- list(
    list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)
)

cm <- c('b_rboff' = 'Retrospective better off', 'b_epast' = 'Economy better', 'b_unpast' = 'Unemployment better', 'b_edu' = 'Education', 'b_income' = 'Income', 'b_losejob' = 'Worried about losing job', 'b_ecfamily' = 'Worried about family finances', 'b_ecjob' = 'Know someone lost job', 'b_tradbreak' = 'New lifestyles breaking society', 'b_govbias' = 'Gov. biased against whites', 'b_inflwhite' = 'whites influence politics', 'b_whitediscrim' = 'whites are discriminated against', 'b_raceResent' = 'Racial resentment', 'b_pid' = 'Party ID', 'b_female' = 'Female', 'b_Intercept[1]' = 'Threshold 1', 'b_Intercept[2]' = 'Threshold 2', 'b_Intercept[3]' = 'Threshold 3', 'b_Intercept[4]' = 'Threshold 4')
```

```{r}
#| label: tbl-2012-lasso-simple
#| tbl-cap: Predictors of white Identity in 2012

rows <- tribble(~term, ~`2012 LWD`, 'Economic', '', 'Non-material', '', 'Demographics', '', 'Thresholds', '')
attr(rows, 'position') <- c(1, 10, 19, 28)

modelsummary(lasso[['2012 LWD']], coef_map = cm, statistic = 'conf.int', gof_map = gm, add_rows = rows, notes = list('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2012 American National Election Study.'))
```

```{r}
#| label: tbl-2016-lasso-simple
#| tbl-cap: Predictors of white identity in 2012

rows <- tribble(~term, ~`2016 LWD`, 'Economic', '', 'Non-material', '', 'Demographics', '', 'Thresholds', '')
attr(rows, 'position') <- c(1, 13, 22, 31)

modelsummary(lasso[['2016 LWD']], coef_map = cm, statistic = 'conf.int', gof_map = gm, add_rows = rows, notes = list('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2016 American National Election Study.'))
```

```{r}
#| label: tbl-2020-lasso-simple
#| tbl-cap: Predictors of white identity in 2020

rows <- tribble(~term, ~`2020 LWD`, 'Economic', '', 'Non-material', '', 'Demographics', '', 'Thresholds', '')
attr(rows, 'position') <- c(1, 13, 22, 31)

modelsummary(lasso[['2020 LWD']], coef_map = cm, statistic = 'conf.int', gof_map = gm, add_rows = rows, notes = list('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2020 American National Election Study.'))
```

```{r}
#| label: tbl-2012-2016-2020-ordinal-lasso
#| tbl-cap: Predictors of white identity in 2012, 2016, and 2020

model <- list(
    '2012 ANES' = lasso[['2012']],
    '2016 ANES' = lasso[['2016']],
    '2020 ANES' = lasso[['2020']]
)
rows <- tribble(~term, ~`2012 ANES`, ~`2016 ANES`, ~`2020 ANES`, 'Economic', '', '', '','Non-material', '', '', '', 'Demographics', '', '', '', 'Thresholds', '','', '')
attr(rows, 'position') <- c(1, 13, 22, 31)

modelsummary(model, coef_map = cm, statistic = 'conf.int', gof_map = gm, add_rows = rows, notes = list('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.' ,'Data Source: 2012, 2016, and 2020 American National Election Study.'))
```

### 2022-10-11 code/dcr_anes_2012_cleaning.R

@@ -14,7 +14,7 @@ box::use(
     #* create list object
 anes_2012 = list()
     #* Load original ANES file ----
-anes_2012[['original']] = read_dta('../data/anes-2012/anes-2012.dta')
+anes_2012[['original']] = read_dta('../../data/anes-2012/anes-2012.dta')
 # Cleaning ----
 anes_2012[['clean']] = anes_2012[['original']] |>
     #* Filter dataset to include only white respondents ----
@@ -56,6 +56,10 @@ anes_2012[['clean']] = anes_2012[['original']] |>
         #** Coded as: dem_edugroup_x - 1 < highschool, 2 hs, 3 some post-hs, 4 Bachelors, 5 graduate, < 1 Missing not asked, etc.
         #** Recoded to: edu - 1 < highschool, 2 hs, 3 some post-hs, 4 Bachelors, 5 graduate, NA = missing, not asked, etc
         edu = ifelse(dem_edugroup_x <= 0, NA, dem_edugroup_x),
+    #* income - Income
+        #** Coded as: incgroup_prepost_x - < 1 Missing, not asked, etc
+        #** Recoded to: NA < 1 Missing, not asked, etc
+        income = ifelse(incgroup_prepost_x >= 1, incgroup_prepost_x, NA),
     #* losejob - Worried about losing job
         #** Coded as: dem_losejob - 1 not at all - 5 extremely worried, < 1 missing, not asked, etc
         #** Recoded to: losejob - -2 = not at all - 2  = extremely worried, NA = missing, not asked, etc
@@ -159,4 +163,4 @@ anes_2012[['clean']] = anes_2012[['original']] |>
     )
 
 # Save data
-write.csv(anes_2012[['clean']], '../data/anes-2012/anes-2012-updated.csv')
+write.csv(anes_2012[['clean']], '../../data/anes-2012/anes-2012-updated.csv')

### 2022-10-11 code/dcr_anes_2016_cleaning.R

@@ -9,7 +9,7 @@
     #* create list object
 anes_2016 = list()
     #* load dataset
-anes_2016[['original']] = read_dta('../data/anes-2016/anes-2016.dta')
+anes_2016[['original']] = read_dta('../../data/anes-2016/anes-2016.dta')
 
 # Cleaning
 anes_2016[['clean']] = anes_2016[['original']] |>
@@ -52,6 +52,10 @@ anes_2016[['clean']] = anes_2016[['original']] |>
         #** Coded as: V161270 - 1 = < high school - 16 = Doctorate degree, < 1 missing, not asked, etc
         #** Recoded to: 1 = < high school - 16 = Doctorate degree, NA = missing, not asked, etc
     edu = ifelse(V161270 <= 0, NA, V161270),
+    #* income - Income
+        #** Coded as: V162309x - < 1 Missing, not asked, etc
+        #** Recoded to: NA = Missing, not asked, etc
+    income = ifelse(V162309x >= 1, V162309x, NA),
     #* losejob - Worried about losing job
         #** Coded as: V161297 - 1 not at all - 5 extremely worried, < 1 missing, not asked, etc
         #** Recoded to: losejob - -2 = not at all - 2  = extremely worried, NA = missing, not asked, etc
@@ -164,4 +168,4 @@ anes_2016[['clean']] = anes_2016[['original']] |>
     )
 
 # Save clean data
-write.csv(anes_2016[['clean']], '../data/anes-2016/anes-2016-updated.csv')
+write.csv(anes_2016[['clean']], '../../data/anes-2016/anes-2016-updated.csv')

### 2022-10-11 code/dcr_anes_2020_cleaning.r

@@ -9,7 +9,7 @@
     #* Create list object
 anes_2020 = list()
     #* Load dataset
-anes_2020[['original']] = read_dta('../data/anes-2020/anes_2020_original.dta')
+anes_2020[['original']] = read_dta('../../data/anes-2020/anes_2020_original.dta')
 
 # Clean the data
 anes_2020[['cleaned']] = anes_2020[['original']] |>
@@ -52,6 +52,10 @@ anes_2020[['cleaned']] = anes_2020[['original']] |>
         #** Coded as: V201511x - 1 < highschool, 2 hs, 3 some post-hs, 4 Bachelors, 5 graduate, < 1 Missing not asked, etc.
         #** Recoded to: edu - 1 < highschool, 2 hs, 3 some post-hs, 4 Bachelors, 5 graduate, NA = missing, not asked, etc
         edu = ifelse(V201511x <= 0, NA, V201511x),
+    #* income - Income
+        #** Coded as: V202468x - < 1 Missing, not asked, etc
+        #** Recoded to: NA = < 1 Missing, not asked, etc
+        income = ifelse(V202468x >= 1, V202468x, NA),
     #* losejob - Worried about losing job
         #** Coded as: V201540 - 1 not at all - 5 extremely worried, < 1 missing, not asked, etc
         #** Recoded to: losejob - -2 = not at all - 2  = extremely worried, NA = missing, not asked, etc
@@ -147,4 +151,4 @@ anes_2020[['cleaned']] = anes_2020[['original']] |>
     )
 
 # Save the clean data
-write.csv(anes_2020[['cleaned']], '../data/anes-2020/anes-2020-updated.csv')
+write.csv(anes_2020[['cleaned']], '../../data/anes-2020/anes-2020-updated.csv')

### 2022-10-12 drafts/fa_2022/dcr_white_identity_relative_deprivation_fa_2022.qmd
SUBMITTED!!!

@@
 abstract: |
-    Does economic or social and political threat explain white identity?
-Two leading arguments suggest that white identity results from a disruption of the racial status quo.
-However, they disagree on which disrupted element drives it.
-While one suggests that white identity comes from concern about whites' political influence, the other suggests that it is the result of concern with increased economic competition between whites and other racial groups.
-Considering the place that whiteness has in historical and contemporary institutions, I argue that white identity likely reflects concern about changes to their influence in the country's institutions.
-I use data from the American National Election Study that measure egotropic and sociotropic economic attitudes and leveraging questions gauging attitudes about social and political changes.
-Using a penalized regression, I examine whether proxies for economic threat appear to be plausible correlates of white identity.
-On balance, the models suggest that the hypothesis of economic threat is weaker than some suggest.
-I take this as evidence that white political identity, as it works today, is largely the result of whites' concerns with protecting the social and political status quo - and not so much economic.
+    Do economic or political threats explain reported white identity?
+Overall, the social identity literature would suggest that white identity would increase in response to economic threats.
+However, a number of those that study white identity, specifically, argue that it results from concern about political influence.
+Considering what whiteness means historically and contemporaneously, I argue that we should expect that political threats reflect stronger associations with white identity.
+Using data from the 2012, 2016, and 2020 American National Election Study, I consider a single model using penalized regression containing proxies of economic and political threats.
+I find evidence suggesting that in the post-Trump era, white political identity is strongly associated with reported feelings of Whites’ loss of political influence as opposed to economic threats, as some suggest and may expect.
 thanks: |

### 2022-12-28 drafts/fa_2022/dcr_white_identity_relative_deprivation_fa_2022.qmd
fix languages

@@
 data[['anes_2012']] <- read.csv('../../data/anes-2012/anes-2012-updated.csv') |>
-    select(wid, rboff, epast, unpast, edu, income, losejob, immtakejobs, ecfamily, ecjob, tradbreak, govbias, inflwhite, whitediscrim, raceResent, pid, female)
+    select(wid, rboff, epast, unpast, edu, income, losejob, ecfamily, ecjob, tradbreak, govbias, inflwhite, whitediscrim, raceResent, pid, female)
     #* 2016
 data[['anes_2016']] <- read.csv('../../data/anes-2016/anes-2016-updated.csv') |>
-    select(wid, rboff, epast, unpast, edu, income, losejob, immtakejobs, ecfamily, ecjob, tradbreak, govbias, inflwhite, whitediscrim, raceResent, pid, female)
+    select(wid, rboff, epast, unpast, edu, income, losejob, ecfamily, ecjob, tradbreak, govbias, inflwhite, whitediscrim, raceResent, pid, female)
     #* 2020
 data[['anes_2020']] <- read.csv('../../data/anes-2020/anes-2020-updated.csv') |>
-    select(wid, rboff, epast, unpast, edu, income, losejob, immtakejobs, ecfamily, ecjob, govbias, inflwhite, whitediscrim, raceResent, pid, female)
+    select(wid, rboff, epast, unpast, edu, income, losejob, ecfamily, ecjob, govbias, inflwhite, whitediscrim, raceResent, pid, female)
 ```
@@
 
-In many eras of American history, there are several instances where we have sought to reclassify who is white.
-During the antebellum period, whiteness was no longer simply one's economic class.
-However, now it had become a reflection of a racial hierarchy that afforded some various rights and excluded others from enjoying the same rights (e.g., literacy tests).
-During this era, rather than conceptualizing whites as landowners of Caucasian descent, the definition of what was white shifted to include those previously labeled non-white Europeans, such as the Irish.
-Changes to definitions of who is white illustrate that the primary motivator was race and not class as to who holds some rights and who does not [@rile-hayward_2013].
-
-More contemporaneously, whites hold several strong political attitudes racialized by politicians.
-Whites' attitudes toward welfare programs are not too distinct from those toward Black Americans [@gilens_1999].
-A provided explanation is that social welfare policies are more unpopular among whites when the media frames these policies as benefiting non-whites [@winter_2008;@van-doorn_2015].
-Explicit cues reminding individuals of the racial hierarchy entrench the recipients' attitudes toward Black Americans [@lyle_2014; see also @wellman_et-al_2016].
-Conservatives argued that Barack Obama, the first Black president, expressed support for the Affordable Care Act as a way to disproportionately help Black Americans [@tesler_2012;@tesler_2016].
-These messages corresponded with a rise of survey reports expressing old-fashioned racism among whites during his administration [@tesler_2013].
-While whites tend to report negative views about government aid for those in need when they are non-white recipients, evidence suggests that whites who live in communities with a higher proportion of poor whites report much more positive views of the poor than those that perceive the poor as mostly non-white [@hopkins_2009].
-Additionally, telling whites that other whites will be the beneficiaries of an entitlement program increases their support for such a program [@wetts_willer_2018].
-
 As an identity, much of the literature view these attitudes as coming from the presence or focus on economic threats.
@@
 Others argue that deindustrialization and job loss activate whites' concerns about the racial hierarchy and predict their support for Republican challengers that express support for keeping the racial status quo [@baccini_weymouth_2021_apsr].
-
-Not all agree, however.
-Some suggest that the negative attitudes that whites express toward immigrants result from concerns about cultural changes in those living in border towns [@hainmueller_hopkins_2014].
-Others suggest that simply reminding whites of these demographic shifts activates these racial identity-based attitudes [@craig_richeson_2014].
-Some other evidence suggests that these identity-based attitudes are better predicted by ethnic antagonism [@bartels_2020] and by racial animus [@fahey_et-al_2020] and explicitly take the position that a reaction to economic threats does not drive these.
-However, these attitudes tend to focus on prejudiced views rather than a reflection of a connection to other whites.
-
-# Is white identity non-material?
 
@@
 As this is an open question and is rather a complex one, I attempt to wade into this by parsing through which of these two perspectives are more plausible.
-As I discuss in the concluding section, I argue that scholars should continue taking the conceptualization and measurement of white political identification seriously and examine the mechanisms that connect economic and political threats to white political identity.
 In the next section, I examine whether attitudes expressing concern with economic or political threats are more robust correlates of the degree to which someone feels that being white is important to their identity.
@@
 data[['anes_2012']] |>
-  rename(`white identity` = wid, `Retrospective better off` = rboff, `Economy better` = epast,    `Unemployment better` = unpast, `Education` = edu, `Income` = income, `Worried about losing job` = losejob, `Immigrants take jobs` = immtakejobs, `Worried about family finances` = ecfamily, `Know someone lost job` = ecjob, `New lifestyles breaking society` = tradbreak, `Government biased against whites` = govbias, `whites influence politics` = inflwhite, `whites are discriminated against` = whitediscrim, `Racial resentment` = raceResent, `Party ID` = pid, `Female` = female) |>
+  rename(`white identity` = wid, `Retrospective better off` = rboff, `Economy better` = epast,    `Unemployment better` = unpast, `Education` = edu, `Income` = income, `Worried about losing job` = losejob, `Worried about family finances` = ecfamily, `Know someone lost job` = ecjob, `New lifestyles breaking society` = tradbreak, `Government biased against whites` = govbias, `whites influence politics` = inflwhite, `whites are discriminated against` = whitediscrim, `Racial resentment` = raceResent, `Party ID` = pid, `Female` = female) |>
   datasummary_skim(notes = c('NAs are included in Unique column.', 'Data source: 2012 American National Election Study, unweighted.'))
@@
 data[['anes_2016']] |>
-  rename(`white identity` = wid, `Retrospective better off` = rboff, `Economy better` = epast,    `Unemployment better` = unpast, `Education` = edu, `Income` = income, `Worried about losing job` = losejob, `Immigrants take jobs` = immtakejobs, `Worried about family finances` = ecfamily, `Know someone lost job` = ecjob, `New lifestyles breaking society` = tradbreak, `Government biased against whites` = govbias, `whites influence politics` = inflwhite, `whites are discriminated against` = whitediscrim, `Racial resentment` = raceResent, `Party ID` = pid, `Female` = female) |>
+  rename(`white identity` = wid, `Retrospective better off` = rboff, `Economy better` = epast,    `Unemployment better` = unpast, `Education` = edu, `Income` = income, `Worried about losing job` = losejob, `Worried about family finances` = ecfamily, `Know someone lost job` = ecjob, `New lifestyles breaking society` = tradbreak, `Government biased against whites` = govbias, `whites influence politics` = inflwhite, `whites are discriminated against` = whitediscrim, `Racial resentment` = raceResent, `Party ID` = pid, `Female` = female) |>
   datasummary_skim(notes = c('NAs are included in Unique column.', 'Data source: 2016 American National Election Study, unweighted.'))
@@
 data[['anes_2020']] |>
-  rename(`white identity` = wid, `Retrospective better off` = rboff, `Economy better` = epast,    `Unemployment better` = unpast, `Education` = edu, `Income` = income, `Worried about losing job` = losejob, `Immigrants take jobs` = immtakejobs, `Worried about family finances` = ecfamily, `Know someone lost job` = ecjob, `Government biased against whites` = govbias, `whites influence politics` = inflwhite, `whites are discriminated against` = whitediscrim, `Racial resentment` = raceResent, `Party ID` = pid, `Female` = female) |>
+  rename(`white identity` = wid, `Retrospective better off` = rboff, `Economy better` = epast,    `Unemployment better` = unpast, `Education` = edu, `Income` = income, `Worried about losing job` = losejob, `Worried about family finances` = ecfamily, `Know someone lost job` = ecjob, `Government biased against whites` = govbias, `whites influence politics` = inflwhite, `whites are discriminated against` = whitediscrim, `Racial resentment` = raceResent, `Party ID` = pid, `Female` = female) |>
   datasummary_skim(notes = c('NAs are included in Unique column.', 'Data source: 2020 American National Election Study, unweighted.'))
@@
 rows <- tribble(~term, ~`2012 LWD`, 'Economic', '', 'Non-material', '', 'Demographics', '', 'Thresholds', '')
-attr(rows, 'position') <- c(1, 10, 19, 28)
+attr(rows, 'position') <- c(1, 18, 27, 32)
 
@@
 #| label: tbl-2016-lasso-simple
-#| tbl-cap: Predictors of white identity in 2012
+#| tbl-cap: Predictors of white identity in 2016
 
 rows <- tribble(~term, ~`2016 LWD`, 'Economic', '', 'Non-material', '', 'Demographics', '', 'Thresholds', '')
-attr(rows, 'position') <- c(1, 13, 22, 31)
+attr(rows, 'position') <- c(1, 18, 27, 32)
 
@@
 rows <- tribble(~term, ~`2020 LWD`, 'Economic', '', 'Non-material', '', 'Demographics', '', 'Thresholds', '')
-attr(rows, 'position') <- c(1, 13, 22, 31)
+attr(rows, 'position') <- c(1, 18, 25, 30)
 
@@
 rows <- tribble(~term, ~`2012 ANES`, ~`2016 ANES`, ~`2020 ANES`, 'Economic', '', '', '','Non-material', '', '', '', 'Demographics', '', '', '', 'Thresholds', '','', '')
-attr(rows, 'position') <- c(1, 13, 22, 31)
+attr(rows, 'position') <- c(1, 18, 27, 32)
 

### 2022-12-28 code/dcr_anes_2016_cleaning.R

@@ -53,9 +53,9 @@ anes_2016[['clean']] = anes_2016[['original']] |>
         #** Recoded to: 1 = < high school - 16 = Doctorate degree, NA = missing, not asked, etc
     edu = ifelse(V161270 <= 0, NA, V161270),
     #* income - Income
-        #** Coded as: V162309x - < 1 Missing, not asked, etc
+        #** Coded as: V161361x - < 1 Missing, not asked, etc
         #** Recoded to: NA = Missing, not asked, etc
-    income = ifelse(V162309x >= 1, V162309x, NA),
+    income = ifelse(V161361x >= 1, V161361x, NA),
     #* losejob - Worried about losing job
         #** Coded as: V161297 - 1 not at all - 5 extremely worried, < 1 missing, not asked, etc
         #** Recoded to: losejob - -2 = not at all - 2  = extremely worried, NA = missing, not asked, etc

### 2022-12-29 code/dcr_anes_2020_cleaning.r
readjust file extension for 2020 cleaning

@@ -1,154 +1,0 @@
-# Title: 2020 ANES Cleaning for White Identity Project
-
-# Notes:
-    #* Description: R Script to do the cleaning for the 2020 ANES 
-    #* Updated: 2022-10-01
-    #* Updated by: dcr
-
-# Setup
-    #* Create list object
-anes_2020 = list()
-    #* Load dataset
-anes_2020[['original']] = read_dta('../../data/anes-2020/anes_2020_original.dta')
-
-# Clean the data
-anes_2020[['cleaned']] = anes_2020[['original']] |>
-    #* Filter only white respondents
-    filter(V201549x == 1) |>
-    mutate(
-    #* wid - White Identity
-        #** Coded as: V202499x - 1 = Extremely Important - 5 Not at all important, < 1 Missing, not asked, etc
-        #** Recoded to: wid - 1 = Not at all important - 5 = Extremely important, NA = Missing, not asked, etc
-        wid = ifelse(V202499x == 1, 5,
-                ifelse(V202499x == 2, 4,
-                    ifelse(V202499x == 3, 3,
-                        ifelse(V202499x == 4, 2,
-                            ifelse(V202499x == 5, 1, NA))))),
-    #* rboff - Retrospective better off: ----
-        #** Coded as: V201502 - 1 much better, 5 much worse, < 1 Missing, not asked, etc
-        #** Recoded to: rboff - -2 much worse, 2 much better, NA = missing, not asked, etc.
-        rboff = ifelse(V201502 == 1, 2,
-                    ifelse(V201502 == 2, 1,
-                        ifelse(V201502 == 3, 0,
-                            ifelse(V201502 == 4, -1,
-                                ifelse(V201502 == 5, -2, NA))))),
-    #* epast - Economy better/worse than 1 year ago: ----
-        #** Coded as: V201327x - 1 much better - 5 much worse, < 1 Missing, not asked, etc
-        #** Recoded to: epast - -2 much worse - 2 much better, NA = missing, not asked, etc
-        epast = ifelse(V201327x == 1, 2,
-                    ifelse(V201327x == 2, 1,
-                        ifelse(V201327x == 3, 0,
-                            ifelse(V201327x == 4, -1,
-                                ifelse(V201327x == 5, -2, NA))))),
-    #* unpast - Unemployment better/worse than 1 year ago
-        #** Coded as: V201333x - 1 much better - 5 much worse, < 1 Missing, not asked, etc
[…]

### 2022-12-29 code/dcr_anes_2020_cleaning.R
adjust file extension for 2020 cleaning

@@ -1,0 +1,154 @@
+# Title: 2020 ANES Cleaning for White Identity Project
+
+# Notes:
+    #* Description: R Script to do the cleaning for the 2020 ANES 
+    #* Updated: 2022-10-01
+    #* Updated by: dcr
+
+# Setup
+    #* Create list object
+anes_2020 = list()
+    #* Load dataset
+anes_2020[['original']] = read_dta('../../data/anes-2020/anes_2020_original.dta')
+
+# Clean the data
+anes_2020[['cleaned']] = anes_2020[['original']] |>
+    #* Filter only white respondents
+    filter(V201549x == 1) |>
+    mutate(
+    #* wid - White Identity
+        #** Coded as: V202499x - 1 = Extremely Important - 5 Not at all important, < 1 Missing, not asked, etc
+        #** Recoded to: wid - 1 = Not at all important - 5 = Extremely important, NA = Missing, not asked, etc
+        wid = ifelse(V202499x == 1, 5,
+                ifelse(V202499x == 2, 4,
+                    ifelse(V202499x == 3, 3,
+                        ifelse(V202499x == 4, 2,
+                            ifelse(V202499x == 5, 1, NA))))),
+    #* rboff - Retrospective better off: ----
+        #** Coded as: V201502 - 1 much better, 5 much worse, < 1 Missing, not asked, etc
+        #** Recoded to: rboff - -2 much worse, 2 much better, NA = missing, not asked, etc.
+        rboff = ifelse(V201502 == 1, 2,
+                    ifelse(V201502 == 2, 1,
+                        ifelse(V201502 == 3, 0,
+                            ifelse(V201502 == 4, -1,
+                                ifelse(V201502 == 5, -2, NA))))),
+    #* epast - Economy better/worse than 1 year ago: ----
+        #** Coded as: V201327x - 1 much better - 5 much worse, < 1 Missing, not asked, etc
+        #** Recoded to: epast - -2 much worse - 2 much better, NA = missing, not asked, etc
+        epast = ifelse(V201327x == 1, 2,
+                    ifelse(V201327x == 2, 1,
+                        ifelse(V201327x == 3, 0,
+                            ifelse(V201327x == 4, -1,
+                                ifelse(V201327x == 5, -2, NA))))),
+    #* unpast - Unemployment better/worse than 1 year ago
+        #** Coded as: V201333x - 1 much better - 5 much worse, < 1 Missing, not asked, etc
[…]

### 2023-06-26 src/R/plot.R
getting back at this project

@@ -1,0 +1,31 @@
+#' @title plotting LASSO results
+#' 
+#' @description
+#' This is a function that takes the results from the fitted LASSO models
+#' and makes a plot of the posterior distributions.
+#' 
+#' @details
+#' 
+#' @param
+#' 
+#' @returns
+#' 
+#' @example
+#' 
+lasso_plot <- function (
+  model
+  , level = 0.95
+  , year = "2012"
+) {
+  #* grab the posterior draws
+  df_posterior <- parameters(
+    model = model
+    , level = level
+    , verbose = FALSE
+  ) |>
+  standardize_names(style = "broom") |>
+  data.frame() |>
+  mutate(model = year) |>
+  filter(!is.na(term))
+
+}

### 2023-06-26 src/eda/dcr_anes_2012_cleaning.R

@@ -2,21 +2,25 @@
 
 # Notes: ----
     #* Description: Cleaning script for the 2012 ANES ----
-    #* Updated: 2022-10-01 ----
+    #* Updated: 2023-06-26 ----
     #* Updated by: dcr -----
 
 # Setup: -----
     #* Modularly load packages -----
 box::use(
     haven = haven[read_dta],
-    dplyr = dplyr[mutate, filter, case_when]
+    dplyr = dplyr[
+      case_when
+      , filter
+      , mutate
+    ]
 )
     #* create list object
-anes_2012 = list()
+anes_2012 <- list()
     #* Load original ANES file ----
-anes_2012[['original']] = read_dta('../../data/anes-2012/anes-2012.dta')
+anes_2012[["original"]] <- read_dta("../../../data/original/anes_2012/anes-2012.dta")
 # Cleaning ----
-anes_2012[['clean']] = anes_2012[['original']] |>
+anes_2012[["clean"]] <- anes_2012[["original"]] |>
     #* Filter dataset to include only white respondents ----
     filter(dem_raceeth_x == 1) |>
     mutate(
@@ -163,4 +167,4 @@ anes_2012[['clean']] = anes_2012[['original']] |>
     )
 
 # Save data
-write.csv(anes_2012[['clean']], '../../data/anes-2012/anes-2012-updated.csv')
+write.csv(anes_2012[["clean"]], '../../../data/clean/anes_2012_updated.csv')

### 2023-06-26 src/eda/dcr_anes_2016_cleaning.R

@@ -6,13 +6,22 @@
     #* Updated by: dcr 
 
 # Setup
+    #* Modularly load packages -----
+box::use(
+    haven = haven[read_dta],
+    dplyr = dplyr[
+      case_when
+      , filter
+      , mutate
+    ]
+)
     #* create list object
-anes_2016 = list()
+anes_2016 <- list()
     #* load dataset
-anes_2016[['original']] = read_dta('../../data/anes-2016/anes-2016.dta')
+anes_2016[["original"]] <- read_dta("../../../data/original/anes_2016/anes-2016.dta")
 
 # Cleaning
-anes_2016[['clean']] = anes_2016[['original']] |>
+anes_2016[["clean"]] <- anes_2016[["original"]] |>
     #* Filter only white respondents
     filter(V161310x == 1) |>
     mutate(
@@ -168,4 +177,4 @@ anes_2016[['clean']] = anes_2016[['original']] |>
     )
 
 # Save clean data
-write.csv(anes_2016[['clean']], '../../data/anes-2016/anes-2016-updated.csv')
+write.csv(anes_2016[["clean"]], "../../../data/clean/anes_2016_updated.csv")

### 2023-06-26 src/eda/dcr_anes_2020_cleaning.R

@@ -6,13 +6,22 @@
     #* Updated by: dcr
 
 # Setup
+    #* Modularly load packages -----
+box::use(
+    haven = haven[read_dta],
+    dplyr = dplyr[
+      case_when
+      , filter
+      , mutate
+    ]
+)
     #* Create list object
-anes_2020 = list()
+anes_2020 <- list()
     #* Load dataset
-anes_2020[['original']] = read_dta('../../data/anes-2020/anes_2020_original.dta')
+anes_2020[["original"]] <- read_dta("../../../data/original/anes_2020/anes_2020_original.dta")
 
 # Clean the data
-anes_2020[['cleaned']] = anes_2020[['original']] |>
+anes_2020[["clean"]] <- anes_2020[["original"]] |>
     #* Filter only white respondents
     filter(V201549x == 1) |>
     mutate(
@@ -151,4 +160,4 @@ anes_2020[['cleaned']] = anes_2020[['original']] |>
     )
 
 # Save the clean data
-write.csv(anes_2020[['cleaned']], '../../data/anes-2020/anes-2020-updated.csv')
+write.csv(anes_2020[["clean"]], "../../../data/clean/anes_2020_updated.csv")

### 2023-06-26 src/models/model_fitting.R

@@ -1,0 +1,186 @@
+# Title: model fitting
+
+# Notes
+  #* Description
+    #** fitting the models in the manuscript
+  #* Updated
+    #** 2023-06-26
+    #** dcr
+
+# Setup
+  #* execute cleaning scripts
+source("./src/eda/dcr_anes_2012_cleaning.R")
+source("./src/eda/dcr_anes_2016_cleaning.R")
+source("./src/eda/dcr_anes_2020_cleaning.R")
+  #* modularly load handy functions
+box::use(
+  , mice[
+    mice
+  ]
+  , brms[
+    brm
+    , brm_multiple
+    , bf
+    , set_prior
+    , lasso
+    , cumulative
+  ]
+)
+library(data.table)
+  #* define number of datasets to impute
+data_sets <- 10
+  #* create empty list objects
+data <- list()
+lasso <- list()
+  #* load cleaned data
+    #** 2012
+data[["anes_2012"]] <- read.csv("./data/clean/anes_2012_updated.csv")
+data[["anes_2012"]] <- data[["anes_2012"]][
+  , c(
+    "wid", "rboff", "epast", "unpast", "edu", "income", "losejob"
+    , "ecfamily", "ecjob", "tradbreak", "govbias", "inflwhite"
+    , "whitediscrim", "raceResent", "pid", "female"
+    )
+]
+    #** 2016
+data[["anes_2016"]] <- read.csv("./data/clean/anes_2016_updated.csv")
+data[["anes_2016"]] <- data[["anes_2016"]][
+  , c(
+    "wid", "rboff", "epast", "unpast", "edu", "income"
+    , "losejob", "ecfamily", "ecjob", "tradbreak"
+    , "govbias", "inflwhite", "whitediscrim", "raceResent", "pid", "female"
+  )
+]
+    #** 2020
+data[["anes_2020"]] <- read.csv("./data/clean/anes_2020_updated.csv")
+data[["anes_2020"]] <- data[["anes_2020"]][
+  , c(
+    "wid", "rboff", "epast", "unpast", "edu"
+    , "income", "losejob", "ecfamily", "ecjob"
+    , "govbias", "inflwhite", "whitediscrim"
+    , "raceResent", "pid", "female"
+  )
+]
+
+# Impute the data
+    #* 2012
+data[["anes_2012_imputed"]] <- mice(
+  data[["anes_2012"]]
+  , m = data_sets
+  , method = "rf"
+)
+    #* 2016
[…]

### 2023-06-27 src/manuscript.qmd
draft done

--- 2022-12-28 drafts/fa_2022/dcr_white_identity_relative_deprivation_fa_2022.qmd
@@
 thanks: |
+    Word count: 3575.
 I would like to thank Jennifer Wolak for her feedback on the many drafts of this manuscript.
@@
 date: today
-bibliography: "../../assets/wid.bib"
+bibliography: "../assets/wid.bib"
 format:
   hikmah-pdf:
-    biblatex-chicago: true
-    biblio-style: authordate
-    biblatexoptions:
-      - backend=biber
-      - autolang=hyphen
-      - isbn=false
-      - uniquename=false
+    cite-method: citeproc
+    latex-output-dir: ../out/submissions/bjps
+    filters:
+    - "_extensions/andrewheiss/wordcount/citeproc.lua"
+    - "_extensions/andrewheiss/wordcount/wordcount.lua"
+    geometry:
+      - top=1in
+      - bottom=1in
+      - left=1in
+      - right=1in
+      - heightrounded
+    linestretch: 2
+    fontsize: 12pt
+    appendix-style: none
+    include-in-header:
+      text: |
+        \usepackage{pdflscape}
+        \newcommand{\blandscape}{\begin{landscape}}
+        \newcommand{\elandscape}{\end{landscape}}
 linestretch: 2
@@
   message: false
-header-includes:  
-- \usepackage{lscape}
+params:
+  replicate: false
 ---
@@
 #| label: block-setup
-
 # set seed
-set.seed(90210)
+set.seed(12102022)
 # modularly load functions
 box::use(
-    dplyr = dplyr[select, filter, mutate, case_when, rename],
-    tidyr = tidyr[drop_na],
-    mice = mice[mice, pool],
-    miceadds = miceadds[mids2datlist],
-    brms = brms[brm, brm_multiple, set_prior, lasso, cumulative, prior],
-    tibble = tibble[tribble],
-    modelsummary = modelsummary[datasummary_skim, datasummary_correlation, modelsummary],
-    parameters = parameters[parameters, standardize_names],
-    ggplot2 = ggplot2[ggplot, aes, position_dodge, theme_minimal, geom_vline, labs, theme, element_text, scale_y_discrete],
-    tidybayes = tidybayes[geom_pointinterval],
-    performance = performance[check_model]
+  dplyr = dplyr[
+    mutate
+    , filter
+    , case_when
+    , select
+  ]
+  , tibble = tibble[
+    tribble
+  ]
+  , modelsummary = modelsummary[
+    datasummary_skim
+    , modelsummary
+  ]
+  , parameters = parameters[
+    parameters
+    , standardize_names
+  ]
+  , ggplot2 = ggplot2[
+    ggplot
+    , aes
+    , position_dodge
+    , theme_minimal
+    , geom_vline
+    , labs
+    , theme
+    , element_text
+    , scale_y_discrete
+  ]
+  , tidybayes = tidybayes[
+    geom_pointinterval
+  ]
+  , data.table[
+    setnames
+  ]
 )
+# If doing a full replication, change replicate option in yaml section
+if (params$replicate == TRUE) {
+  source("./models/model_fitting.R")
+}
 ```
@@
 ```{r}
-#| label: cleaning-block
-
-# source cleaning scripts
-source('../../code/dcr_anes_2012_cleaning.R')
-source('../../code/dcr_anes_2016_cleaning.R')
-source('../../code/dcr_anes_2020_cleaning.R')
-
-```
-
-```{r}
-#| label: load-data-block
-# create list object for datasets
-data = list()
-# Create empty lasso list to store fitted models
-lasso = list()
-# Specify number of iterations for imputation
-data_sets = 10
-# load the cleaned data
-    #* 2012
-data[['anes_2012']] <- read.csv('../../data/anes-2012/anes-2012-updated.csv') |>
-    select(wid, rboff, epast, unpast, edu, income, losejob, ecfamily, ecjob, tradbreak, govbias, inflwhite, whitediscrim, raceResent, pid, female)
-    #* 2016
-data[['anes_2016']] <- read.csv('../../data/anes-2016/anes-2016-updated.csv') |>
-    select(wid, rboff, epast, unpast, edu, income, losejob, ecfamily, ecjob, tradbreak, govbias, inflwhite, whitediscrim, raceResent, pid, female)
-    #* 2020
-data[['anes_2020']] <- read.csv('../../data/anes-2020/anes-2020-updated.csv') |>
-    select(wid, rboff, epast, unpast, edu, income, losejob, ecfamily, ecjob, govbias, inflwhite, whitediscrim, raceResent, pid, female)
+#| label: load-fitted-model-block
+load("../data/temp/fitted_models.RData")
 ```
@@
 Analyses by political scientists largely attribute Trump's success in attracting voters to those who see their whiteness as a politically relevant identity where they are distinct from Americans of other racial groups [@jardina_2020_pp].
-For a long time, political scientists saw white political behavior and attitudes through the lens of prejudice.
-These analyses exclude another dimension of whites' racial attitudes: their affinity for, feelings of linked fate with other whites, and an expression that their race is a politically-relevant identity.
-More recently, scholars using the @tajfel_turner_1986 social psychological framework of identity challenge this tendency to focus on whites' prejudice by arguing that some whites may see their race as an identity essential and relevant to politics.
-As a result, while some may express out-group attitudes, they also express an in-group preference for other whites [@jardina_2019_cup].
-Despite the importance of these attitudes, the source of these in-group preferences remains somewhat unclear.
+What remains an open question is whether it actually is whites' concerns about economic status that motivates whites to use their race as a lens to evaluate politics or whether it is a backlash to changes in the racial groups reflected by those occupying visible political offices.
 
-There appears to be a mismatch between those who study sources of a variety of social identities and those that study white political identity.
+While this is an ongoing debate, part of this disagreement between scholars about where white identity comes from may be an artifact of the traditions of the theoretical frameworks we use in this area.
 Scholars studying identity, more broadly, often focus on economic threats as a critical motivator for an expressed connection to a social in-group [see @huddy_2001].
@@
 This mismatch leads to some conceptualizing white identity as originating from economic views in the tradition of the identity literature.
-In contrast, others conceptualize white identity as coming from political and cultural threats.
-
-I agree with those who focus on the connection between white political identity and political threats.
-Considering identity's historical and contextual roots [@trawalter_et-al_2022_cdps], I argue that white identifiers are concerned about their status in politics and society.
-Though I acknowledge that economic threat is a potent activation of identity and encourages engagement in group-based behavior, white identity holds decisive cultural and institutional significance in the U.S.
-[@rile-hayward_2013].
-One's whiteness is not a descriptive characteristic, but it reflects a history of allowances by institutions and advantages in one's ability to influence politics.
-Though race and economics are undoubtedly correlated, we must be careful in specifying the data-generating process.
+In contrast, others conceptualize white identity as motivated by political threats.
 
@@
 With the introduction of competition, these boys reported fierce hostility toward the members of the other group and expressed positive attitudes toward those within the group.
-
+<!--
 This example illustrates several features of social identities.
@@
 Whiteness and one's willingness to identify with that group is important [@wong_cho_2005] as it reflects a desire for an association with the history and affordances of being white [@mcdermott_samson_2005].
-
-As an identity, much of the literature view these attitudes as coming from the presence or focus on economic threats.
-Some studying white political identity find economic predictors as substantively and statistically significant [see @perez_et-al_2022_pp].
-From non-experimental evidence, others make similar arguments about what predicts white political identity.
-@bunyasi_2019 provides evidence suggesting that Trump's messages highlighting increased job market competition harming whites because of affirmative action was a successful strategy and helped build his base of support.
+-->
+Recent work makes the case that we should seriously consider whites' racial identity as an identity [@jardina_2019_cup].
+As an identity, much of the literature view these attitudes resulting from whites' racial identity as stemming from economic threat.
+Some studying white political identity find economic predictors as substantively and statistically significant [see @perez_et-al_2022_pp;@lindsay_2023_apr on their racial attitudes writ large].
+Observational evidence supports this claim as well.
+For example, Trump's messaging about job market competition with immigrants appeared to be successful at attracting white voters [@bunyasi_2019].
 After all, the identity literature, more broadly, often focuses on economic threats.
@@
 
-Historical accounts that contextualize attitudes stemming from white racial identity seem to suggest that, though economics correlates with racial attitudes, what it meant to be white was no longer about class.
-However, it became about what it meant for access and privileges afforded by institutions [@rile-hayward_2013].
-In the $19^{th}$ century, privilege came from one's freedom and wealth.
-Those who were not enslaved but were poor were not considered white and were of a lower strata than whites who owned land [@rile-hayward_2013].
-With the abolition of slavery, the expressed (explicit) concern became not about wealth but now about race.
-This motivated a shift in how the institutions and how whites defined themselves.
-It became the focus for whites to be a higher status group, regardless of whether or not they were poorer than formerly enslaved Black Americans [@rile-hayward_2013].
+# Why might white identity be a backlash to loss of institutional privelege?
 
-Many conceptualizations of white political identity align with its primary source as concern for the group's political power.
 While a critical component of identity is a feeling of your fate as linked with others in your group [@leach_et-al_2008], this linked fate is the connection between attitudes and material forms of threat but that identity, in and of itself, is better conceptualized as a result of cultural or ideational threat [@donnelly_2020].
@@
 
-As this is an open question and is rather a complex one, I attempt to wade into this by parsing through which of these two perspectives are more plausible.
+Whites' institutional privelege and threats to it are the more potent threats that whites may be less used to and as a result, coalesce as a group to defend.
+@winter_2008 demonstrates that whites' negative attitudes toward social welfare programs are motivated by perceptions that beneficiaries are Black.
+In this book length examination of the question, it is not that the people are poor and could use government assistance, it is that those receiving help are not white.
+Further, the increasing racial polarization is not something that is occuring among whites who are most economically threatened, but the Republican party is whiter with both wealthy and less wealthy whites -- political considerations are causally prior to racial attitudes [@engelhardt_2021_bjps; see also @schaffner_2022_bjps].
+
+An example of a contemporary institutional threat to whites is the election of Barack Obama.
+Evidence suggests that America did not shift into a post-racial era as some political commentators suspected.
+In fact, race became an extremely salient issue for whites [@tesler_2016].
+White Republicans exhibited higher levels of old fashioned racism [@tesler_2013].
+Issues were racialized during Obama's presidency and many Whites exhibited lower levels of support for those policies in response [@tesler_2012;@tesler_2016].
+As a result of whites' intense concern with race, non-white political candidates, such as Obama, often had to distance themselves from their racial groups in speeches before white audiences as a way to not activate concerns about being to beholden to their racial group [@stephens-dougan_2020].
+The reason for such potent responses to indicators of change for what groups have influence in political instittuions is simply that whites are not used to it.
+They have spent the entirety of this country's existence defending the idea that whites have higher status in American society regardless of wealth, education, or other forms of status.
+
+My argument is not a new one.
+Historical accounts outlining the complicated history of what it meant to be "white" from a legal standpoint boils down to: did other whites see you as being good enough to be called white?
+There are many examples of groups that are now considered white but have not always been so.
+For example groups from eastern europe were often considered as non-white.
+While these definitions often correlated with economic status (e.g., German immigrants were often considered white but not Irish immigrants), they often changed in response to concerns about weakening white political influence with other groups growing in political influence such as the Irish in Boston.
+During these shifts in definitions of whiteness were no longer about class.
+Instead, it became about what it meant for access and privileges afforded by institutions [@rile-hayward_2013].
+In the $19^{th}$ century, privilege came from one's freedom and wealth.
+Those who were not enslaved but were poor were not considered white and were of a lower strata than whites who owned land.
+With the abolition of slavery, the expressed (explicit) concern became not about wealth but now about race.
+Whites could no longer use the classification of "free" and "land-owning" as an excuse to exclude Blacks from full access to America's institutions.
+This motivated a shift in how the institutions and how whites defined themselves [@rile-hayward_2013].
+It became the focus for whites to be a higher status group, regardless of whether or not they were poorer than formerly enslaved Black Americans [@rile-hayward_2013].
+In other words, the white American tradition is to hold one's whiteness as a status symbol regardless of other forms of status such as class.
+
+Not one single paper or effort to provide empirical evidence can satisfactorily address such a large and complex question.
+The causal mechanism here is extremely difficult to pin down.
+In an effort to address this question in a bite-sized chunk, I first start with examining whether or not correlational evidence aligns with my claim that individual-level economic predictors are weaker than perceptions of political influence.
 In the next section, I examine whether attitudes expressing concern with economic or political threats are more robust correlates of the degree to which someone feels that being white is important to their identity.
@@
 The studies are nationally representative questions that capture several questions that lend themselves to testing the relationship between economic and political concerns while also asking white respondents to report how important being white is to their identity.
-
+<!--
 Of course, I cannot *prove* that these questions reflect the two types of threats I purport to represent.
@@
 Nevertheless, the paper aims to offer an argument that contextualizes white identity in history and how that leads to different predictions about the correlates of white identity than many might assume from understanding the literature on the origins of social identity.
-
-Each study asks white respondents, "How important is being white to your Identity" and offers them a five-point Likert scale with the options reflecting their level of agreement with the question ^[Details about question-wording, original ANES coding, and my coding for all variables I include in my models are included in @tbl-anes-measures located in the Appendix.
-I include descriptive statistics for the measures in these samples in @tbl-2012-descriptive-stats, @tbl-2016-descriptive-stats, and @tbl-2020-descriptive-stats in the Appendix].
-They do not ask white respondents to report on items from the white political identity measures of @jardina_2019_cup or the more complex battery of items that @helms_carter_1990 discusses.
-However, this question does capture the degree to which one sees their whiteness as part of who they are and reflects the conceptualization that others have used [see @schildkraut_2017;@jardina_2019_cup;@perez_et-al_2022_pp].
+-->
+Each study asks white respondents, "How important is being white to your Identity" and offers them a five-point Likert scale with the options reflecting their level of agreement with the question.^[Details about question-wording, original ANES coding, and my coding for all variables I include in my models are included in @tbl-anes-measures located in the Supplementary Information.
+I include descriptive statistics for the measures in these samples in @tbl-2012-descriptive-stats, @tbl-2016-descriptive-stats, and @tbl-2020-descriptive-stats located in the Supplementary Information.] They do not ask white respondents to report on items from the white political identity measures of @jardina_2019_cup or the more complex battery of items that @helms_carter_1990 discusses.
+However, this question does capture the degree to which one sees their whiteness as part of who they are and reflects the operationalization that others have used [see @schildkraut_2017;@jardina_2019_cup;@perez_et-al_2022_pp].
 
@@
 I also want to examine the degree to which respondents feel that the institutions benefit non-whites rather than whites by including the question that whites face discrimination and that the government is biased against whites.
-I include a question asking respondents about their belief that traditions in this country are breaking^[The ANES only included this question in the 2012 and 2016 questionnaires].
-While these, like the economic attitudes, are just proxies of threat, they reflect the extent to which respondents may feel insecure about the political influence of whites.
+I include a question asking respondents about their belief that traditions in this country are breaking.^[The ANES only included this question in the 2012 and 2016 questionnaires.] While these, like the economic attitudes, are just proxies of threat, they reflect the extent to which respondents may feel insecure about the political influence of whites.
 
@@
 
-One option for implementation is to use the "classical" LASSO, which relies on an optimization procedure through cross-validation.
-It is a frequentist approach to LASSO penalized regression.
-However, there are several positive benefits to the second implementation of the LASSO penalized regression in a Bayesian framework [see @van-erp_et-al_2019_jmp].
-From my review of the literature, I have \textit{apriori} expectations for the distribution of effects on my predictors in the form of a Laplace distribution which I explicitly define in my prior density function.
-I also have the additional benefit of having the ability to define its hyperparameters.
-Meaning I suspect that there is a peak around zero and that there are relatively large tails.
-This characterization of my prior fits with that of the LASSO - a large number of included predictors have zero effect, with a relatively smaller proportion of included predictors that do have an effect but are relatively unsure as to how large those coefficients might be.
-I can specify the scale of that uncertainty by specifying a hyperparameter for this LASSO prior.
-Without having to perform cross-validation as I would with the classical LASSO, my posterior distribution naturally reflects the uncertainty around the size of my L2 shrinkage parameter and can estimate models with different L2 shrinkage parameters simultaneously.
-As $\verb|stan|$ includes implementations of this model, I rely on Hamiltonian Monte Carlo, a type of Markov Chain Monte Carlo sampling which allows for non-convexity in the L2 shrinkage parameters.
-As classical LASSO models are challenging to interpret due to cross-validation, using credible intervals offers a much less error-prone interpretation.
+Rather than use the classical LASSO with Leave One Out Cross Validation (LOOCV), I take advantage of a type of Markov Chain Monte Carlos (MCMC) called a Hamiltonian Monte Carlo for my LASSO due to a number of benefits outlined by @van-erp_et-al_2019_jmp.
+As the literature leads me to hold *apriori* expectations about the distribution of predictors that meaningfully effect white political identity, I specify a Laplace distribution as my prior density function.
+I suspect that there are a large number of included predictors that have null effects and a relatively smaller proportion of predictors that *do* have an effect but am relatively unsure about how large their coefficients might be.
+This characterization of my prior fits with that of an L2 Shrinkage parameter specified in a LASSO regression.
+As I am not using LOOCV, I am able to interpret the results of the model through my credible intervals.
 
 My measure of white identity is ordinal.
-Though white identity as a concept is likely continuous, through the specification of a cumulative density function with a logit link, my model bins this latent continuous space based on the number of response options.
+Therefore, I apply a ordered logistic link function.
 Though it creates a more complicated computational task, it is a more appropriate model specification given the available measure and my assumptions about the concept [@liddell_kruschke_2018_jesp].
 
-To account for potential problems stemming from missing data, I impute my data by using multiple imputation with chained equations (MICE) where the particular model I use is a random forest model to allow for more flexibility and to reduce model dependence [@buuren_goothuis-oudshoorn_2011; @marbach_2022_pa]^[I impute ```r sprintf('%.0f', data_sets)``` datasets and simultaneously fit my Bayesian LASSO on each dataset and pool the results.
+To account for potential problems stemming from missing data, I impute my data by using multiple imputation with chained equations (MICE) where the particular model I use is a random forest model to allow for more flexibility and to reduce model dependence [@buuren_goothuis-oudshoorn_2011; @marbach_2022_pa].^[I impute 10 datasets and simultaneously fit my Bayesian LASSO on each dataset and pool the results.
 Uncertainty reflected in my reported credible intervals not only reflects the uncertainty from one model, but my reported uncertainty also reflects the uncertainty generated from the imputation procedure.
-One drawback with this particular approach is that my normalized split-$\hat{R}$ will likely be greater than the widely recognized 1.01 cutoff that indicates model convergence due to the pooling of my multiply imputed datasets.].
-I report the results of the model fitted using listwise deletion in @tbl-2012-lasso-simple, @tbl-2016-lasso-simple, and @tbl-2020-lasso-simple.
-
-```{r}
-#| label: imputation-block
-#| eval: false
-# Imputation
-    #* 2012
-data[['anes_2012_imputed']] <- mice(data[['anes_2012']], m = data_sets, method = 'rf')
-saveRDS(data[['anes_2012_imputed']], file = '../data/anes-2012/anes-2012-imputed.rds')
-    #* 2016
-data[['anes_2016_imputed']] <- mice(data[['anes_2016']], m = data_sets, method = 'rf')
-saveRDS(data[['anes_2016_imputed']], file = '../data/anes-2016/anes-2016-imputed.rds')
-    #* 2020
-data[['anes_2020_imputed']] <- mice(data[['anes_2020']], m = data_sets, method = 'rf')
-saveRDS(data[['anes_2020_imputed']], file = '../data/anes-2020/anes-2020-imputed.rds')
-```
-
-```{r}
-#| label: imputation-loading-block
-
-data[['anes_2012_imputed']] <- readRDS('../..//data/anes-2012/anes-2012-imputed.rds')
-
-data[['anes_2016_imputed']] <- readRDS('../../data/anes-2016/anes-2016-imputed.rds')
-
-data[['anes_2020_imputed']] <- readRDS('../../data/anes-2020/anes-2020-imputed.rds')
-```
+One drawback with this particular approach is that my normalized split-$\hat{R}$ will likely be greater than the widely recognized 1.01 cutoff that indicates model convergence due to the pooling of my multiply imputed datasets.] I report the results of the model fitted using listwise deletion in the Supplementary Information.
 
@@
 
-@fig-2012-2016-2020-ordinal-lasso presents the estimates of my fitted ordinal models with the LASSO shrinkage prior^[@tbl-2012-2016-2020-ordinal-lasso presents the table of results.].
-These models are fit on 6 chains with 2000 iterations each.
+@fig-2012-2016-2020-ordinal-lasso presents the estimates of my fitted ordinal models with the LASSO shrinkage prior.^[The full tables of results are included in the Supplementary Information.] These models are fit on 6 chains with 2000 iterations each.
 To examine the evidence of whether a particular predictor "matters" for predicting white political identity, I examine the estimated credible intervals, which report the probability that the true estimate falls within the estimated range.
@@
 ```{r}
-#| label: models-block
-#| eval: false
+#| label: fig-2012-2016-2020-ordinal-lasso
+#| layout-ncol: 2
+#| fig-cap: Economic factors are weak predictors of white identity
+#| fig-subcap:
+#|   - Economic
+#|   - Political
+# Graphs
+  #* Tidyfy posterior draws
+    #** 2012
+lasso[['2012_ordinal_tidy']] <- parameters(
+  lasso[['2012_ordinal']]
+  , conf.int = 0.95
+  , verbose = FALSE
+) |> # grab the posterior draws and construct CI's at the 95% level
+  standardize_names(style = 'broom') |> # standardize the coefficient names
+  data.frame() |> # convert it to a data.frame
+  mutate(model = '2012') |> # add a column called model with the value of 2012
+  filter(!is.na(term)) # get rid of any rows that are na
+    #** 2016
+lasso[['2016_ordinal_tidy']] <- parameters(
+  lasso[['2016_ordinal']]
+  , conf.int = 0.95
+  , verbose = FALSE
+) |> # grab the posterior draws and construct CI's at the 95% level
+  standardize_names(style = 'broom') |> # standardize coefficient names
+  data.frame() |> # convert it to a data.frame
+  mutate(model = '2016') |> # add a column called model with the value of 2016
+  filter(!is.na(term)) # get rid of any rows that are na
+    #** 2020
+lasso[['2020_ordinal_tidy']] <- parameters(
+  lasso[['2020_ordinal']]
+  , conf.int = 0.95
+  , verbose = FALSE
+) |> # grab the posterior drawsw and construct CI's at the 95% level
+  standardize_names(style = 'broom') |> # standardize coefficient names
+  data.frame() |> # convert it to a data.frame
+  mutate(model = '2020') |> # add a column called model with the value of 2020
+  filter(!is.na(term)) # get rid of any rows that are na
 
-# Models
-    #* 2012
-lasso[['2012_simple']] <- brm(wid ~ rboff + epast + unpast + edu + income + losejob + ecfamily + ecjob + tradbreak + govbias + whitediscrim + raceResent + pid + female, data = data[['anes_2012']], prior = set_prior(lasso(1)), chains = 6, iter = 2000, family = cumulative(link = 'logit'))
-saveRDS(lasso[['2012_simple']], file = '../data/lasso_2012_simple.rds')
-
-lasso[['2012_ordinal']] <- brm_multiple(wid ~ rboff + epast + unpast + edu + income + losejob +  ecfamily + ecjob + tradbreak + govbias + whitediscrim + raceResent + pid + female, data = data[['anes_2012_imputed']], prior = set_prior(lasso(1)), chains = 6, iter = 2000, family = cumulative(link = 'logit'))
-saveRDS(lasso[['2012_ordinal']], file = '../data/lasso_2012_ordinal.rds')
-    #* 2016
-lasso[['2016_simple']] <- brm(wid ~ rboff + epast + unpast + edu + income + losejob + ecfamily + ecjob + tradbreak + govbias + whitediscrim + raceResent + pid + female, data = data[['anes_2016']], prior = set_prior(lasso(1)), chains = 6, iter = 2000, family = cumulative(link = 'logit'))
-saveRDS(lasso[['2016_simple']], file = '../data/lasso_2016_simple.rds')
-
-lasso[['2016_ordinal']] <- brm_multiple(wid ~ rboff + epast + unpast + edu + income + losejob + ecfamily + ecjob + tradbreak + govbias + whitediscrim + raceResent + pid + female, data = data[['anes_2016_imputed']], prior = set_prior(lasso(1)), chains = 6, iter = 2000, family = cumulative(link = 'logit'))
-saveRDS(lasso[['2016_ordinal']], file = '../data/lasso_2016_ordinal.rds')
-    #* 2020
-lasso[['2020_simple']] <- brm(wid ~ rboff + epast + unpast + edu + income + losejob + ecfamily + ecjob + govbias + whitediscrim + raceResent + pid + female, data = data[['anes_2020']], prior = set_prior(lasso(1)), chains = 6, iter = 2000, family = cumulative(link = 'logit'))
-saveRDS(lasso[['2020_simple']], file = '../data/lasso_2020_simple.rds')
-
-lasso[['2020_ordinal']] <- brm_multiple(wid ~ rboff + epast + unpast + edu + income + losejob + ecfamily + ecjob + govbias + whitediscrim + raceResent + pid + female, data = data[['anes_2020_imputed']], prior = set_prior(lasso(1)), chains = 6, iter = 2000, family = cumulative(link = 'logit'))
-saveRDS(lasso[['2020_ordinal']], file = '../data/lasso_2020_ordinal.rds')
-```
-
-```{r}
-#| label: loading-fitted-models-block
-
-# 2012
-lasso[['2012 LWD']] <- readRDS('../../data/lasso_2012_simple.rds') 
-lasso[['2012']] <- readRDS('../../data/lasso_2012_ordinal.rds') 
-# 2016
-lasso[['2016 LWD']] <- readRDS('../../data/lasso_2016_simple.rds') 
-lasso[['2016']] <- readRDS('../../data/lasso_2016_ordinal.rds')
-# 2020
-lasso[['2020 LWD']] <- readRDS('../../data/lasso_2020_simple.rds')
-lasso[['2020']] <- readRDS('../../data/lasso_2020_ordinal.rds')
-```
-
-```{r}
-#| label: fig-2012-2016-2020-ordinal-lasso
-#| fig-cap: Economic factors are weak predictors of white identity
-#| fig-height: 9
-#| fig-width: 9
-# Graphs
-    #* Posterior draws
-lasso[['2012_ordinal_tidy']] <- parameters(lasso[['2012']], conf.int = 0.95, verbose = FALSE) |>
-    standardize_names(style = 'broom') |>
-    data.frame() |>
-    mutate(model = '2012') |>
-    filter(!is.na(term))
-
-lasso[['2016_ordinal_tidy']] <- parameters(lasso[['2016']], conf.int = 0.95, verbose = FALSE) |>
-    standardize_names(style = 'broom') |>
-    data.frame() |>
-    mutate(model = '2016') |>
-    filter(!is.na(term))
-
-lasso[['2020_ordinal_tidy']] <- parameters(lasso[['2020']], conf.int = 0.95, verbose = FALSE) |>
-    standardize_names(style = 'broom') |>
-    data.frame() |>
-    mutate(model = '2020') |>
-    filter(!is.na(term))
-
-lasso[['ordinal_tidy']] <- rbind(lasso[['2012_ordinal_tidy']], lasso[['2016_ordinal_tidy']], lasso[['2020_ordinal_tidy']]) |>
+lasso[['ordinal_tidy']] <- rbind(
+  lasso[['2012_ordinal_tidy']]
+  , lasso[['2016_ordinal_tidy']]
+  , lasso[['2020_ordinal_tidy']]
+) |> # append the three models' results together
   mutate(
-    term = case_when(term == 'b_rboff' ~ 'Retrospective better off', term == 'b_epast' ~ 'Economy better', term == 'b_unpast' ~ 'Unemployment better', term == 'b_edu' ~ 'Education', term == 'b_income' ~ 'Income', term == 'b_losejob' ~ 'Worried about losing job', term == 'b_ecfamily' ~ 'Worried about family finances', term == 'b_ecjob' ~ 'Know someone lost job', term == 'b_tradbreak' ~ 'New lifestyles breaking society', term == 'b_govbias' ~ 'Gov.
-biased against whites', term == 'b_inflwhite' ~ 'whites influence politics', term == 'b_whitediscrim' ~ 'whites are discriminated against', term == 'b_raceResent' ~ 'Racial resentment', term == 'b_pid' ~ 'Party ID', term == 'b_female' ~ 'Female'),
-    term = factor(term, levels = c('Retrospective better off', 'Economy better', 'Unemployment better', 'Education', 'Income', 'Worried about losing job', 'Worried about family finances', 'Know someone lost job', 'New lifestyles breaking society', 'Gov.
-biased against whites', 'whites influence politics', 'whites are discriminated against', 'Racial resentment', 'Party ID', 'Female'))
+    term = case_when(
+      term == 'b_rboff' ~ 'Retrospective better off'
+      , term == 'b_epast' ~ 'Economy better'
+      , term == 'b_unpast' ~ 'Unemployment better'
+      , term == 'b_edu' ~ 'Education'
+      , term == 'b_income' ~ 'Income'
+      , term == 'b_losejob' ~ 'Worried about losing job'
+      , term == 'b_ecfamily' ~ 'Worried about family finances'
+      , term == 'b_ecjob' ~ 'Know someone lost job'
+      , term == 'b_tradbreak' ~ 'New lifestyles breaking society'
+      , term == 'b_govbias' ~ 'Gov. biased against whites'
+      , term == 'b_inflwhite' ~ 'whites influence politics'
+      , term == 'b_whitediscrim' ~ 'whites are discriminated against'
+      , term == 'b_raceResent' ~ 'Racial resentment'
+      , term == 'b_pid' ~ 'Party ID'
+      , term == 'b_female' ~ 'Female'
+    ), # rename the parameters
+    term = factor(
+      term
+      , levels = c(
+        'Retrospective better off'
+        , 'Economy better'
+        , 'Unemployment better'
+        , 'Education'
+        , 'Income'
+        , 'Worried about losing job'
+        , 'Worried about family finances'
+        , 'Know someone lost job'
+        , 'New lifestyles breaking society'
+        , 'Gov. biased against whites'
+        , 'whites influence politics'
+        , 'whites are discriminated against'
+        , 'Racial resentment'
+        , 'Party ID'
+        , 'Female'
+      )
+    ) # convert it to a factor variable
   ) |> 
-  filter(!is.na(term))
-
-ggplot(data = lasso[['ordinal_tidy']], aes(y = term, x = estimate, xmin = conf.low, xmax = conf.high, linetype = factor(model), shape = factor(model))) +
-    geom_pointinterval(position = position_dodge(width = 0.4), size = 3) +
-    geom_vline(xintercept = 0, linetype = 2) +
+  filter(!is.na(term)) # get rid of na rows
+    #** make a data.frame of results for the econ predictors
+lasso[["ordinal_tidy_econ"]] <- lasso[["ordinal_tidy"]] |>
+  filter(
+    term == "Retrospective better off" | term == "Economy better" | 
+    term == "Unemployment better" | term == "Education" | term == "Income" | 
+    term == "Worried about losing job" | term == "Wrroeid about family finances" |
+    term == "Know someone lost job"
+  )
+    #** make a data.frame of results for the political predictors
+lasso[["ordinal_tidy_poli"]] <- lasso[["ordinal_tidy"]] |>
+  filter(
+    term == "New lifestyles breaking society" | term == "Gov. biased against whites" |
+    term == "whites influence politics" | term == "whites are discriminated against" |
+    term == "Racial resentment"
+  )
+  #* plot for econ predictors
+econ <- ggplot(
+  data = lasso[['ordinal_tidy_econ']]
+  , aes(
+    y = term
+    , x = estimate
+    , xmin = conf.low
+    , xmax = conf.high
+    , linetype = factor(model)
+    , shape = factor(model)
+  )
+) +
+  geom_pointinterval(
+    position = position_dodge(width = 0.4)
+    , size = 3
+) +
+  geom_vline(
+    xintercept = 0
+    , linetype = 2
+) +
   scale_y_discrete(limits = rev) +
   theme_minimal() +
-    labs(linetype = 'Model', shape = 'Model', x = 'Estimate', y = '', caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.') +
+  labs(
+    linetype = 'Model'
+    , shape = 'Model'
+    , x = 'Estimate'
+    , y = ''
+    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
+) + 
   theme(text=element_text(size=12, family = 'sans'))
+  #* plot for political predictors
+poli <- ggplot(
+  data = lasso[['ordinal_tidy_poli']]
+  , aes(
+    y = term
+    , x = estimate
+    , xmin = conf.low
+    , xmax = conf.high
+    , linetype = factor(model)
+    , shape = factor(model)
+  )
+) +
+  geom_pointinterval(
+    position = position_dodge(width = 0.4)
+    , size = 3
+) +
+  geom_vline(
+    xintercept = 0
+    , linetype = 2
+) +
+  scale_y_discrete(limits = rev) +
+  theme_minimal() +
+  labs(
+    linetype = 'Model'
+    , shape = 'Model'
+    , x = 'Estimate'
+    , y = ''
+    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
+) + 
+  theme(text=element_text(size=12, family = 'sans'))
+
+# Print plots
+econ
+poli
 ```
@@
 
-## Discussion
+# Discussion and conclusions
 
@@
 Though in 2020, the global economy faced significant uncertainty due to the COVID-19 pandemic, economic factors appear to be relatively less important.
-These results suggest that white political identity results from a political threat in the post-Trump era.
+These results suggest that white political identity results from perceptions of a political threat in the post-Trump era.
 
-# Conclusion
-
-The literature on social identity suggests that the centrality of one's identity and group-based attitudes and behaviors result from economic threat.
-However, white political identity has a historical and institutional significance that, while tied up in economics, reacts to changes in whites' influence over these institutions.
-
-As pitting two theories against one another is a methodological challenge, I use data from the 2012, 2016, and 2020 ANES to demonstrate whether the relationships between attitudes proxying political and economic threats are more plausible.
-While this exercise is not and should not be interpreted as causal, I do demonstrate that economic factors appear to have a weaker relationship with white identity than political attitudes - particularly in the post-Trump era.
-
-The narrative I advocate for suggests that white political identity is most responsive to threats of whites' institutional influence.
-The descriptive analysis of the relationships between political and economic threat fall short of causal identification.
-With the results of these analyses, scholars have evidence suggesting that economic threats appear to weakly associate with one's expression of a white political identity when including political threats in the same model.
-
-The task for scholars is to examine whether political threats have any causal bearing on explaining the source of white political identity.
-This undoubtedly is a difficult task as threats are hard to treat experimentally and are reasonably suspect with observational data.
+These findings provide a direction for scholars to take when examining the factors driving Whites to take on a white political identity.
+While this particular analysis is not causal, it provides a basis for those wondering whether, when put together, economic or political threats are more fruitful to engender a rise in white political identity.
 
 {{< pagebreak >}}
-<!--
-TODO:
-  - Make a strong assertion about my results
-  - Discuss why I think economic factors don't matter when they matter in most identity papers
-  - What does this mean for how we understand white identity? Elaborate on the takeaway of the paper and what I added but also what questions are still there.
--->
 
-# Appendix{#sec-appendix}
+# References
 
-## Measures{#sec-measures}
+::: {#refs}
+:::
 
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
+{{< pagebreak >}}
+
+# Supplementary Information {.appendix}
+
+## Measures {.appendix}
+
+\blandscape
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 |                                              | Abbreviated question                                       | ANES Coding                                                                                                                                                  | Author coding                                                                                                                                                             | Years            |
-+==========================================+============================================================+==================================================================================================================================================================+===========================================================================================================================================================================+==================+
++==============================================+============================================================+==============================================================================================================================================================+===========================================================================================================================================================================+==================+
 | white identity                               | How important is being white to your identity              | 1 = Extremely important; 2 = Very important; 3 = Moderately important; 4 = A little important; 5 = Not at all important; \<1 Missing, not asked, etc         | 1 = Not at all important; 2 = A little important; 3 = Moderately important; 4 = Very important; 5 = Extremely important; NA = Missing, not asked, etc                     | 2012, 2016, 2020 |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 | **Economics**                                |                                                            |                                                                                                                                                              |                                                                                                                                                                           |                  |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 | Retrospective better off                     | How much better worse off than 1 year ago                  | 1 = Much better; 2 = Somewhat better; 3 = The same; 4 = Somewhat worse; 5 = Much worse, \< 1 Missing, not asked, etc                                         | -2 = Much worse; -1 = Somewhat worse; 0 = The same; 1 = Somewhat better; 2 = Much better; NA = Missing, not asked, etc                                                    | 2012, 2016, 2020 |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 | Economy better                               | U.S.
 economy better worse off than 1 year ago              | 1 = Much better; 2 = somewhat better; 3 = Stayed about the same, 4 = Somewhat worse; 5 = Much worse; \< 1 Missing, not asked, etc                            | -2 = Much worse; -1 = Somewhat worse; 0 = The same; 1 = Somewhat better; 2 = Much better; NA = Missing, not asked, etc                                                    | 2012, 2016, 2020 |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 | Unemployment better                          | Unemployment better or worse than 1 year ago               | 1 = Much better; 2 = somewhat better; 3 = Stayed about the same, 4 = Somewhat worse; 5 = Much worse; \< 1 Missing, not asked, etc                            | -2 = Much worse; -1 = Somewhat worse; 0 = The same; 1 = Somewhat better; 2 = Much better; NA = Missing, not asked, etc                                                    | 2012, 2016, 2020 |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 | Education                                    | Educational attainment                                     | 1 = \< than High school; 2 = High school; 3 = Some post-High school; 4 = Bachelor's degree; \< 1 Missing, not asked, etc                                     | 1 = \< than High school; 2 = High school; 3 = Some post-High school; 4 = Bachelor's degree; NA = Missing, not asked, etc                                                  | 2012, 2020       |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 | Education (2016)                             | Educational attainment                                     | See ANES codebook                                                                                                                                            | See ANES codebook; NA = MIssing, not asked, etc                                                                                                                           | 2016             |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 | Income                                       | Total family income                                        | See ANES codebook; \< Missing, not asked, etc                                                                                                                | See ANES codebook; NA = MIssing, not asked, etc                                                                                                                           | 2012, 2016, 2020 |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 | Worried about losing job                     | How worried about losing job in near future                | 1 = Not at all; 2 = A little; 3 = Moderately; 4 = Very; 5 = Extremely; \< 1 Missing, not asked, etc                                                          | -2 = Not at all; -1 = A little; 0 = Moderately; 1 = Very; 2 = Extremely                                                                                                   | 2012, 2016, 2020 |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 | Immigrants take jobs                         | How likely immigration will take away jobs                 | 1 = Extremely; 2 = Very; 3 = Somewhat; 4 = Not at all; \< 1 Missing, not asked, etc                                                                          | 1 = Not at all; 2 = Somewhat; 3 = Very; 4 = Extremely; NA = Missing, not asked, etc                                                                                       | 2012, 2016, 2020 |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 | Worried about family finances                | Worry about family financial situation                     | 1 = Extremely; 2 = Very; 3 = Moderately; 4 = A little; 5 = Not at all; \< 1 Missing, not asked, etc                                                          | 1 = Not at all, 2 = A little; 3 = Moderately; 4 = Very; 5 = Extremely; NA = Missing, not asked, etc                                                                       | 2012, 2016, 2020 |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 | Know someone lost job                        | Know someone who lost job                                  | 1 = Someone lost job; 2 = No one lost job; \< 1 Missing, not asked, etc                                                                                      | 0 = No one lost job; 1 = Someone lost job; NA = Missing, not asked, etc                                                                                                   | 2012, 2016, 2020 |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 | **Political**                                |                                                            |                                                                                                                                                              |                                                                                                                                                                           |                  |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 | New lifestyles breaking society              | Newer lifestyles breaking down society                     | 1 = Agree strongly; 2 = Agree somewhat; 3 = Neither; 4 = Disagree somewhat; 5 = Disagree strongly; \< 1 Missing, not asked, etc                              | -2 = Disagree strongly; -1 = Disagree somewhat; 0 = Neither, 1 = Agree somewhat; 2 = Agree strongly                                                                       | 2012, 2016       |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 | Government biased against whites             | Does the Administration favor Blacks or whites (2012 only) | 1 = Favors whites; 2 = Favors Blacks; 3 = Treats both the same; \< 1 Missing, not asked, etc                                                                 | -1 = Favors whites; 0 = Treats both the same; 1 = Favors Blacks; NA = Missing, not asked, etc                                                                             | 2012, 2016, 2020 |
@@
 |                                              | Does the Federal Gov treat Blacks or whites Better         |                                                                                                                                                              |                                                                                                                                                                           |                  |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 | whites influence politics                    | How much influence do whites have on U.S.
 politics         | 1 = Too much; 2 = Just about right; 3 = Too little; \< 1 Missing, not asked, etc                                                                             | -1 = Too little; 0 = Just about right; 1 = Too much; NA = Missing, not asked, etc                                                                                         | 2012, 2016, 2020 |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 | whites are discriminated against             | Discrimination in U.S.
 against whites                      | 1 = A great deal; 2 = A lot; 3 = A moderate amount; 4 = A little; 5 = None at all; \< 1 Missing, not asked, etc                                              | 1 = None at all; 2 = A little; 3 = A moderate amount; 4 = A lot; 5 = A great deal; NA = Missing, not asked, etc                                                           | 2012, 2016, 2020 |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 | **Controls**                                 |                                                            |                                                                                                                                                              |                                                                                                                                                                           |                  |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 | Racial Resentment                            | Standard, 4-item, battery                                  | 5-point Likert scale                                                                                                                                         | Avg of four questions; NA = Missing, not asked, etc                                                                                                                       | 2012, 2016, 2020 |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
-| Party ID                                 | Standard battery                                           | 1 = Strong Democrat; 2 = Democrat; 3 = Leans Democrat; 4 = Independent; 5 = Leans Republican; 6 = Republican; 7 = Strong Republican; \< 1 Missing, not asked, etc| 1 = Strong Democrat; 2 = Democrat; 3 = Leans Democrat; 4 = Independent/Neither; 5 = Leans Republican; 6 = Republican; 7 = Strong Republican; \< 1 Missing, not asked, etc | 2012, 2016, 2020 |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
+| Party ID                                     | Standard battery                                           | 1 = Strong  Dem; 2 = De; 3 = Leans Dem; 4 = Independent; 5 = Leans Rep; 6 = Rep; 7 = Strong Rep; \< 1 Missing, not asked, etc                                | 1 = Strong Democrat; 2 = Democrat; 3 = Leans Democrat; 4 = Independent/Neither; 5 = Leans Republican; 6 = Republican; 7 = Strong Republican; \< 1 Missing, not asked, etc | 2012, 2016, 2020 |
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
 | Female                                       | Sex of respondent                                          | 1 = Male, 2 = Female; \< 1 Missing, not asked, etc                                                                                                           | 0 = Male, 1 = Female; NA = Missing, not asked, etc                                                                                                                        | 2012, 2016, 2020 |
-+------------------------------------------+------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
-: ANES Measures {#tbl-anes-measures}
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
+: ANES Measures {#tbl-anes-measures tbl-colwidths="[20,20,25,25,10]"}
 
+\elandscape
 {{< pagebreak >}}
 
-## Descriptive statistics {#sec-desc-stats}
+## Descriptive statistics {.appendix}
+
+```{r}
+#| label: loading-data-block
+#create empty list object to hold datasets in
+data <- list()
+# load the cleaned data
+    #* 2012
+data[["anes_2012"]] <- read.csv("../data/clean/anes_2012_updated.csv")
+
+data[["anes_2012"]] <- data[["anes_2012"]][
+  , c(
+    "wid", "rboff", "epast", "unpast", "edu", "income", "losejob"
+    , "ecfamily", "ecjob", "tradbreak", "govbias", "inflwhite"
+    , "whitediscrim", "raceResent", "pid", "female"
+    )
+]
+data[["anes_2012"]] <- setnames(
+  data[["anes_2012"]]
+  , old = c(
+    "wid", "rboff", "epast", "unpast", "edu", "income", "losejob"
+    , "ecfamily", "ecjob", "tradbreak", "govbias", "inflwhite"
+    , "whitediscrim", "raceResent", "pid", "female"
+  )
+  , new = c(
+    "White identity", "Retrospective better off", "Economy better", "Unemployment better", "Education", "Income", "Worried about losing job"
+    , "Worried about family finances", "Know someone lost job", "New lifestyles breaking tradition", "Government biased against whites", "Whites influence politics"
+    , "Whites discrimminated against", "Racial resentment", "Party ID", "Female"
+  ))
+
+    #* 2016
+data[["anes_2016"]] <- read.csv("../data/clean/anes_2016_updated.csv") 
+data[["anes_2016"]] <- data[["anes_2016"]][
+  , c(
+    "wid", "rboff", "epast", "unpast", "edu", "income", "losejob"
+    , "ecfamily", "ecjob", "tradbreak", "govbias", "inflwhite"
+    , "whitediscrim", "raceResent", "pid", "female"
+    )
+]
+data[["anes_2016"]] <- setnames(
+  data[["anes_2016"]]
+  , old = c(
+    "wid", "rboff", "epast", "unpast", "edu", "income", "losejob"
+    , "ecfamily", "ecjob", "tradbreak", "govbias", "inflwhite"
+    , "whitediscrim", "raceResent", "pid", "female"
+  )
+  , new = c(
+    "White identity", "Retrospective better off", "Economy better", "Unemployment better", "Education", "Income", "Worried about losing job"
+    , "Worried about family finances", "Know someone lost job", "New lifestyles breaking tradition", "Government biased against whites", "Whites influence politics"
+    , "Whites discrimminated against", "Racial resentment", "Party ID", "Female"
+  ))
+    #* 2020
+data[["anes_2020"]] <- read.csv("../data/clean/anes_2020_updated.csv") 
+data[["anes_2020"]] <- data[["anes_2020"]][
+  , c(
+    "wid", "rboff", "epast", "unpast", "edu", "income", "losejob"
+    , "ecfamily", "ecjob", "govbias", "inflwhite"
+    , "whitediscrim", "raceResent", "pid", "female"
+    )
+]
+data[["anes_2020"]] <- setnames(
+  data[["anes_2020"]]
+  , old = c(
+    "wid", "rboff", "epast", "unpast", "edu", "income", "losejob"
+    , "ecfamily", "ecjob", "govbias", "inflwhite"
+    , "whitediscrim", "raceResent", "pid", "female"
+  )
+  , new = c(
+    "White identity", "Retrospective better off", "Economy better", "Unemployment better", "Education", "Income", "Worried about losing job"
+    , "Worried about family finances", "Know someone lost job", "Government biased against whites", "Whites influence politics"
+    , "Whites discrimminated against", "Racial resentment", "Party ID", "Female"
+  ))
+```
 
@@
 #| tbl-cap: 2012 ANES Descriptive Statistics
-
-data[['anes_2012']] |>
-  rename(`white identity` = wid, `Retrospective better off` = rboff, `Economy better` = epast,    `Unemployment better` = unpast, `Education` = edu, `Income` = income, `Worried about losing job` = losejob, `Worried about family finances` = ecfamily, `Know someone lost job` = ecjob, `New lifestyles breaking society` = tradbreak, `Government biased against whites` = govbias, `whites influence politics` = inflwhite, `whites are discriminated against` = whitediscrim, `Racial resentment` = raceResent, `Party ID` = pid, `Female` = female) |>
-  datasummary_skim(notes = c('NAs are included in Unique column.', 'Data source: 2012 American National Election Study, unweighted.'))
+datasummary_skim(
+  data[["anes_2012"]]
+  , histogram = FALSE
+  , notes = c(
+    "NAs are included in Unique column."
+    , "Data source: 2012 American National Election Study, unweighted."
+  )
+  , output = "latex"
+)
 ```
@@
 
-data[['anes_2016']] |>
-  rename(`white identity` = wid, `Retrospective better off` = rboff, `Economy better` = epast,    `Unemployment better` = unpast, `Education` = edu, `Income` = income, `Worried about losing job` = losejob, `Worried about family finances` = ecfamily, `Know someone lost job` = ecjob, `New lifestyles breaking society` = tradbreak, `Government biased against whites` = govbias, `whites influence politics` = inflwhite, `whites are discriminated against` = whitediscrim, `Racial resentment` = raceResent, `Party ID` = pid, `Female` = female) |>
-  datasummary_skim(notes = c('NAs are included in Unique column.', 'Data source: 2016 American National Election Study, unweighted.'))
+datasummary_skim(
+  data[["anes_2016"]]
+  , histogram = FALSE
+  , notes = c(
+    "NAs are included in Unique column."
+    , "Data source: 2016 American National Election Study, unweighted."
+  )
+  , output = "latex"
+)
 ```
@@
 
-data[['anes_2020']] |>
-  rename(`white identity` = wid, `Retrospective better off` = rboff, `Economy better` = epast,    `Unemployment better` = unpast, `Education` = edu, `Income` = income, `Worried about losing job` = losejob, `Worried about family finances` = ecfamily, `Know someone lost job` = ecjob, `Government biased against whites` = govbias, `whites influence politics` = inflwhite, `whites are discriminated against` = whitediscrim, `Racial resentment` = raceResent, `Party ID` = pid, `Female` = female) |>
-  datasummary_skim(notes = c('NAs are included in Unique column.', 'Data source: 2020 American National Election Study, unweighted.'))
+datasummary_skim(
+  data[["anes_2020"]]
+  , histogram = FALSE
+  , notes = c(
+    "NAs are included in Unique column."
+    , "Data source: 2020 American National Election Study, unweighted."
+  )
+  , output = "latex"
+)
 ```
@@
 
-## Model results
+## Model results {.appendix}
+
 ```{r}
@@
 #| tbl-cap: Predictors of white Identity in 2012
-
-rows <- tribble(~term, ~`2012 LWD`, 'Economic', '', 'Non-material', '', 'Demographics', '', 'Thresholds', '')
+model_2012 <- list(
+  '2012 LWD' = lasso[["2012_simple"]],
+  '2012' = lasso[["2012_ordinal"]]
+)
+rows <- tribble(~term, ~`2012 LWD`, ~`2012`, 'Economic', '', '', 'Non-material', '', '', 'Demographics', '', '', 'Thresholds', '', '')
 attr(rows, 'position') <- c(1, 18, 27, 32)
 
-modelsummary(lasso[['2012 LWD']], coef_map = cm, statistic = 'conf.int', gof_map = gm, add_rows = rows, notes = list('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2012 American National Election Study.'))
+modelsummary(model_2012, coef_map = cm, statistic = 'conf.int', gof_map = NA, add_rows = rows, notes = list('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2012 American National Election Study.'))
+
 ```
@@
 #| tbl-cap: Predictors of white identity in 2016
-
-rows <- tribble(~term, ~`2016 LWD`, 'Economic', '', 'Non-material', '', 'Demographics', '', 'Thresholds', '')
+model_2016 <- list(
+  '2016 LWD' = lasso[["2016_simple"]],
+  '2016' = lasso[["2016_ordinal"]]
+)
+rows <- tribble(~term, ~`2016 LWD`, ~`2016`, 'Economic', '', '', 'Non-material', '', '', 'Demographics', '', '', 'Thresholds', '', '')
 attr(rows, 'position') <- c(1, 18, 27, 32)
 
-modelsummary(lasso[['2016 LWD']], coef_map = cm, statistic = 'conf.int', gof_map = gm, add_rows = rows, notes = list('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2016 American National Election Study.'))
+modelsummary(model_2016, coef_map = cm, statistic = 'conf.int', gof_map = NA, add_rows = rows, notes = list('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2016 American National Election Study.'))
+
 ```
@@
 #| tbl-cap: Predictors of white identity in 2020
-
-rows <- tribble(~term, ~`2020 LWD`, 'Economic', '', 'Non-material', '', 'Demographics', '', 'Thresholds', '')
+model_2020 <- list(
+  '2020 LWD' = lasso[["2020_simple"]],
+  '2020' = lasso[["2020_ordinal"]]
+)
+rows <- tribble(~term, ~`2020 LWD`, ~`2020`, 'Economic', '', '', 'Non-material', '', '', 'Demographics', '', '', 'Thresholds', '', '')
 attr(rows, 'position') <- c(1, 18, 25, 30)
 
-modelsummary(lasso[['2020 LWD']], coef_map = cm, statistic = 'conf.int', gof_map = gm, add_rows = rows, notes = list('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2020 American National Election Study.'))
+modelsummary(model_2020, coef_map = cm, statistic = 'conf.int', gof_map = NA, add_rows = rows, notes = list('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2020 American National Election Study.'))
 ```
-
-```{r}
-#| label: tbl-2012-2016-2020-ordinal-lasso
-#| tbl-cap: Predictors of white identity in 2012, 2016, and 2020
-
-model <- list(
-    '2012 ANES' = lasso[['2012']],
-    '2016 ANES' = lasso[['2016']],
-    '2020 ANES' = lasso[['2020']]
-)
-rows <- tribble(~term, ~`2012 ANES`, ~`2016 ANES`, ~`2020 ANES`, 'Economic', '', '', '','Non-material', '', '', '', 'Demographics', '', '', '', 'Thresholds', '','', '')
-attr(rows, 'position') <- c(1, 18, 27, 32)
-
-modelsummary(model, coef_map = cm, statistic = 'conf.int', gof_map = gm, add_rows = rows, notes = list('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.' ,'Data Source: 2012, 2016, and 2020 American National Election Study.'))
-```

### 2023-06-27 src/manuscript.qmd
make a general draft

@@
     cite-method: citeproc
-    latex-output-dir: ../out/submissions/bjps
+    latex-output-dir: ../out/
     filters:

### 2023-07-07 src/manuscript.qmd
doing some adjustments to front half and adding extra models

@@
 
-Those conducting post-mortem analyses of the 2016 Presidential election of Donald J.
-Trump, a presidential candidate that personifies white grievance politics, contend that the substantial presence of non-college-educated white supporters explains his electoral success [@griffin_et-al_2017].
-Analyses by political scientists largely attribute Trump's success in attracting voters to those who see their whiteness as a politically relevant identity where they are distinct from Americans of other racial groups [@jardina_2020_pp].
-What remains an open question is whether it actually is whites' concerns about economic status that motivates whites to use their race as a lens to evaluate politics or whether it is a backlash to changes in the racial groups reflected by those occupying visible political offices.
+Are White identifiers those who are percieve economic threats or is it a backlash to a loss of influence over politics?
+The existing work offers an unclear answer to this question.
+This note attempts to provide an ombnibus test to determine not only which of these questions has more empirical support but also which among common proxies of economic and political power in the United States appear to be reliable predictors of a White individual's expression of racial identity.
 
-While this is an ongoing debate, part of this disagreement between scholars about where white identity comes from may be an artifact of the traditions of the theoretical frameworks we use in this area.
-Scholars studying identity, more broadly, often focus on economic threats as a critical motivator for an expressed connection to a social in-group [see @huddy_2001].
-However, some scholars studying white political attitudes and racial identity tend to focus on the role of political threats [see @bartels_2020].
-This mismatch leads to some conceptualizing white identity as originating from economic views in the tradition of the identity literature.
-In contrast, others conceptualize white identity as motivated by political threats.
-
-Along with providing a narrative that scholars might use to identify potential mechanisms connecting political threats to white identity, I examine the correlates of white identity.
-Using data from the 2012, 2016, and 2020 American National Election Study, I utilize penalized regression to test the claim that a handful of predictors that reflect attitudes about political power are more influential than general (sociotropic) and pocketbook (egotropic) attitudes about the economy.
-I find that proxies of political threat appear stronger than economic attitudes in the post-Trump era.
-As this particular analysis is not a causal test of the mechanism connecting threat to white identity, I close by arguing that scholars should continue to examine the mechanisms that explain this connection.
+A common source of data and of measures for researchers in the United States is the American National Election Study.
+I use a penalized regression using data from the 2012, 2016, and 2020 American National Election Studies to examine the degree to which individual measures of economic and political influence explain an individual's uptake of White identity as well as to examine the degree to which changes in the political context may increase or decrease the explanatory power of these measures.
+I find that proxies reporting concerns about Whites' political influence are stronger explantory factors than proxies of economic influence in the post-Trump era.
 
@@
 
+One perspective on what explains racial identity among Whites in the United States are that they are responding to threats to their economic status and power.
+Realistic Conflict Theory suggests that a racial identity and racial prejudice come as a response to a collective concern among members of a particular racial group about the loss of economic influence and status a group holds relative to other racial groups [@blumer_1958_psr;@bobo_hutchings_1996_asr].
+Some work in support of this perspective sees White racial identity as a response to concerns about increased job market competition through immigration [@pardos-prado_xena_2019].
+Other work finds support for this argument in lab experiments [@perez_et-al_2022_pp] and in surveys [@lindsay_2023_apr].
+In one of the most thorough treatments on the use of racial identity by Whites in politics, @jardina_2019_cup includes indicators of economic status and power in the primary model to predict White political identity.
+If this perspective holds, we might expect that in times of economic hardship, Whites may be most sensitive to economic threats as the collective and the individual see economics as most salient to them.
+
+Social identity theory, and its offshoots, would suggest that Whites expressing a racial identity are responding to political, cultural, and societal threats.
+While these threats are a bit more hard to define, there are a number of examples of Whites expressing prejudice towards other racial groups in response to these sorts of challenges to their power.
+@tesler_2016 demonstrates that Whites heavily racialized American politics in response to Barack Obama's ascention to the White House.
+Other work demonstrates that the saliency and reprocussions of aligning with competing racial groups drove non-White political candidates to distance themselves from their racial confederates in an attempt to appease Whites [@stephens-dougan_2020].
+Some pundits on Fox News suggest that Whites are in direct competition for jobs with immigrants from South and Central America.
+Despite these claims, some evidence suggests that Whites' hostility directed toward non-White immigrant groups depends on the degree to which they assimilate [@green_staerkle_2013; see also @hainmueller_hopkins_2014].
+In terms of attitudes oriented toward the in-group, @dawson_1994_pup argues that Black Americans hold a strong racial identity that supercedes their class-based identity due to the persistent focus on Black individual's race in their personal and political life.
+For Whites, their racial identity has long been central to their behavior and the way in which they mobilize to shape legal and political institutions to maintain power [@rile-hayward_2013].
+
+We may also see that economic or political threat may appear differently for partisans and for female respondents.
+Recent evidence suggests that racial attitudes are guided by partisanship [@engelhardt_2021_bjps].
+Specifically, some evidence suggests that White Democrats are much more sensitive to their racial identity than White Republicans due to the higher amount of racial heterogeneity within the Democratic party.
+This may mean for White Democrats that they have more exposure to intergroup competition for power within the party (such as the choice to support Hillary Clinton versus Barack Obama) and may have higher White identification as a result.
+The second way that these threats may matter more or less for Whites is on the dimension of sex.
+Gender-based differences in economic influence and status remains quite high.
+We therefore may expect that female individuals are more sensitive to economic threats than their male counterparts.
+In addition to the ombnibus model, I also include two additional models where I split the samples along the dimension of partisanship (including leaners in each) and on the dimension of sex.
+
+# Political versus economic threats
+
+Identity is largely an elusive concept to operationalize.
+Though there are implicit measures of White identity [see @knowles_peng_2005_jpsp], the common and measure of white identity is a response to the question, "How important is being White to your identity?" This measure captures a significant feature of an identity: the centrality or the importance of that identity to an individual.
+Though a rather rough measure of something as complex as identity, it has been shown to produce similar results to more complex measures of white identity and holds a significant amount of face validity [@jardina_2019_cup].
+
+Threat is also quite difficult to measure directly.
+However, we can make a number of reasonable guesses about who may feel economic and political threat.
+There are two key things that threats affect.
+Status and power.
+Status refers to the respectability of the group whereas power refers to the influence that the group has [@craig_phillips_2023_ohpp].
+Threat here refers to either a manifestation of economic or political challenges to the current racial hierarchy.
+I do not distinguish between these two sources of threat as common survey measures are too crude to distinguish between how the threat manifests.
+Here, I am focusing on what types of non-material threats activate White identity in the context of politics and will, as a shorthand refer to political, cultural, and societal threats as political threats.
+
+A reasonable approach to measuring economic threat is to measure the economic circumstances that White individuals find themselves in.
+We may expect that Whites who are facing bottom-up economic competition, will be those that experience the strongest desire to identify themselves with their racial group.
+As the dominant racial group in the United States, Whites that are unemployed, are surrounded by other Whites who are experiencing economic hardship and those who have limited capacity to switch careers are those that may be the most motivated to identify with the dominant racial group to defend Whites' influence and status in society [@bobo_hutchings_1996_asr].
+Therefore, measures of employment, finances, family finances, finances of those they know, and educational attainment all seem likely predictors of racial identity for Whites.
+This fits with measures that @dawson_1994_pup used to measure the influence of economics in Black American's identity.
+These measures also fit with @lindsay_2023_apr who found that many of these items predicted racial resentment among Whites.
+
+Here I conceptualize of political threat as perceptions that politics and political institutions no longer serves White people.
+Existing work demonstrates that Whites who express a desire to maintain the racial status quo with Whites on top are more likely to express that Whites are discriminated against at similar levels as Blacks [@marshburn_et-al_2022_gpir].
+As Whiteness is often a central component of Whites' conceptualization of American society [see @green_staerkle_2013], changes to "traditional American values" may reflect concern about changes to the influences of White values.
+Finally, a common finding is that expression of White identity is heavily wrapped up in the perception that racial stratifications are by meritocratic means [@lowery_et-al_2012_jpsp].
+As members of a dominant social group, when given information that either threatens one's belief that their merits are earned or that their group operates as a way to maintain inequality, we expect lower levels of expression in identification with that group.
+The classic items of racial resentment construct capture these ideas in that they revolve around the idea that Blacks are not disadvantaged as the result of systematic oppression by Whites.
+Though many have often used racial resentment as a measure of prejudice which would be a possible outcome of White identity, some make the compelling argument that it is distinct as a concept and operationalization [@davis_wilson_2022_ucp].
+
+<!--
 Social identity theory suggests that individuals form attitudes based on groups they associate with, even if those groups are subjectively defined [see @tajfel_turner_1986].
@@
 Whiteness and one's willingness to identify with that group is important [@wong_cho_2005] as it reflects a desire for an association with the history and affordances of being white [@mcdermott_samson_2005].
--->
+->
 Recent work makes the case that we should seriously consider whites' racial identity as an identity [@jardina_2019_cup].
@@
 The goal is not to test particular causal mechanisms but to examine whether my belief and others' that white identity reflects reactions to political threats more than it does economic.
-
+-->
 # Data
@@
 The studies are nationally representative questions that capture several questions that lend themselves to testing the relationship between economic and political concerns while also asking white respondents to report how important being white is to their identity.
-<!--
-Of course, I cannot *prove* that these questions reflect the two types of threats I purport to represent.
-This leaves any evidence of the particular mechanism connecting these predictors to reported levels of white identity -- among other reasons -- suspect.
-However, the goal of this paper is not necessarily to test a particular causal mechanism -- this is a much larger project that we should continue to work on collectively.
-Nevertheless, the paper aims to offer an argument that contextualizes white identity in history and how that leads to different predictions about the correlates of white identity than many might assume from understanding the literature on the origins of social identity.
--->
-Each study asks white respondents, "How important is being white to your Identity" and offers them a five-point Likert scale with the options reflecting their level of agreement with the question.^[Details about question-wording, original ANES coding, and my coding for all variables I include in my models are included in @tbl-anes-measures located in the Supplementary Information.
-I include descriptive statistics for the measures in these samples in @tbl-2012-descriptive-stats, @tbl-2016-descriptive-stats, and @tbl-2020-descriptive-stats located in the Supplementary Information.] They do not ask white respondents to report on items from the white political identity measures of @jardina_2019_cup or the more complex battery of items that @helms_carter_1990 discusses.
-However, this question does capture the degree to which one sees their whiteness as part of who they are and reflects the operationalization that others have used [see @schildkraut_2017;@jardina_2019_cup;@perez_et-al_2022_pp].
 
-The ANES does not ask questions that explicitly capture the attitudes of respondents about perceptions of the economy for different groups.
-That is, there are no questions asking respondents about their perceptions of the economy for Black Americans relative to white Americans.
-The literature suggests, however, that those who have lower incomes are more likely to feel insecure about their financial status and are more likely to use this economic threat as a motivation for their negative attitudes toward immigrants [@pardos-prado_xena_2019].
-Therefore, one measure of economic threat that I use is the respondent's reported level of total family income.
-Those with lower levels of education are also likely to feel similar levels of insecurity about their economic status, as their ability to seek higher-paying jobs is likely more limited.
-Whites who report that they feel that their family is doing worse relative to the previous year and knowing others who lost a job also reflect a degree of insecurity about their economic status and may reflect perceptions of economic threat [@jardina_2019_cup].
-I include an item reflecting respondents' perceptions of economic mobility today relative to the past.
-In 2012, whites harboring negative attitudes toward Obama rooted in racial animous were more likely to express more negative views of the economy relative to how it had been in the past [@sides_et-al_2018].
-Others demonstrate that those who are reacting to economic threats and are concerned about their and other whites' status evaluate the economy and their economic position as worse [@baccini_weymouth_2021_apsr].
-Therefore, I additionally include questions about retrospective sociotropic evaluations of the economy.
+Though I am using the more crude measure of White identity and am limited in the predictors I can choose from, a larger concern is that the degree to which these threats may vary as a result of changes to the political landscape.
+Specifically, in 2008 the United States experienced a massive recession that lead hit the housing market quite hard.
+In 2012, the economy was still recovering.
+In 2016, Donald Trump became infamous for an election where he made race and gender extremely salient [@banda_cassese_2022_pb].
+Finally, in 2020, Trump had been in office for four years with an improving economy.
+Each of these three years reflect a different context and may offer the opportunity to observe how economic and political threats operate differrently.
 
-The ANES includes several questions that reflect perceptions of whites losing political power.
-First, I include a question that asks respondents to report whether or not they felt that whites currently have too little or too much influence on politics today.
-I also want to examine the degree to which respondents feel that the institutions benefit non-whites rather than whites by including the question that whites face discrimination and that the government is biased against whites.
-I include a question asking respondents about their belief that traditions in this country are breaking.^[The ANES only included this question in the 2012 and 2016 questionnaires.] While these, like the economic attitudes, are just proxies of threat, they reflect the extent to which respondents may feel insecure about the political influence of whites.
-
-Furthermore, the ANES includes the racial resentment battery, which asks questions reflecting whites' beliefs that Blacks are individually responsible for their circumstances and undeserving of assistance.
-I control for racial resentment as the individual items likely confound the relationship between both economic and political threat measures on expressed white identity.
-As economic evaluations often reflect partisans reporting a better economy when the government is run by co-partisans and a worse economy when the other party runs the government [@achen_bartels_2017], I additionally control for a 7-point item measure of partisanship.
-Finally, I control for whether or not the respondent is female, as Sex affects one's socioeconomic status.
+Question wording for each of the items in the following models are included in @tbl-anes-measures which is located in the Supplementary Materials.
+I include descriptive statistics for the measures in these samples in @tbl-2012-descriptive-stats, @tbl-2016-descriptive-stats, and @tbl-2020-descriptive-stats located in the Supplementary Information.
 
@@
 
-The summary of the literature points to a large number of potential predictors that influence the degree to which one takes on white political identity.
-To adjudicate between hypotheses, there is value in a comprehensive model that considers the plausibility of each.
-Most models consider those predictors they believe are most plausible and do not include others that the researcher believes to be less plausible [see @clarke_primo_2012].
-They understandably do this out of concern with increases in the inefficiency of one's regression estimates: the more predictors you include, the risk of overfitting your model to your particular sample and from multicollinearity increases, thereby erroneously generalizing conclusions drawn from the fitted statistical model.
-Though this concern is reasonable, such a model neglects to examine the effects of the predictors a researcher believes to be less plausible.
-
-Therefore, I use a penalized regression model to balance these concerns.
-The literature surveyed above suggests that there are a small number of variables that do predict white political identity, and there are many others that have no effect.
-Imagining an assumed hypothesis about the distribution of coefficients, this implies that the L2 regularization parameter equals 0.
-This is because the L2 regularization parameter shrinks coefficients asymptotically toward zero.
-The L1 regularization parameter, however, does allow for a shrinkage parameter to pull coefficients to 0.
-Given this, it suggests that the more appropriate model to capture this debate would be the LASSO penalized regression.
-
-Rather than use the classical LASSO with Leave One Out Cross Validation (LOOCV), I take advantage of a type of Markov Chain Monte Carlos (MCMC) called a Hamiltonian Monte Carlo for my LASSO due to a number of benefits outlined by @van-erp_et-al_2019_jmp.
-As the literature leads me to hold *apriori* expectations about the distribution of predictors that meaningfully effect white political identity, I specify a Laplace distribution as my prior density function.
-I suspect that there are a large number of included predictors that have null effects and a relatively smaller proportion of predictors that *do* have an effect but am relatively unsure about how large their coefficients might be.
-This characterization of my prior fits with that of an L2 Shrinkage parameter specified in a LASSO regression.
-As I am not using LOOCV, I am able to interpret the results of the model through my credible intervals.
+The primary contribution of this paper is to examine which of a number of competing measures of different types of threat may predict white identity for individuals.
+As this requires an ombnibus test, I elect to use a penalized regression.
+I provide a more detailed discussion of the reasons why I use an estimator with an L2 regularization parameter, often refered to as a LASSO model in the supplementary materials.
+For a brief comment as to why I use this approach is because approach is designed to take models with a large number of predictors and to shrink the coefficients toward zero.
+In practice, this often means that predictors that has relatively fewer explanatory power have much smaller coefficients than the variables that do provide significant predictive power.
 
@@
     , xmax = conf.high
-    , linetype = factor(model)
     , shape = factor(model)
@@
   geom_pointinterval(
-    position = position_dodge(width = 0.4)
-    , size = 3
+    position = position_dodge(width = 0.6)
+    , size = 4
 ) +
@@
   labs(
-    linetype = 'Model'
-    , shape = 'Model'
+    shape = 'Model'
     , x = 'Estimate'
@@
 
-Examining the figure, it is readily apparent that in the face of indicators of attitudes about political status, economic attitudes inconsistently associate with white political identity.
-In all three years, reported attitudes about the nation's economy being better relative to the previous year, reported feelings about the nation's unemployment rate, income, concern about one's family finances, and knowing someone who lost a job are not meaningfully correlated with white political identity.
-For other indicators of economic threat, none appear to be consistently meaningful.
-Retrospective evaluations of one's financial situation and education matter in 2012, but these effects were not present in 2016 and 2020.
+Examining the figure, we see that none of the proxies for economic threat consistently correlate with White political identity; however there is at least one proxy of political threat that does.
+For economic threats, we see that when the economy was recovering from the recession and Obama was in Office, those who perceived that the economy was better off retrospectively and with a lower reported level of education were correlated with expressing a higher White identity.
+However, when the economy was better and Trump was in office, we see that those less worried about losing a job were reported a higher white identity.
 
-Turning to indicators of political threat, we still see significant inconsistency.
-In both 2012 and 2016, whether someone feels that new lifestyles are breaking society positively affects reported levels of white political identity.
-This relationship is meaningful in 2012 and 2016 (this question is not present in the 2020 questionnaire).
-In 2020, those who reported that the government favored Black Americans also reported a higher white political identity, and those who feel that whites are discriminated against; however, neither of these effects are meaningful in 2012 or 2016.
+For political threats, we see that in both instances where Obama was in the White House that those who reported being concerned with "new lifestyles" as well as those with a higher score on the racial resentment measures were associated with having higher white political identities.
+When Trump was in the White House, we see that those who felt that the government was biased against Whites, that Whites were discriminated against, and who scored higher on the racial resentment measures all reported higher levels of White political identity.
 
@@
 
-Overall, these results suggest that the effects of economic threat on white political identity are more robust in the pre-Trump era.
-Again, this is not confirmatory as there is no test of a causal mechanism for either economic or political threats.
-The goal of this analysis is to examine whether, in an omnibus test, either of these threats appears to be a more plausible predictor of white political identity.
-Though the relationship between political threats is inconsistent, their relationships appear more likely; particularly when we consider how many of these attitudes are explained by a rather complex data-generating process.
+Overall, these results suggest that common proxies of economic threat are at least inconsistent measures of White identity.
+Even in a time where the economy is recovering from a collapse, more proxies of political threat seemed to predict White identity.
+When a White president well-known for his penchance for racial dog-whistles takes office, we see support for the idea that White grievance politics was at a high -- high those who felt that political institutions and that the direction of society was turning away from the racial status quo were those that reported their racial identity as more central to who they are.
 
-Economic threats tend to play a more significant part in explaining white identification in 2012.
-Political threats, however, appear to be a stronger predictor of white identification in the post-Trump era.
-In 2012, the U.S.
-economy was recovering from a massive recession and housing market collapse.
-However, as Obama's administration continued, conservatives became more outspoken about Obama's race [@tesler_2016].
-This shift to focus on Obama's race continued through the 2016 Presidential election and may reflect a much more explicit message to whites about political threats [@sides_et-al_2018].
-Though in 2020, the global economy faced significant uncertainty due to the COVID-19 pandemic, economic factors appear to be relatively less important.
-These results suggest that white political identity results from perceptions of a political threat in the post-Trump era.
+There are a few important points that are illustrated here.
+First, suggestions that White political identity is something occuring for those who are most economically threatened appear to have weak theoretical and empirical backing when directly pitted against other forms of threat.
+Second, that measures of cultural or societal threat are important variables to include if one is using white identity as a variable in a model.
 
@@
 
-## Model results {.appendix}
+## Model{.appendix}
+
+The literature surveyed above suggests that there are a small number of variables that do predict white political identity, and there are many others that have no effect.
+Imagining an assumed hypothesis about the distribution of coefficients, this implies that the L2 regularization parameter equals 0.
+This is because the L2 regularization parameter shrinks coefficients asymptotically toward zero.
+The L1 regularization parameter, however, does allow for a shrinkage parameter to pull coefficients to 0.
+Given this, it suggests that the more appropriate model to capture this debate would be the LASSO penalized regression.
+
+Rather than use the classical LASSO with Leave One Out Cross Validation (LOOCV), I take advantage of a type of Markov Chain Monte Carlos (MCMC) called a Hamiltonian Monte Carlo for my LASSO due to a number of benefits outlined by @van-erp_et-al_2019_jmp.
+As the literature leads me to hold *apriori* expectations about the distribution of predictors that meaningfully effect white political identity, I specify a Laplace distribution as my prior density function.
+I suspect that there are a large number of included predictors that have null effects and a relatively smaller proportion of predictors that *do* have an effect but am relatively unsure about how large their coefficients might be.
+This characterization of my prior fits with that of an L2 Shrinkage parameter specified in a LASSO regression.
+As I am not using LOOCV, I am able to interpret the results of the model through my credible intervals.
+
+### Full tables of results {.appendix}
 

### 2023-07-07 src/eda/dcr_anes_2012_cleaning.R

@@ -18,7 +18,7 @@ box::use(
     #* create list object
 anes_2012 <- list()
     #* Load original ANES file ----
-anes_2012[["original"]] <- read_dta("../../../data/original/anes_2012/anes-2012.dta")
+anes_2012[["original"]] <- read_dta("../data/original/anes_2012/anes-2012.dta")
 # Cleaning ----
 anes_2012[["clean"]] <- anes_2012[["original"]] |>
     #* Filter dataset to include only white respondents ----
@@ -167,4 +167,4 @@ anes_2012[["clean"]] <- anes_2012[["original"]] |>
     )
 
 # Save data
-write.csv(anes_2012[["clean"]], '../../../data/clean/anes_2012_updated.csv')
+write.csv(anes_2012[["clean"]], '../data/clean/anes_2012_updated.csv')

### 2023-07-07 src/eda/dcr_anes_2016_cleaning.R

@@ -18,7 +18,7 @@ box::use(
     #* create list object
 anes_2016 <- list()
     #* load dataset
-anes_2016[["original"]] <- read_dta("../../../data/original/anes_2016/anes-2016.dta")
+anes_2016[["original"]] <- read_dta("../data/original/anes_2016/anes-2016.dta")
 
 # Cleaning
 anes_2016[["clean"]] <- anes_2016[["original"]] |>
@@ -177,4 +177,4 @@ anes_2016[["clean"]] <- anes_2016[["original"]] |>
     )
 
 # Save clean data
-write.csv(anes_2016[["clean"]], "../../../data/clean/anes_2016_updated.csv")
+write.csv(anes_2016[["clean"]], "../data/clean/anes_2016_updated.csv")

### 2023-07-07 src/eda/dcr_anes_2020_cleaning.R

@@ -18,7 +18,7 @@ box::use(
     #* Create list object
 anes_2020 <- list()
     #* Load dataset
-anes_2020[["original"]] <- read_dta("../../../data/original/anes_2020/anes_2020_original.dta")
+anes_2020[["original"]] <- read_dta("../data/original/anes_2020/anes_2020_original.dta")
 
 # Clean the data
 anes_2020[["clean"]] <- anes_2020[["original"]] |>
@@ -160,4 +160,4 @@ anes_2020[["clean"]] <- anes_2020[["original"]] |>
     )
 
 # Save the clean data
-write.csv(anes_2020[["clean"]], "../../../data/clean/anes_2020_updated.csv")
+write.csv(anes_2020[["clean"]], "../data/clean/anes_2020_updated.csv")

### 2023-07-07 src/models/model_fitting.R

@@ -9,9 +9,9 @@
 
 # Setup
   #* execute cleaning scripts
-source("./src/eda/dcr_anes_2012_cleaning.R")
-source("./src/eda/dcr_anes_2016_cleaning.R")
-source("./src/eda/dcr_anes_2020_cleaning.R")
+source("./eda/dcr_anes_2012_cleaning.R")
+source("./eda/dcr_anes_2016_cleaning.R")
+source("./eda/dcr_anes_2020_cleaning.R")
   #* modularly load handy functions
 box::use(
   , mice[
@@ -34,7 +34,8 @@ data <- list()
 lasso <- list()
   #* load cleaned data
     #** 2012
-data[["anes_2012"]] <- read.csv("./data/clean/anes_2012_updated.csv")
+data[["anes_2012"]] <- read.csv("../data/clean/anes_2012_updated.csv") |>
+  as.data.table()
 data[["anes_2012"]] <- data[["anes_2012"]][
   , c(
     "wid", "rboff", "epast", "unpast", "edu", "income", "losejob"
@@ -42,8 +43,21 @@ data[["anes_2012"]] <- data[["anes_2012"]][
     , "whitediscrim", "raceResent", "pid", "female"
     )
 ]
+data[["anes_2012_female"]] <- data[["anes_2012"]][
+  female == 1,
+]
+data[["anes_2012_male"]] <- data[["anes_2012"]][
+  female == 0,
+]
+data[["anes_2012_republican"]] <- data[["anes_2012"]][
+  pid > 4,
+]
+data[["anes_2012_democrat"]] <- data[["anes_2012"]][
+  pid < 4,
+]
     #** 2016
-data[["anes_2016"]] <- read.csv("./data/clean/anes_2016_updated.csv")
+data[["anes_2016"]] <- read.csv("../data/clean/anes_2016_updated.csv") |>
+  as.data.table()
 data[["anes_2016"]] <- data[["anes_2016"]][
   , c(
     "wid", "rboff", "epast", "unpast", "edu", "income"
@@ -51,8 +65,21 @@ data[["anes_2016"]] <- data[["anes_2016"]][
     , "govbias", "inflwhite", "whitediscrim", "raceResent", "pid", "female"
   )
 ]
+data[["anes_2016_female"]] <- data[["anes_2016"]][
+  female == 1,
+]
+data[["anes_2016_male"]] <- data[["anes_2016"]][
+  female == 0,
+]
[…]

### 2023-07-07 src/manuscript.qmd
another draft'

@@
 Realistic Conflict Theory suggests that a racial identity and racial prejudice come as a response to a collective concern among members of a particular racial group about the loss of economic influence and status a group holds relative to other racial groups [@blumer_1958_psr;@bobo_hutchings_1996_asr].
-Some work in support of this perspective sees White racial identity as a response to concerns about increased job market competition through immigration [@pardos-prado_xena_2019].
-Other work finds support for this argument in lab experiments [@perez_et-al_2022_pp] and in surveys [@lindsay_2023_apr].
-In one of the most thorough treatments on the use of racial identity by Whites in politics, @jardina_2019_cup includes indicators of economic status and power in the primary model to predict White political identity.
-If this perspective holds, we might expect that in times of economic hardship, Whites may be most sensitive to economic threats as the collective and the individual see economics as most salient to them.
-
-Social identity theory, and its offshoots, would suggest that Whites expressing a racial identity are responding to political, cultural, and societal threats.
+Some evidence suggests that Whites living in areas with shrinking job opportunities respond to this with an increased desire to protect the interests of their racial confederates [@baccini_weymouth_2021_apsr].
+Others have experimental evidence supporting this perspective that White identity is motivated by material concerns [@perez_et-al_2022_pp].
+Further, @jardina_2019_cup who provides one of the most comprehensive treatments of conceptualizing, measuring, and examining the outcomes of White identity in politics includes indicators of economic status and power in the primary model.
+
+Social identity theory, and its offshoots, would suggest an alternative perspective: that Whites expressing a racial identity are responding to political, cultural, and societal threats.
 While these threats are a bit more hard to define, there are a number of examples of Whites expressing prejudice towards other racial groups in response to these sorts of challenges to their power.
@@
 For Whites, their racial identity has long been central to their behavior and the way in which they mobilize to shape legal and political institutions to maintain power [@rile-hayward_2013].
-
-We may also see that economic or political threat may appear differently for partisans and for female respondents.
-Recent evidence suggests that racial attitudes are guided by partisanship [@engelhardt_2021_bjps].
-Specifically, some evidence suggests that White Democrats are much more sensitive to their racial identity than White Republicans due to the higher amount of racial heterogeneity within the Democratic party.
-This may mean for White Democrats that they have more exposure to intergroup competition for power within the party (such as the choice to support Hillary Clinton versus Barack Obama) and may have higher White identification as a result.
-The second way that these threats may matter more or less for Whites is on the dimension of sex.
-Gender-based differences in economic influence and status remains quite high.
-We therefore may expect that female individuals are more sensitive to economic threats than their male counterparts.
-In addition to the ombnibus model, I also include two additional models where I split the samples along the dimension of partisanship (including leaners in each) and on the dimension of sex.
 
@@
 
+An additional question worth examining is that economic or political threat may appear differently for partisans and for female respondents.
+Recent evidence suggests that racial attitudes are guided by partisanship [@engelhardt_2021_bjps].
+Specifically, some evidence suggests that White Democrats are much more sensitive to their racial identity than White Republicans due to the higher amount of racial heterogeneity within the Democratic party.
+This may mean for White Democrats that they have more exposure to intergroup competition for power within the party (such as the choice to support Hillary Clinton versus Barack Obama) and may have higher White identification as a result.
+The second way that these threats may matter more or less for Whites is on the dimension of sex.
+Gender-based differences in economic influence and status remains quite high.
+We therefore may expect that female individuals are more sensitive to economic threats than their male counterparts.
+In addition to the ombnibus model, I also include two additional models where I split the samples along the dimension of partisanship (including leaners in each) and on the dimension of sex.
+
+```{r}
+#| label: fig-2012-2016-2020-female-male-lasso
+#| layout-ncol: 2
+#| fig-cap: Economic factors are weak predictors of white identity
+#| fig-subcap:
+#|   - Female, Economic
+#|   - Female, Political
+#|   - Male, Economic
+#|   - Female, Economicß
+# Graphs
+  #* Tidyfy posterior draws
+    #** 2012
+lasso[["2012_female_tidy"]] <- parameters(
+  lasso[["2012_female"]]
+  , conf.int = 0.95
+  , verbose = FALSE
+) |>
+  standardize_names(style = "broom") |>
+  data.frame() |>
+  mutate(model = "2012") |>
+  filter(!is.na(term))
+lasso[["2012_male_tidy"]] <- parameters(
+  lasso[["2012_male"]]
+  , conf.int = 0.95
+  , verbose = FALSE
+) |>
+  standardize_names(style = "broom") |>
+  data.frame() |>
+  mutate(model = "2012") |>
+  filter(!is.na(term))
+    #** 2016
+  lasso[["2016_female_tidy"]] <- parameters(
+  lasso[["2016_female"]]
+  , conf.int = 0.95
+  , verbose = FALSE
+) |>
+  standardize_names(style = "broom") |>
+  data.frame() |>
+  mutate(model = "2016") |>
+  filter(!is.na(term))
+lasso[["2016_male_tidy"]] <- parameters(
+  lasso[["2016_male"]]
+  , conf.int = 0.95
+  , verbose = FALSE
+) |>
+  standardize_names(style = "broom") |>
+  data.frame() |>
+  mutate(model = "2016") |>
+  filter(!is.na(term))
+    #** 2020
+lasso[["2020_female_tidy"]] <- parameters(
+  lasso[["2020_female"]]
+  , conf.int = 0.95
+  , verbose = FALSE
+) |>
+  standardize_names(style = "broom") |>
+  data.frame() |>
+  mutate(model = "2020") |>
+  filter(!is.na(term))
+lasso[["2020_male_tidy"]] <- parameters(
+  lasso[["2020_male"]]
+  , conf.int = 0.95
+  , verbose = FALSE
+) |>
+  standardize_names(style = "broom") |>
+  data.frame() |>
+  mutate(model = "2020") |>
+  filter(!is.na(term))
+
+lasso[['female_tidy']] <- rbind(
+  lasso[['2012_female_tidy']]
+  , lasso[['2016_female_tidy']]
+  , lasso[['2020_female_tidy']]
+) |> # append the three models' results together
+  mutate(
+    term = case_when(
+      term == 'b_rboff' ~ 'Retrospective better off'
+      , term == 'b_epast' ~ 'Economy better'
+      , term == 'b_unpast' ~ 'Unemployment better'
+      , term == 'b_edu' ~ 'Education'
+      , term == 'b_income' ~ 'Income'
+      , term == 'b_losejob' ~ 'Worried about losing job'
+      , term == 'b_ecfamily' ~ 'Worried about family finances'
+      , term == 'b_ecjob' ~ 'Know someone lost job'
+      , term == 'b_tradbreak' ~ 'New lifestyles breaking society'
+      , term == 'b_govbias' ~ 'Gov. biased against whites'
+      , term == 'b_inflwhite' ~ 'whites influence politics'
+      , term == 'b_whitediscrim' ~ 'whites are discriminated against'
+      , term == 'b_raceResent' ~ 'Racial resentment'
+      , term == 'b_pid' ~ 'Party ID'
+    ), # rename the parameters
+    term = factor(
+      term
+      , levels = c(
+        'Retrospective better off'
+        , 'Economy better'
+        , 'Unemployment better'
+        , 'Education'
+        , 'Income'
+        , 'Worried about losing job'
+        , 'Worried about family finances'
+        , 'Know someone lost job'
+        , 'New lifestyles breaking society'
+        , 'Gov. biased against whites'
+        , 'whites influence politics'
+        , 'whites are discriminated against'
+        , 'Racial resentment'
+        , 'Party ID'
+      )
+    ) # convert it to a factor variable
+  ) |> 
+  filter(!is.na(term)) # get rid of na rows
+lasso[['male_tidy']] <- rbind(
+  lasso[['2012_male_tidy']]
+  , lasso[['2016_male_tidy']]
+  , lasso[['2020_male_tidy']]
+) |> # append the three models' results together
+  mutate(
+    term = case_when(
+      term == 'b_rboff' ~ 'Retrospective better off'
+      , term == 'b_epast' ~ 'Economy better'
+      , term == 'b_unpast' ~ 'Unemployment better'
+      , term == 'b_edu' ~ 'Education'
+      , term == 'b_income' ~ 'Income'
+      , term == 'b_losejob' ~ 'Worried about losing job'
+      , term == 'b_ecfamily' ~ 'Worried about family finances'
+      , term == 'b_ecjob' ~ 'Know someone lost job'
+      , term == 'b_tradbreak' ~ 'New lifestyles breaking society'
+      , term == 'b_govbias' ~ 'Gov. biased against whites'
+      , term == 'b_inflwhite' ~ 'whites influence politics'
+      , term == 'b_whitediscrim' ~ 'whites are discriminated against'
+      , term == 'b_raceResent' ~ 'Racial resentment'
+      , term == 'b_pid' ~ 'Party ID'
+    ), # rename the parameters
+    term = factor(
+      term
+      , levels = c(
+        'Retrospective better off'
+        , 'Economy better'
+        , 'Unemployment better'
+        , 'Education'
+        , 'Income'
+        , 'Worried about losing job'
+        , 'Worried about family finances'
+        , 'Know someone lost job'
+        , 'New lifestyles breaking society'
+        , 'Gov. biased against whites'
+        , 'whites influence politics'
+        , 'whites are discriminated against'
+        , 'Racial resentment'
+        , 'Party ID'
+      )
+    ) # convert it to a factor variable
+  ) |> 
+  filter(!is.na(term)) # get rid of na rows
+    #** make a data.frame of results for the econ predictors
+lasso[["female_tidy_econ"]] <- lasso[["female_tidy"]] |>
+  filter(
+    term == "Retrospective better off" | term == "Economy better" | 
+    term == "Unemployment better" | term == "Education" | term == "Income" | 
+    term == "Worried about losing job" | term == "Wrroeid about family finances" |
+    term == "Know someone lost job"
+  )
+    #** make a data.frame of results for the political predictors
+lasso[["female_tidy_poli"]] <- lasso[["female_tidy"]] |>
+  filter(
+    term == "New lifestyles breaking society" | term == "Gov. biased against whites" |
+    term == "whites influence politics" | term == "whites are discriminated against" |
+    term == "Racial resentment"
+  )
+lasso[["male_tidy_econ"]] <- lasso[["male_tidy"]] |>
+  filter(
+    term == "Retrospective better off" | term == "Economy better" | 
+    term == "Unemployment better" | term == "Education" | term == "Income" | 
+    term == "Worried about losing job" | term == "Wrroeid about family finances" |
+    term == "Know someone lost job"
+  )
+    #** make a data.frame of results for the political predictors
+lasso[["male_tidy_poli"]] <- lasso[["male_tidy"]] |>
+  filter(
+    term == "New lifestyles breaking society" | term == "Gov. biased against whites" |
+    term == "whites influence politics" | term == "whites are discriminated against" |
+    term == "Racial resentment"
+  )
+female_econ <- ggplot(
+  data = lasso[['female_tidy_econ']]
+  , aes(
+    y = term
+    , x = estimate
+    , xmin = conf.low
+    , xmax = conf.high
+    , shape = factor(model)
+  )
+) +
+  geom_pointinterval(
+    position = position_dodge(width = 0.6)
+    , size = 4
+) +
+  geom_vline(
+    xintercept = 0
+    , linetype = 2
+) +
+  scale_y_discrete(limits = rev) +
+  theme_minimal() +
+  labs(
+    shape = 'Model'
+    , x = 'Estimate'
+    , y = ''
+    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
+) + 
+  theme(text=element_text(size=12, family = 'sans'))
+  #* plot for political predictors
+female_poli <- ggplot(
+  data = lasso[['female_tidy_poli']]
+  , aes(
+    y = term
+    , x = estimate
+    , xmin = conf.low
+    , xmax = conf.high
+    , shape = factor(model)
+  )
+) +
+  geom_pointinterval(
+    position = position_dodge(width = 0.6)
+    , size = 4
+) +
+  geom_vline(
+    xintercept = 0
+    , linetype = 2
+) +
+  scale_y_discrete(limits = rev) +
+  theme_minimal() +
+  labs(
+    shape = 'Model'
+    , x = 'Estimate'
+    , y = ''
+    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
+) + 
+  theme(text=element_text(size=12, family = 'sans'))
+male_econ <- ggplot(
+  data = lasso[['male_tidy_econ']]
+  , aes(
+    y = term
+    , x = estimate
+    , xmin = conf.low
+    , xmax = conf.high
+    , shape = factor(model)
+  )
+) +
+  geom_pointinterval(
+    position = position_dodge(width = 0.6)
+    , size = 4
+) +
+  geom_vline(
+    xintercept = 0
+    , linetype = 2
+) +
+  scale_y_discrete(limits = rev) +
+  theme_minimal() +
+  labs(
+    shape = 'Model'
+    , x = 'Estimate'
+    , y = ''
+    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
+) + 
+  theme(text=element_text(size=12, family = 'sans'))
+  #* plot for political predictors
+male_poli <- ggplot(
+  data = lasso[['male_tidy_poli']]
+  , aes(
+    y = term
+    , x = estimate
+    , xmin = conf.low
+    , xmax = conf.high
+    , shape = factor(model)
+  )
+) +
+  geom_pointinterval(
+    position = position_dodge(width = 0.6)
+    , size = 4
+) +
+  geom_vline(
+    xintercept = 0
+    , linetype = 2
+) +
+  scale_y_discrete(limits = rev) +
+  theme_minimal() +
+  labs(
+    shape = 'Model'
+    , x = 'Estimate'
+    , y = ''
+    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
+) + 
+  theme(text=element_text(size=12, family = 'sans'))
+
+female_econ
+female_poli
+male_econ
+male_poli
+```
+
+```{r}
+#| label: fig-2012-2016-2020-pid-lasso
+#| layout-ncol: 2
+#| fig-cap: Economic factors are weak predictors of white identity
+#| fig-subcap:
+#|   - Republican, Economic
+#|   - Republican, Political
+#|   - Democrat, Economic
+#|   - Democrat, Economic
+# Graphs
+  #* Tidyfy posterior draws
+    #** 2012
+lasso[["2012_republican_tidy"]] <- parameters(
+  lasso[["2012_republican"]]
+  , conf.int = 0.95
+  , verbose = FALSE
+) |>
+  standardize_names(style = "broom") |>
+  data.frame() |>
+  mutate(model = "2012") |>
+  filter(!is.na(term))
+lasso[["2012_democrat_tidy"]] <- parameters(
+  lasso[["2012_democrat"]]
+  , conf.int = 0.95
+  , verbose = FALSE
+) |>
+  standardize_names(style = "broom") |>
+  data.frame() |>
+  mutate(model = "2012") |>
+  filter(!is.na(term))
+    #** 2016
+  lasso[["2016_republican_tidy"]] <- parameters(
+  lasso[["2016_republican"]]
+  , conf.int = 0.95
+  , verbose = FALSE
+) |>
+  standardize_names(style = "broom") |>
+  data.frame() |>
+  mutate(model = "2016") |>
+  filter(!is.na(term))
+lasso[["2016_democrat_tidy"]] <- parameters(
+  lasso[["2016_democrat"]]
+  , conf.int = 0.95
+  , verbose = FALSE
+) |>
+  standardize_names(style = "broom") |>
+  data.frame() |>
+  mutate(model = "2016") |>
+  filter(!is.na(term))
+    #** 2020
+lasso[["2020_republican_tidy"]] <- parameters(
+  lasso[["2020_republican"]]
+  , conf.int = 0.95
+  , verbose = FALSE
+) |>
+  standardize_names(style = "broom") |>
+  data.frame() |>
+  mutate(model = "2020") |>
+  filter(!is.na(term))
+lasso[["2020_democrat_tidy"]] <- parameters(
+  lasso[["2020_democrat"]]
+  , conf.int = 0.95
+  , verbose = FALSE
+) |>
+  standardize_names(style = "broom") |>
+  data.frame() |>
+  mutate(model = "2020") |>
+  filter(!is.na(term))
+
+lasso[['republican_tidy']] <- rbind(
+  lasso[['2012_republican_tidy']]
+  , lasso[['2016_republican_tidy']]
+  , lasso[['2020_republican_tidy']]
+) |> # append the three models' results together
+  mutate(
+    term = case_when(
+      term == 'b_rboff' ~ 'Retrospective better off'
+      , term == 'b_epast' ~ 'Economy better'
+      , term == 'b_unpast' ~ 'Unemployment better'
+      , term == 'b_edu' ~ 'Education'
+      , term == 'b_income' ~ 'Income'
+      , term == 'b_losejob' ~ 'Worried about losing job'
+      , term == 'b_ecfamily' ~ 'Worried about family finances'
+      , term == 'b_ecjob' ~ 'Know someone lost job'
+      , term == 'b_tradbreak' ~ 'New lifestyles breaking society'
+      , term == 'b_govbias' ~ 'Gov. biased against whites'
+      , term == 'b_inflwhite' ~ 'whites influence politics'
+      , term == 'b_whitediscrim' ~ 'whites are discriminated against'
+      , term == 'b_raceResent' ~ 'Racial resentment'
+      , term == 'b_female' ~ 'Female'
+    ), # rename the parameters
+    term = factor(
+      term
+      , levels = c(
+        'Retrospective better off'
+        , 'Economy better'
+        , 'Unemployment better'
+        , 'Education'
+        , 'Income'
+        , 'Worried about losing job'
+        , 'Worried about family finances'
+        , 'Know someone lost job'
+        , 'New lifestyles breaking society'
+        , 'Gov. biased against whites'
+        , 'whites influence politics'
+        , 'whites are discriminated against'
+        , 'Racial resentment'
+        , 'Female'
+      )
+    ) # convert it to a factor variable
+  ) |> 
+  filter(!is.na(term)) # get rid of na rows
+lasso[['democrat_tidy']] <- rbind(
+  lasso[['2012_democrat_tidy']]
+  , lasso[['2016_democrat_tidy']]
+  , lasso[['2020_democrat_tidy']]
+) |> # append the three models' results together
+  mutate(
+    term = case_when(
+      term == 'b_rboff' ~ 'Retrospective better off'
+      , term == 'b_epast' ~ 'Economy better'
+      , term == 'b_unpast' ~ 'Unemployment better'
+      , term == 'b_edu' ~ 'Education'
+      , term == 'b_income' ~ 'Income'
+      , term == 'b_losejob' ~ 'Worried about losing job'
+      , term == 'b_ecfamily' ~ 'Worried about family finances'
+      , term == 'b_ecjob' ~ 'Know someone lost job'
+      , term == 'b_tradbreak' ~ 'New lifestyles breaking society'
+      , term == 'b_govbias' ~ 'Gov. biased against whites'
+      , term == 'b_inflwhite' ~ 'whites influence politics'
+      , term == 'b_whitediscrim' ~ 'whites are discriminated against'
+      , term == 'b_raceResent' ~ 'Racial resentment'
+      , term == 'b_female' ~ 'Female'
+    ), # rename the parameters
+    term = factor(
+      term
+      , levels = c(
+        'Retrospective better off'
+        , 'Economy better'
+        , 'Unemployment better'
+        , 'Education'
+        , 'Income'
+        , 'Worried about losing job'
+        , 'Worried about family finances'
+        , 'Know someone lost job'
+        , 'New lifestyles breaking society'
+        , 'Gov. biased against whites'
+        , 'whites influence politics'
+        , 'whites are discriminated against'
+        , 'Racial resentment'
+        , 'Female'
+      )
+    ) # convert it to a factor variable
+  ) |> 
+  filter(!is.na(term)) # get rid of na rows
+    #** make a data.frame of results for the econ predictors
+lasso[["republican_tidy_econ"]] <- lasso[["republican_tidy"]] |>
+  filter(
+    term == "Retrospective better off" | term == "Economy better" | 
+    term == "Unemployment better" | term == "Education" | term == "Income" | 
+    term == "Worried about losing job" | term == "Wrroeid about family finances" |
+    term == "Know someone lost job"
+  )
+    #** make a data.frame of results for the political predictors
+lasso[["republican_tidy_poli"]] <- lasso[["republican_tidy"]] |>
+  filter(
+    term == "New lifestyles breaking society" | term == "Gov. biased against whites" |
+    term == "whites influence politics" | term == "whites are discriminated against" |
+    term == "Racial resentment"
+  )
+lasso[["democrat_tidy_econ"]] <- lasso[["democrat_tidy"]] |>
+  filter(
+    term == "Retrospective better off" | term == "Economy better" | 
+    term == "Unemployment better" | term == "Education" | term == "Income" | 
+    term == "Worried about losing job" | term == "Wrroeid about family finances" |
+    term == "Know someone lost job"
+  )
+    #** make a data.frame of results for the political predictors
+lasso[["democrat_tidy_poli"]] <- lasso[["democrat_tidy"]] |>
+  filter(
+    term == "New lifestyles breaking society" | term == "Gov. biased against whites" |
+    term == "whites influence politics" | term == "whites are discriminated against" |
+    term == "Racial resentment"
+  )
+republican_econ <- ggplot(
+  data = lasso[['republican_tidy_econ']]
+  , aes(
+    y = term
+    , x = estimate
+    , xmin = conf.low
+    , xmax = conf.high
+    , shape = factor(model)
+  )
+) +
+  geom_pointinterval(
+    position = position_dodge(width = 0.6)
+    , size = 4
+) +
+  geom_vline(
+    xintercept = 0
+    , linetype = 2
+) +
+  scale_y_discrete(limits = rev) +
+  theme_minimal() +
+  labs(
+    shape = 'Model'
+    , x = 'Estimate'
+    , y = ''
+    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
+) + 
+  theme(text=element_text(size=12, family = 'sans'))
+  #* plot for political predictors
+republican_poli <- ggplot(
+  data = lasso[['republican_tidy_poli']]
+  , aes(
+    y = term
+    , x = estimate
+    , xmin = conf.low
+    , xmax = conf.high
+    , shape = factor(model)
+  )
+) +
+  geom_pointinterval(
+    position = position_dodge(width = 0.6)
+    , size = 4
+) +
+  geom_vline(
+    xintercept = 0
+    , linetype = 2
+) +
+  scale_y_discrete(limits = rev) +
+  theme_minimal() +
+  labs(
+    shape = 'Model'
+    , x = 'Estimate'
+    , y = ''
+    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
+) + 
+  theme(text=element_text(size=12, family = 'sans'))
+democrat_econ <- ggplot(
+  data = lasso[['democrat_tidy_econ']]
+  , aes(
+    y = term
+    , x = estimate
+    , xmin = conf.low
+    , xmax = conf.high
+    , shape = factor(model)
+  )
+) +
+  geom_pointinterval(
+    position = position_dodge(width = 0.6)
+    , size = 4
+) +
+  geom_vline(
+    xintercept = 0
+    , linetype = 2
+) +
+  scale_y_discrete(limits = rev) +
+  theme_minimal() +
+  labs(
+    shape = 'Model'
+    , x = 'Estimate'
+    , y = ''
+    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
+) + 
+  theme(text=element_text(size=12, family = 'sans'))
+  #* plot for political predictors
+democrat_poli <- ggplot(
+  data = lasso[['democrat_tidy_poli']]
+  , aes(
+    y = term
+    , x = estimate
+    , xmin = conf.low
+    , xmax = conf.high
+    , shape = factor(model)
+  )
+) +
+  geom_pointinterval(
+    position = position_dodge(width = 0.6)
+    , size = 4
+) +
+  geom_vline(
+    xintercept = 0
+    , linetype = 2
+) +
+  scale_y_discrete(limits = rev) +
+  theme_minimal() +
+  labs(
+    shape = 'Model'
+    , x = 'Estimate'
+    , y = ''
+    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
+) + 
+  theme(text=element_text(size=12, family = 'sans'))
+
+republican_econ
+republican_poli
+democrat_econ
+democrat_poli
+```
+
+We see in both @fig-2012-2016-2020-female-male-lasso and @fig-2012-2016-2020-pid-lasso that indicators of economic and political threat do not have much substantial variation between male and female respondents and Republican and Democrat respondents.
+The full table of results for the model are included in @tbl-2012-gender-lasso, @tbl-2016-gender-lasso, @tbl-2020-gender-lasso, @tbl-2012-pid-lasso, @tbl-2016-pid-lasso, and @tbl-2020-pid-lasso in the Supplimentary Materials.
+
 # Discussion and conclusions
@@
 biased against whites', 'b_inflwhite' = 'whites influence politics', 'b_whitediscrim' = 'whites are discriminated against', 'b_raceResent' = 'Racial resentment', 'b_pid' = 'Party ID', 'b_female' = 'Female', 'b_Intercept[1]' = 'Threshold 1', 'b_Intercept[2]' = 'Threshold 2', 'b_Intercept[3]' = 'Threshold 3', 'b_Intercept[4]' = 'Threshold 4')
+
+cm_gender <- c('b_rboff' = 'Retrospective better off', 'b_epast' = 'Economy better', 'b_unpast' = 'Unemployment better', 'b_edu' = 'Education', 'b_income' = 'Income', 'b_losejob' = 'Worried about losing job', 'b_ecfamily' = 'Worried about family finances', 'b_ecjob' = 'Know someone lost job', 'b_tradbreak' = 'New lifestyles breaking society', 'b_govbias' = 'Gov.
+biased against whites', 'b_inflwhite' = 'whites influence politics', 'b_whitediscrim' = 'whites are discriminated against', 'b_raceResent' = 'Racial resentment', 'b_pid' = 'Party ID', 'b_Intercept[1]' = 'Threshold 1', 'b_Intercept[2]' = 'Threshold 2', 'b_Intercept[3]' = 'Threshold 3', 'b_Intercept[4]' = 'Threshold 4')
+
+cm_pid <- c('b_rboff' = 'Retrospective better off', 'b_epast' = 'Economy better', 'b_unpast' = 'Unemployment better', 'b_edu' = 'Education', 'b_income' = 'Income', 'b_losejob' = 'Worried about losing job', 'b_ecfamily' = 'Worried about family finances', 'b_ecjob' = 'Know someone lost job', 'b_tradbreak' = 'New lifestyles breaking society', 'b_govbias' = 'Gov.
+biased against whites', 'b_inflwhite' = 'whites influence politics', 'b_whitediscrim' = 'whites are discriminated against', 'b_raceResent' = 'Racial resentment', 'b_female' = 'Female', 'b_Intercept[1]' = 'Threshold 1', 'b_Intercept[2]' = 'Threshold 2', 'b_Intercept[3]' = 'Threshold 3', 'b_Intercept[4]' = 'Threshold 4')
 ```
@@
 model_2012 <- list(
-  '2012 LWD' = lasso[["2012_simple"]],
-  '2012' = lasso[["2012_ordinal"]]
+  "2012 LWD" = lasso[["2012_simple"]]
+  , "2012" = lasso[["2012_ordinal"]]
 )
@@
 
-modelsummary(model_2012, coef_map = cm, statistic = 'conf.int', gof_map = NA, add_rows = rows, notes = list('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2012 American National Election Study.'))
-
+modelsummary(model_2012, coef_map = cm, statistic = 'conf.int', gof_map = NA, add_rows = rows, notes = c('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2012 American National Election Study.'))
+
+```
+
+```{r}
+#| label: tbl-2012-gender-lasso
+#| tbl-cap: Predictors of Whit Identity in 2012 conditional on gender
+
+model_2012_gender <- list(
+  "2012 Female" = lasso[["2012_female"]]
+  , "2012 Male" = lasso[["2012_male"]]
+)
+rows <- tribble(~term, ~`2012 Female`, ~`2012 Male`, 'Economic', '', '', 'Non-material', '', '', 'Demographics', '', '', 'Thresholds', '', '')
+attr(rows, 'position') <- c(1, 18, 27, 30)
+
+modelsummary(model_2012_gender, coef_map = cm_gender, statistic = "conf.int", gof_map = NA, add_rows = rows, notes = c('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2012 American National Election Study.'))
+```
+
+```{r}
+#| label: tbl-2012-pid-lasso
+#| tbl-cap: Predictors of Whit Identity in 2012 conditional on partisanship
+
+model_2012_pid <- list(
+  "2012 Republican" = lasso[["2012_republican"]]
+  , "2012 Democrat" = lasso[["2012_democrat"]]
+)
+rows <- tribble(~term, ~`2012 Republican`, ~`2012 Democrat`, 'Economic', '', '', 'Non-material', '', '', 'Demographics', '', '', 'Thresholds', '', '')
+attr(rows, 'position') <- c(1, 18, 27, 30)
+
+modelsummary(model_2012_pid, coef_map = cm_pid, statistic = "conf.int", gof_map = NA, add_rows = rows, notes = c('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2012 American National Election Study.'))
 ```
@@
 model_2016 <- list(
-  '2016 LWD' = lasso[["2016_simple"]],
-  '2016' = lasso[["2016_ordinal"]]
+  "2016 LWD" = lasso[["2016_simple"]]
+  , "2016" = lasso[["2016_ordinal"]]
 )
@@
 
-modelsummary(model_2016, coef_map = cm, statistic = 'conf.int', gof_map = NA, add_rows = rows, notes = list('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2016 American National Election Study.'))
-
-```
-
+modelsummary(model_2016, coef_map = cm, statistic = 'conf.int', gof_map = NA, add_rows = rows, notes = c('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2016 American National Election Study.'))
+```
+
+```{r}
+#| label: tbl-2016-gender-lasso
+#| tbl-cap: Predictors of Whit Identity in 2016 conditional on gender
+
+model_2016_gender <- list(
+  "2016 Female" = lasso[["2016_female"]]
+  , "2016 Male" = lasso[["2016_male"]]
+)
+rows <- tribble(~term, ~`2016 Female`, ~`2016 Male`, 'Economic', '', '', 'Non-material', '', '', 'Demographics', '', '', 'Thresholds', '', '')
+attr(rows, 'position') <- c(1, 18, 27, 30)
+
+modelsummary(model_2016_gender, coef_map = cm_gender, statistic = "conf.int", gof_map = NA, add_rows = rows, notes = c('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2016 American National Election Study.'))
+```
+
+```{r}
+#| label: tbl-2016-pid-lasso
+#| tbl-cap: Predictors of Whit Identity in 2016 conditional on partisanship
+
+model_2016_pid <- list(
+  "2016 Republican" = lasso[["2016_republican"]]
+  , "2016 Democrat" = lasso[["2016_democrat"]]
+)
+rows <- tribble(~term, ~`2016 Republican`, ~`2016 Democrat`, 'Economic', '', '', 'Non-material', '', '', 'Demographics', '', '', 'Thresholds', '', '')
+attr(rows, 'position') <- c(1, 18, 27, 30)
+
+modelsummary(model_2016_pid, coef_map = cm_pid, statistic = "conf.int", gof_map = NA, add_rows = rows, notes = c('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2016 American National Election Study.'))
+```
 ```{r}
@@
 model_2020 <- list(
-  '2020 LWD' = lasso[["2020_simple"]],
-  '2020' = lasso[["2020_ordinal"]]
+  "2020 LWD" = lasso[["2020_simple"]]
+  , "2020" = lasso[["2020_ordinal"]]
 )
@@
 
-modelsummary(model_2020, coef_map = cm, statistic = 'conf.int', gof_map = NA, add_rows = rows, notes = list('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2020 American National Election Study.'))
-```
+modelsummary(model_2020, coef_map = cm, statistic = 'conf.int', gof_map = NA, add_rows = rows, notes = c('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2020 American National Election Study.'))
+```
+
+```{r}
+#| label: tbl-2020-gender-lasso
+#| tbl-cap: Predictors of Whit Identity in 2020 conditional on gender
+
+model_2020_gender <- list(
+  "2020 Female" = lasso[["2020_female"]]
+  , "2020 Male" = lasso[["2020_male"]]
+)
+rows <- tribble(~term, ~`2020 Female`, ~`2020 Male`, 'Economic', '', '', 'Non-material', '', '', 'Demographics', '', '', 'Thresholds', '', '')
+attr(rows, 'position') <- c(1, 18, 27, 30)
+
+modelsummary(model_2020_gender, coef_map = cm_gender, statistic = "conf.int", gof_map = NA, add_rows = rows, notes = c('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2020 American National Election Study.'))
+```
+
+```{r}
+#| label: tbl-2020-pid-lasso
+#| tbl-cap: Predictors of Whit Identity in 2012 conditional on partisanship
+
+model_2020_pid <- list(
+  "2020 Republican" = lasso[["2020_republican"]]
+  , "2020 Democrat" = lasso[["2020_democrat"]]
+)
+rows <- tribble(~term, ~`2020 Republican`, ~`2020 Democrat`, 'Economic', '', '', 'Non-material', '', '', 'Demographics', '', '', 'Thresholds', '', '')
+attr(rows, 'position') <- c(1, 18, 27, 30)
+
+modelsummary(model_2020_pid, coef_map = cm_pid, statistic = "conf.int", gof_map = NA, add_rows = rows, notes = c('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2020 American National Election Study.'))
+```

### 2023-07-07 src/manuscript.qmd
send off a draft of this

@@
 
-Are White identifiers those who are percieve economic threats or is it a backlash to a loss of influence over politics?
+Do economic or political threats motivate the expression of a racial identity among Whites in politics?
 The existing work offers an unclear answer to this question.
@@
 
-A common source of data and of measures for researchers in the United States is the American National Election Study.
-I use a penalized regression using data from the 2012, 2016, and 2020 American National Election Studies to examine the degree to which individual measures of economic and political influence explain an individual's uptake of White identity as well as to examine the degree to which changes in the political context may increase or decrease the explanatory power of these measures.
-I find that proxies reporting concerns about Whites' political influence are stronger explantory factors than proxies of economic influence in the post-Trump era.
+As the American National Election Study is a long-running survey of Americans', it provides some of the most consistent data on common political attitudes and behaviors.
+Using data from the 2012, 2016, and 2020 administrations of the ANES, I fit an ombnibus model with penalized regression to examine the degree to which individual measures of economic and political influence explain an individual's uptake of White identity as well as to examine the degree to which changes in the political context may increase or decrease the explanatory power of these measures.
+From this test, I find that proxies reporting concerns about Whites' political influence are stronger explantory factors than proxies of economic influence in the post-Trump era.
 
@@
 Social identity theory, and its offshoots, would suggest an alternative perspective: that Whites expressing a racial identity are responding to political, cultural, and societal threats.
-While these threats are a bit more hard to define, there are a number of examples of Whites expressing prejudice towards other racial groups in response to these sorts of challenges to their power.
-@tesler_2016 demonstrates that Whites heavily racialized American politics in response to Barack Obama's ascention to the White House.
-Other work demonstrates that the saliency and reprocussions of aligning with competing racial groups drove non-White political candidates to distance themselves from their racial confederates in an attempt to appease Whites [@stephens-dougan_2020].
-Some pundits on Fox News suggest that Whites are in direct competition for jobs with immigrants from South and Central America.
-Despite these claims, some evidence suggests that Whites' hostility directed toward non-White immigrant groups depends on the degree to which they assimilate [@green_staerkle_2013; see also @hainmueller_hopkins_2014].
-In terms of attitudes oriented toward the in-group, @dawson_1994_pup argues that Black Americans hold a strong racial identity that supercedes their class-based identity due to the persistent focus on Black individual's race in their personal and political life.
-For Whites, their racial identity has long been central to their behavior and the way in which they mobilize to shape legal and political institutions to maintain power [@rile-hayward_2013].
+While these threats are a bit more hard to define, there are a number of examples that demonstrate Whites are responsive to cultural or political changes that present as a threat.
+For example, @mutz_status_2018 finds little support for the claim that the 2016 Presidential Election was a response to the economic effects of a globalizing economy and instead argues that it was a response to concerns about the changes brought about by globalization and the racial-diversity it promotes.
+Others demonstrate that the behavioral manifestations of a White constituency in the United States was a reflection of the saliency of race in politics [@agadjanian_lacy_2021].
+A final example demonstrated that exposure to a story about discrimination about a White person lead individuals concerned with the racial hierarchy to report that Whites faced similar amounts of institutional racism as Blacks [@marshburn_et-al_2022_gpir].
+Others have made similar arguments through historical lenses by arguing that while race and class are highly correlated, the saliency of race in society and in our institutions continue to make it a salient group by which we structure our thinking about who has status and power [@helms_carter_1990;@dawson_1994_pup;@rile-hayward_2013]
 
@@
 
-Identity is largely an elusive concept to operationalize.
-Though there are implicit measures of White identity [see @knowles_peng_2005_jpsp], the common and measure of white identity is a response to the question, "How important is being White to your identity?" This measure captures a significant feature of an identity: the centrality or the importance of that identity to an individual.
+Identity is largely an elusive concept to operationalize [see @leach_et-al_2008].
+Though there are implicit measures of White identity [see @knowles_peng_2005_jpsp], the common and measure of white identity is a response to the question, "How important is being White to your identity?" This measure captures a significant feature of an identity: the centrality or the importance of that identity to an individual [@jardina_2019_cup].
 Though a rather rough measure of something as complex as identity, it has been shown to produce similar results to more complex measures of white identity and holds a significant amount of face validity [@jardina_2019_cup].
@@
 
-The primary contribution of this paper is to examine which of a number of competing measures of different types of threat may predict white identity for individuals.
+The primary contribution of this paper is to examine which, of several competing measures, of different types of threat may predict white identity for individuals.
 As this requires an ombnibus test, I elect to use a penalized regression.
 I provide a more detailed discussion of the reasons why I use an estimator with an L2 regularization parameter, often refered to as a LASSO model in the supplementary materials.
-For a brief comment as to why I use this approach is because approach is designed to take models with a large number of predictors and to shrink the coefficients toward zero.
-In practice, this often means that predictors that has relatively fewer explanatory power have much smaller coefficients than the variables that do provide significant predictive power.
+In practice, the model applies a penalization term so predictors that have relatively fewer explanatory power have much smaller coefficients than the variables that do provide significant predictive power.
 
@@
 #| label: tbl-2012-lasso-simple
-#| tbl-cap: Predictors of white Identity in 2012
+#| tbl-cap: Predictors of White Identity in 2012
 model_2012 <- list(
@@
 #| label: tbl-2012-gender-lasso
-#| tbl-cap: Predictors of Whit Identity in 2012 conditional on gender
+#| tbl-cap: Predictors of White Identity in 2012 conditional on gender
 
@@
 #| label: tbl-2012-pid-lasso
-#| tbl-cap: Predictors of Whit Identity in 2012 conditional on partisanship
+#| tbl-cap: Predictors of White Identity in 2012 conditional on partisanship
 
@@
 #| label: tbl-2016-lasso-simple
-#| tbl-cap: Predictors of white identity in 2016
+#| tbl-cap: Predictors of White Identity in 2016
 model_2016 <- list(
@@
 #| label: tbl-2016-gender-lasso
-#| tbl-cap: Predictors of Whit Identity in 2016 conditional on gender
+#| tbl-cap: Predictors of White Identity in 2016 conditional on gender
 
@@
 #| label: tbl-2016-pid-lasso
-#| tbl-cap: Predictors of Whit Identity in 2016 conditional on partisanship
+#| tbl-cap: Predictors of White Identity in 2016 conditional on partisanship
 
@@
 #| label: tbl-2020-lasso-simple
-#| tbl-cap: Predictors of white identity in 2020
+#| tbl-cap: Predictors of White identity in 2020
 model_2020 <- list(
@@
 #| label: tbl-2020-gender-lasso
-#| tbl-cap: Predictors of Whit Identity in 2020 conditional on gender
+#| tbl-cap: Predictors of White Identity in 2020 conditional on gender
 
@@
 #| label: tbl-2020-pid-lasso
-#| tbl-cap: Predictors of Whit Identity in 2012 conditional on partisanship
+#| tbl-cap: Predictors of White Identity in 2012 conditional on partisanship
 

### 2023-07-19 src/manuscript.qmd
slight tweaks to the draft

@@
 thanks: |
-    Word count: 3575.
+    Word count (including references): 3655.
 I would like to thank Jennifer Wolak for her feedback on the many drafts of this manuscript.
@@
 The existing work offers an unclear answer to this question.
-This note attempts to provide an ombnibus test to determine not only which of these questions has more empirical support but also which among common proxies of economic and political power in the United States appear to be reliable predictors of a White individual's expression of racial identity.
-
-As the American National Election Study is a long-running survey of Americans', it provides some of the most consistent data on common political attitudes and behaviors.
-Using data from the 2012, 2016, and 2020 administrations of the ANES, I fit an ombnibus model with penalized regression to examine the degree to which individual measures of economic and political influence explain an individual's uptake of White identity as well as to examine the degree to which changes in the political context may increase or decrease the explanatory power of these measures.
-From this test, I find that proxies reporting concerns about Whites' political influence are stronger explantory factors than proxies of economic influence in the post-Trump era.
+This note provides an omnibus test to determine not only which of these questions has more empirical support but also which among common proxies of economic and political power in the United States appear to be reliable predictors of a White individual's expression of racial identity.
+
+As the American National Election Study is a long-running survey of Americans', it provides some of the most consistent measures on common political attitudes and behaviors.
+Using data from the 2012, 2016, and 2020 administrations of the ANES, I fit an omnibus model with penalized regression to examine the degree to which individual measures of economic and political influence explain an individual's uptake of White identity as well as to examine the degree to which changes in the political context may increase or decrease the explanatory power of these measures.
+From this test, I find that proxies reporting concerns about Whites' political influence are stronger explanatory factors than proxies of economic influence in the post-Trump era.
 
@@
 While these threats are a bit more hard to define, there are a number of examples that demonstrate Whites are responsive to cultural or political changes that present as a threat.
-For example, @mutz_status_2018 finds little support for the claim that the 2016 Presidential Election was a response to the economic effects of a globalizing economy and instead argues that it was a response to concerns about the changes brought about by globalization and the racial-diversity it promotes.
+For example, @mutz_status_2018 finds little support for the claim that the 2016 Presidential Election -- which support for Trump is associated with an expression of racial identity among Whites [@jardina_2020_pb] -- was a response to the economic effects of a globalizing economy and instead argues that it was a response to concerns about the changes brought about by globalization and the racial-diversity it promotes.
 Others demonstrate that the behavioral manifestations of a White constituency in the United States was a reflection of the saliency of race in politics [@agadjanian_lacy_2021].
 A final example demonstrated that exposure to a story about discrimination about a White person lead individuals concerned with the racial hierarchy to report that Whites faced similar amounts of institutional racism as Blacks [@marshburn_et-al_2022_gpir].
-Others have made similar arguments through historical lenses by arguing that while race and class are highly correlated, the saliency of race in society and in our institutions continue to make it a salient group by which we structure our thinking about who has status and power [@helms_carter_1990;@dawson_1994_pup;@rile-hayward_2013]
+Others have made similar arguments through historical lenses by arguing that while race and class are highly correlated, the saliency of race in society and in our institutions continue to make it a salient group by which we structure our thinking about who has status and power [@helms_carter_1990;@dawson_1994_pup;@rile-hayward_2013].
+This perspective fits with many arguments that the United States is not as class-conscious as they are race-conscious [@zoltan_l_hajnal_dangerously_2020].
 
@@
 
+I am using a popular conceptualization of racial identity among Whites: that expression of it reflects homophily toward's one's own racial group [@ellemers_et-al_2002_arp;@jardina_2019_cup].
+As identity conceptualized this way may predict negative intergroup attitudes, it reflects a desire to find ways to find similarities rather than differences between one's self and other in-group members [@ellemers_et-al_2002_arp].
 Identity is largely an elusive concept to operationalize [see @leach_et-al_2008].
@@
 The second way that these threats may matter more or less for Whites is on the dimension of sex.
-Gender-based differences in economic influence and status remains quite high.
+Sex-based differences in economic influence and status remains quite high.
 We therefore may expect that female individuals are more sensitive to economic threats than their male counterparts.
@@
 Even in a time where the economy is recovering from a collapse, more proxies of political threat seemed to predict White identity.
-When a White president well-known for his penchance for racial dog-whistles takes office, we see support for the idea that White grievance politics was at a high -- high those who felt that political institutions and that the direction of society was turning away from the racial status quo were those that reported their racial identity as more central to who they are.
+When a White president well-known for his penchance for racial dog-whistles takes office, we see support for the idea that White grievance politics was at a high as evidenced by those who felt that political institutions and that the direction of society was turning away from the racial status quo were those that reported their racial identity as more central to who they are.
 
 There are a few important points that are illustrated here.
-First, suggestions that White political identity is something occuring for those who are most economically threatened appear to have weak theoretical and empirical backing when directly pitted against other forms of threat.
-Second, that measures of cultural or societal threat are important variables to include if one is using white identity as a variable in a model.
+First, suggestions that White political identity is something occuring for those who are most economically threatened appear to have weak theoretical and empirical backing when directly pitted against symbolic forms of threat.
+Second, when thinking about what is driving racial identity among Whites, evidence from this omnibus test suggests that it is more symbolic forms of threat as opposed to material.
 

### 2023-09-21 src/manuscript.qmd
updated citations and a little bit of the framing

@@
 thanks: |
-    Word count (including references): 3655.
+    Word count (including references): 3898.
 I would like to thank Jennifer Wolak for her feedback on the many drafts of this manuscript.
@@
 date: today
-bibliography: "../assets/wid.bib"
+bibliography: "../assets/references.bib"
 format:
@@
     latex-output-dir: ../out/
-    filters:
-    - "_extensions/andrewheiss/wordcount/citeproc.lua"
-    - "_extensions/andrewheiss/wordcount/wordcount.lua"
     geometry:
@@
     fontsize: 12pt
-    appendix-style: none
+    appendix-style: default
     include-in-header:
@@
 
-Do economic or political threats motivate the expression of a racial identity among Whites in politics?
-The existing work offers an unclear answer to this question.
-This note provides an omnibus test to determine not only which of these questions has more empirical support but also which among common proxies of economic and political power in the United States appear to be reliable predictors of a White individual's expression of racial identity.
-
-As the American National Election Study is a long-running survey of Americans', it provides some of the most consistent measures on common political attitudes and behaviors.
-Using data from the 2012, 2016, and 2020 administrations of the ANES, I fit an omnibus model with penalized regression to examine the degree to which individual measures of economic and political influence explain an individual's uptake of White identity as well as to examine the degree to which changes in the political context may increase or decrease the explanatory power of these measures.
-From this test, I find that proxies reporting concerns about Whites' political influence are stronger explanatory factors than proxies of economic influence in the post-Trump era.
-
-# Identity and white political identity
+There is a growing literature suggesting that White Political Identity predicts a number of outcomes.
+For example evidence suggests that White Political Identity predicts support for Trump in the 2016 Presidential Election [@jardina_disavowing_2021].
+The existing literature is less clear about the predictors of White Political Identity, however.
+I advance an argument as to why we might consider the expression of a White Political Identity to be a reflection of concerns around cultural threat as opposed to economic threat.
+The primary contribution of the note is that I provide an omnibus test to tease out whether cultural or economic threats are more reliable predictors of a White individual's expression of racial identity in the United States.
+
+As the American National Election Study is a long-running survey of those residing in the United States, it provides some of the most consistent measures on common political attitudes and behaviors.
+Using data from the 2012, 2016, and 2020 administrations of the ANES, I fit an omnibus model with penalized regression to examine the degree to which additive scales and individual measures of economic and cultural influence explain an individual's uptake of White identity as well as to examine the degree to which changes in the social context may increase or decrease the explanatory power of these measures.
+From this test, I find that measures reporting concerns about Whites' cultural influence are stronger explanatory factors than proxies of economic influence in the post-Trump era.
+
+# Identity and White Political Identity
 
 One perspective on what explains racial identity among Whites in the United States are that they are responding to threats to their economic status and power.
-Realistic Conflict Theory suggests that a racial identity and racial prejudice come as a response to a collective concern among members of a particular racial group about the loss of economic influence and status a group holds relative to other racial groups [@blumer_1958_psr;@bobo_hutchings_1996_asr].
-Some evidence suggests that Whites living in areas with shrinking job opportunities respond to this with an increased desire to protect the interests of their racial confederates [@baccini_weymouth_2021_apsr].
-Others have experimental evidence supporting this perspective that White identity is motivated by material concerns [@perez_et-al_2022_pp].
-Further, @jardina_2019_cup who provides one of the most comprehensive treatments of conceptualizing, measuring, and examining the outcomes of White identity in politics includes indicators of economic status and power in the primary model.
+Realistic Conflict Theory suggests that a racial identity and racial prejudice come as a response to a collective concern among members of a particular racial group about the loss of economic influence and the status a group holds relative to other racial groups [@blumer_race_1958;@bobo_perceptions_1996].
+Some evidence suggests that Whites living in areas with shrinking job opportunities respond to this with an increased desire to protect the interests of their racial confederates [@baccini_gone_2021].
+Others have experimental evidence supporting this perspective that White Political Identity is motivated by material concerns [@perez_politics_2022].
+Further, @jardina_white_2019 who provides one of the most comprehensive treatments of conceptualizing, measuring, and examining the outcomes of White Political Identity includes indicators of economic status and power in the primary model.
 
@@
 While these threats are a bit more hard to define, there are a number of examples that demonstrate Whites are responsive to cultural or political changes that present as a threat.
-For example, @mutz_status_2018 finds little support for the claim that the 2016 Presidential Election -- which support for Trump is associated with an expression of racial identity among Whites [@jardina_2020_pb] -- was a response to the economic effects of a globalizing economy and instead argues that it was a response to concerns about the changes brought about by globalization and the racial-diversity it promotes.
-Others demonstrate that the behavioral manifestations of a White constituency in the United States was a reflection of the saliency of race in politics [@agadjanian_lacy_2021].
-A final example demonstrated that exposure to a story about discrimination about a White person lead individuals concerned with the racial hierarchy to report that Whites faced similar amounts of institutional racism as Blacks [@marshburn_et-al_2022_gpir].
-Others have made similar arguments through historical lenses by arguing that while race and class are highly correlated, the saliency of race in society and in our institutions continue to make it a salient group by which we structure our thinking about who has status and power [@helms_carter_1990;@dawson_1994_pup;@rile-hayward_2013].
-This perspective fits with many arguments that the United States is not as class-conscious as they are race-conscious [@zoltan_l_hajnal_dangerously_2020].
+A common touchstone of Social Dominance Theory suggests that members of a racial group with the highest status are likely to find ways to maintain the racial hierarchy [@lange_social_2012].
+From this perspective, many argue that Whites in the United States have and continue focus their efforts on maintaining influence in our institutions so that they may continue to wield political influence [@rile_hayward_how_2013;@lopez_white_2006].
+For example, @mutz_status_2018 finds little support for the claim that the 2016 Presidential Election -- which support for Trump is associated with an expression of racial identity among Whites [@jardina_disavowing_2021] -- was a response to the economic effects of a globalizing economy and instead argues that it was a response to concerns about the cultural changes brought about by globalization and the racial-diversity it promotes.
+Others demonstrate that the behavioral manifestations of a White constituency in the United States was a reflection of the saliency of race in politics [@agadjanian_changing_2021].
+
+A final example demonstrated that exposure to a story about discrimination about a White person lead individuals concerned with the racial hierarchy to report that Whites faced similar amounts of institutional racism as Blacks [@marshburn_dominance_2022]; that is, these Whites moved away from a color-blind ideology of race and expressed a more race-conscious ideology of White grievance [@smith_making_2023].
+Others have made similar arguments through historical lenses by arguing that while race and class are highly correlated, but that the correlation is due to a pattern of using class to disenfranchise non-Whites [See @rile_hayward_how_2013].
+Others point to the many examples of our institutions in the United States that continue to make race a salient group by which we structure our thinking about who has status and power [@helms_development_1990;@dawson_behind_1994;@rile_hayward_how_2013].
+This perspective fits with many arguments that the United States is not as class-conscious as they are race-conscious [@hajnal_dangerously_2020].
 
@@
 
-I am using a popular conceptualization of racial identity among Whites: that expression of it reflects homophily toward's one's own racial group [@ellemers_et-al_2002_arp;@jardina_2019_cup].
-As identity conceptualized this way may predict negative intergroup attitudes, it reflects a desire to find ways to find similarities rather than differences between one's self and other in-group members [@ellemers_et-al_2002_arp].
-Identity is largely an elusive concept to operationalize [see @leach_et-al_2008].
-Though there are implicit measures of White identity [see @knowles_peng_2005_jpsp], the common and measure of white identity is a response to the question, "How important is being White to your identity?" This measure captures a significant feature of an identity: the centrality or the importance of that identity to an individual [@jardina_2019_cup].
-Though a rather rough measure of something as complex as identity, it has been shown to produce similar results to more complex measures of white identity and holds a significant amount of face validity [@jardina_2019_cup].
+I am using a popular conceptualization of racial identity among Whites: that expression of it reflects homophily toward's one's own racial group [@ellemers_self_2002;@jardina_white_2019].
+As identity conceptualized this way may predict negative intergroup attitudes, it reflects a desire to find ways to find similarities rather than differences between one's self and other in-group members [@ellemers_self_2002].
+Identity is largely an elusive concept to operationalize [see @leach_group_2008].
+Though there are implicit measures of White identity [see @knowles_white_2005], the common and measure of white identity is a response to the question, "How important is being White to your identity?" This measure captures a significant feature of an identity: the centrality or the importance of that identity to an individual [@jardina_white_2019].
+Though a rather rough measure of something as complex as identity, it has been shown to produce similar results to more complex measures of white identity and holds a significant amount of face validity [@jardina_white_2019].
 
@@
 Status and power.
-Status refers to the respectability of the group whereas power refers to the influence that the group has [@craig_phillips_2023_ohpp].
+Status refers to the respectability of the group whereas power refers to the influence that the group has [@craig_group-based_2023].
 Threat here refers to either a manifestation of economic or political challenges to the current racial hierarchy.
@@
 We may expect that Whites who are facing bottom-up economic competition, will be those that experience the strongest desire to identify themselves with their racial group.
-As the dominant racial group in the United States, Whites that are unemployed, are surrounded by other Whites who are experiencing economic hardship and those who have limited capacity to switch careers are those that may be the most motivated to identify with the dominant racial group to defend Whites' influence and status in society [@bobo_hutchings_1996_asr].
+As the dominant racial group in the United States, Whites that are unemployed, are surrounded by other Whites who are experiencing economic hardship and those who have limited capacity to switch careers are those that may be the most motivated to identify with the dominant racial group to defend Whites' influence and status in society [@bobo_perceptions_1996].
 Therefore, measures of employment, finances, family finances, finances of those they know, and educational attainment all seem likely predictors of racial identity for Whites.
-This fits with measures that @dawson_1994_pup used to measure the influence of economics in Black American's identity.
-These measures also fit with @lindsay_2023_apr who found that many of these items predicted racial resentment among Whites.
+This fits with measures that @dawson_behind_1994 used to measure the influence of economics in Black American's identity.
+These measures also fit with @lindsay_relationship_2023 who found that many of these items predicted racial resentment among Whites.
 
 Here I conceptualize of political threat as perceptions that politics and political institutions no longer serves White people.
-Existing work demonstrates that Whites who express a desire to maintain the racial status quo with Whites on top are more likely to express that Whites are discriminated against at similar levels as Blacks [@marshburn_et-al_2022_gpir].
-As Whiteness is often a central component of Whites' conceptualization of American society [see @green_staerkle_2013], changes to "traditional American values" may reflect concern about changes to the influences of White values.
-Finally, a common finding is that expression of White identity is heavily wrapped up in the perception that racial stratifications are by meritocratic means [@lowery_et-al_2012_jpsp].
+Existing work demonstrates that Whites who express a desire to maintain the racial status quo with Whites on top are more likely to express that Whites are discriminated against at similar levels as Blacks [@marshburn_dominance_2022].
+As Whiteness is often a central component of Whites' conceptualization of American society [see @green_migration_2013], changes to "traditional American values" may reflect concern about changes to the influences of White values.
+Finally, a common finding is that expression of White identity is heavily wrapped up in the perception that racial stratifications are by meritocratic means [@lowery_paying_2012].
 As members of a dominant social group, when given information that either threatens one's belief that their merits are earned or that their group operates as a way to maintain inequality, we expect lower levels of expression in identification with that group.
 The classic items of racial resentment construct capture these ideas in that they revolve around the idea that Blacks are not disadvantaged as the result of systematic oppression by Whites.
-Though many have often used racial resentment as a measure of prejudice which would be a possible outcome of White identity, some make the compelling argument that it is distinct as a concept and operationalization [@davis_wilson_2022_ucp].
+Though many have often used racial resentment as a measure of prejudice which would be a possible outcome of White identity, some make the compelling argument that it is distinct as a concept and operationalization [@davis_racial_2022].
 
@@
 contain a number of these features.
-As a result, a burgeoning literature in political science that has made this connection takes the racial identity of whites more seriously as one that informs their political attitudes [see @jardina_2019_cup].
+As a result, a burgeoning literature in political science that has made this connection takes the racial identity of whites more seriously as one that informs their political attitudes [see @jardina_white_2019].
 Race in America is subjectively defined [see @rile-hayward_2013], and these definitions are deeply embedded and reinforced by our institutions [@lopez_2006].
@@
 ->
-Recent work makes the case that we should seriously consider whites' racial identity as an identity [@jardina_2019_cup].
+Recent work makes the case that we should seriously consider whites' racial identity as an identity [@jardina_white_2019].
 As an identity, much of the literature view these attitudes resulting from whites' racial identity as stemming from economic threat.
-Some studying white political identity find economic predictors as substantively and statistically significant [see @perez_et-al_2022_pp;@lindsay_2023_apr on their racial attitudes writ large].
+Some studying white political identity find economic predictors as substantively and statistically significant [see @perez_et-al_2022_pp;@lindsay_relationship_2023 on their racial attitudes writ large].
 Observational evidence supports this claim as well.
@@
 As much of the literature on whites' racial attitudes rely heavily on social identity theory as a paradigm, a significant portion of it presents the source of these group-based views as responses to economic threat.
-However, some disagree with this position [see @jardina_2019_cup] and view these attitudes as the result of changes to the racial hierarchy.
+However, some disagree with this position [see @jardina_white_2019] and view these attitudes as the result of changes to the racial hierarchy.
 These other perspectives provide explanations for whites' attitudes and conceptualize the definition of whiteness and white identity as a desire to retain political power.
@@
 The prominent account of white political identity and its implications in contemporary American politics argues for a strong connection between white political identity and a desire to protect this racial hierarchy and status.
-This conceptualization reflects the tendency to report in-group-oriented, not necessarily out-group-oriented, attitudes [@jardina_2019_cup].
+This conceptualization reflects the tendency to report in-group-oriented, not necessarily out-group-oriented, attitudes [@jardina_white_2019].
 
@@
 In 2012, the economy was still recovering.
-In 2016, Donald Trump became infamous for an election where he made race and gender extremely salient [@banda_cassese_2022_pb].
+In 2016, Donald Trump became infamous for an election where he made race and gender extremely salient [@banda_hostile_2022].
 Finally, in 2020, Trump had been in office for four years with an improving economy.
-Each of these three years reflect a different context and may offer the opportunity to observe how economic and political threats operate differrently.
+Each of these three years reflect a different context and may offer the opportunity to observe how economic and political threats operate differently.
 
@@
 The primary contribution of this paper is to examine which, of several competing measures, of different types of threat may predict white identity for individuals.
-As this requires an ombnibus test, I elect to use a penalized regression.
-I provide a more detailed discussion of the reasons why I use an estimator with an L2 regularization parameter, often refered to as a LASSO model in the supplementary materials.
+As this requires an omnibus test, I elect to use a penalized regression.
+I provide a more detailed discussion of the reasons why I use an estimator with an L2 regularization parameter, often referred to as a LASSO model in the supplementary materials.
 In practice, the model applies a penalization term so predictors that have relatively fewer explanatory power have much smaller coefficients than the variables that do provide significant predictive power.
@@
 Therefore, I apply a ordered logistic link function.
-Though it creates a more complicated computational task, it is a more appropriate model specification given the available measure and my assumptions about the concept [@liddell_kruschke_2018_jesp].
-
-To account for potential problems stemming from missing data, I impute my data by using multiple imputation with chained equations (MICE) where the particular model I use is a random forest model to allow for more flexibility and to reduce model dependence [@buuren_goothuis-oudshoorn_2011; @marbach_2022_pa].^[I impute 10 datasets and simultaneously fit my Bayesian LASSO on each dataset and pool the results.
+Though it creates a more complicated computational task, it is a more appropriate model specification given the available measure and my assumptions about the concept [@liddell_analyzing_2018].
+
+To account for potential problems stemming from missing data, I impute my data by using multiple imputation with chained equations (MICE) where the particular model I use is a random forest model to allow for more flexibility and to reduce model dependence [@buuren_mice_2011; @marbach_choosing_2022].^[I impute 10 datasets and simultaneously fit my Bayesian LASSO on each dataset and pool the results.
 Uncertainty reflected in my reported credible intervals not only reflects the uncertainty from one model, but my reported uncertainty also reflects the uncertainty generated from the imputation procedure.
@@
 #| label: fig-2012-2016-2020-ordinal-lasso
+#| fig-cap: "Economic factors are weak predictors of White Political Identity"
+#| fig-subcap:
+#|   - "Economic"
+#|   - "Political"
 #| layout-ncol: 2
-#| fig-cap: Economic factors are weak predictors of white identity
-#| fig-subcap:
-#|   - Economic
-#|   - Political
 # Graphs
@@
 An additional question worth examining is that economic or political threat may appear differently for partisans and for female respondents.
-Recent evidence suggests that racial attitudes are guided by partisanship [@engelhardt_2021_bjps].
+Recent evidence suggests that racial attitudes are guided by partisanship [@engelhardt_racial_2021].
 Specifically, some evidence suggests that White Democrats are much more sensitive to their racial identity than White Republicans due to the higher amount of racial heterogeneity within the Democratic party.
@@
 We therefore may expect that female individuals are more sensitive to economic threats than their male counterparts.
-In addition to the ombnibus model, I also include two additional models where I split the samples along the dimension of partisanship (including leaners in each) and on the dimension of sex.
+In addition to the omnibus model, I also include two additional models where I split the samples along the dimension of partisanship (including leaners in each) and on the dimension of sex.
 
@@
 #| label: fig-2012-2016-2020-female-male-lasso
+#| fig-cap: "Economic factors are weak predictors of white identity"
+#| fig-subcap:
+#|   - "Female, Economic"
+#|   - "Female, Political"
+#|   - "Male, Economic"
+#|   - "Female, Economic"
 #| layout-ncol: 2
-#| fig-cap: Economic factors are weak predictors of white identity
-#| fig-subcap:
-#|   - Female, Economic
-#|   - Female, Political
-#|   - Male, Economic
-#|   - Female, Economicß
 # Graphs
@@
 #| layout-ncol: 2
-#| fig-cap: Economic factors are weak predictors of white identity
+#| fig-cap: "Economic factors are weak predictors of white identity"
 #| fig-subcap:
-#|   - Republican, Economic
-#|   - Republican, Political
-#|   - Democrat, Economic
-#|   - Democrat, Economic
+#|   - "Republican, Economic"
+#|   - "Republican, Political"
+#|   - "Democrat, Economic"
+#|   - "Democrat, Economic"
 # Graphs
@@
 #| label: tbl-2012-descriptive-stats
-#| tbl-cap: 2012 ANES Descriptive Statistics
+#| tbl-cap: "2012 ANES Descriptive Statistics"
 datasummary_skim(
@@
 #| label: tbl-2016-descriptive-stats
-#| tbl-cap: 2016 ANES Descriptive Statistics
-
+#| tbl-cap: "2016 ANES Descriptive Statistics"
 datasummary_skim(
@@
 #| label: tbl-2020-descriptive-stats
-#| tbl-cap: 2020 ANES Descriptive Statistics
-
+#| tbl-cap: "2020 ANES Descriptive Statistics"
 datasummary_skim(
@@
 
-## Model{.appendix}
+## Model {.appendix}
 
@@
 
-Rather than use the classical LASSO with Leave One Out Cross Validation (LOOCV), I take advantage of a type of Markov Chain Monte Carlos (MCMC) called a Hamiltonian Monte Carlo for my LASSO due to a number of benefits outlined by @van-erp_et-al_2019_jmp.
+Rather than use the classical LASSO with Leave One Out Cross Validation (LOOCV), I take advantage of a type of Markov Chain Monte Carlos (MCMC) called a Hamiltonian Monte Carlo for my LASSO due to a number of benefits outlined by @van_erp_shrinkage_2019.
 As the literature leads me to hold *apriori* expectations about the distribution of predictors that meaningfully effect white political identity, I specify a Laplace distribution as my prior density function.
@@
 
+{{< pagebreak >}}
+
 ### Full tables of results {.appendix}
@@
 #| label: tbl-2012-lasso-simple
-#| tbl-cap: Predictors of White Identity in 2012
+#| tbl-cap: "Predictors of White Identity in 2012"
 model_2012 <- list(
@@
 #| label: tbl-2012-gender-lasso
-#| tbl-cap: Predictors of White Identity in 2012 conditional on gender
+#| tbl-cap: "Predictors of White Identity in 2012 conditional on gender"
 
@@
 #| label: tbl-2012-pid-lasso
-#| tbl-cap: Predictors of White Identity in 2012 conditional on partisanship
+#| tbl-cap: "Predictors of White Identity in 2012 conditional on partisanship"
 
@@
 #| label: tbl-2016-lasso-simple
-#| tbl-cap: Predictors of White Identity in 2016
+#| tbl-cap: "Predictors of White Identity in 2016"
 model_2016 <- list(
@@
 #| label: tbl-2016-gender-lasso
-#| tbl-cap: Predictors of White Identity in 2016 conditional on gender
+#| tbl-cap: "Predictors of White Identity in 2016 conditional on gender"
 
@@
 #| label: tbl-2016-pid-lasso
-#| tbl-cap: Predictors of White Identity in 2016 conditional on partisanship
+#| tbl-cap: "Predictors of White Identity in 2016 conditional on partisanship"
 
@@
 #| label: tbl-2020-lasso-simple
-#| tbl-cap: Predictors of White identity in 2020
+#| tbl-cap: "Predictors of White identity in 2020"
 model_2020 <- list(
@@
 #| label: tbl-2020-gender-lasso
-#| tbl-cap: Predictors of White Identity in 2020 conditional on gender
+#| tbl-cap: "Predictors of White Identity in 2020 conditional on gender"
 
@@
 #| label: tbl-2020-pid-lasso
-#| tbl-cap: Predictors of White Identity in 2012 conditional on partisanship
+#| tbl-cap: "Predictors of White Identity in 2012 conditional on partisanship"
 

### 2023-09-27 src/eda/dcr_anes_primary_analysis.R
update name of eda file

@@ -1,204 +1,0 @@
-# Title: White identity analyses
-
-# Notes:
-    #* Description: R script for 2012 & 2016 ANES and 2016-2020 ANES panel analyses
-    #* Updated: 2022-08-12
-    #* Updated by: dcr
-
-# Setup
-    #* Set seed
-set.seed(90210)
-    #* Source cleaning scripts
-source('code/dcr_anes_2012_cleaning.R')
-source('code/dcr_anes_2016_cleaning.R')
-source('code/dcr_anes_2020_cleaning.R')
-    #* Modularly load functions and packages
-box::use(
-    dplyr = dplyr[select, filter, mutate, case_when, rename],
-    tidyr = tidyr[drop_na],
-    mice = mice[mice, pool],
-    miceadds = miceadds[mids2datlist],
-    #future = future[plan, multisession],
-    #glmnet = glmnet[cv.glmnet],
-    #rstanarm = rstanarm[...],
-    brms = brms[brm, brm_multiple, set_prior, lasso, cumulative, prior],
-    #bayesreg = bayesreg[bayesreg],
-    tibble = tibble[tribble],
-    modelsummary = modelsummary[datasummary_skim, modelsummary],
-    parameters = parameters[parameters, standardize_names],
-    ggplot2 = ggplot2[ggplot, aes, position_dodge, theme_minimal, geom_vline, labs, theme, element_text, ggsave],
-    tidybayes = tidybayes[geom_pointinterval]
-)
-    #* create list object
-data = list()
-    #* Make sure the data you have loaded is the clean data
-        #** 2012
-data[['anes_2012']] = read.csv('data/anes-2012/anes-2012-updated.csv') |>
-    select(wid, female, rboff, pboff, ideo, epast, efuture, unpast, unnext, pid, age, edu, findjob, losejob, offwork, fairjblacks, immtakejobs, raceResent, ecfamily, linked, tradbreak, toofar, govbias, inflwhite, inflblack, whitediscrim)
-        #** 2016
-data[['anes_2016']] = read.csv('data/anes-2016/anes-2016-updated.csv') |>
-    select(rboff, pboff, epast, efuture, unpast, ecfamily, ecjob, tradbreak, pocare, ee, inflwhite, inflblack, govbias, whitediscrim, wid, raceResent, getahead, employment, pid, female, ideo, age, edu)
-        #** 2020
-data[['anes_2020']] = read.csv('data/anes-2020/anes-2020-updated.csv') |>
[…]

### 2023-09-27 src/eda/dcr_anes_primary_eda.R
pca didn't really work out here

@@ -1,0 +1,204 @@
+# Title: White identity analyses
+
+# Notes:
+    #* Description: R script for 2012 & 2016 ANES and 2016-2020 ANES panel analyses
+    #* Updated: 2022-08-12
+    #* Updated by: dcr
+
+# Setup
+    #* Set seed
+set.seed(90210)
+    #* Source cleaning scripts
+source('./eda/dcr_anes_2012_cleaning.R')
+source('./eda/dcr_anes_2016_cleaning.R')
+source('./eda/dcr_anes_2020_cleaning.R')
+    #* Modularly load functions and packages
+box::use(
+    dplyr = dplyr[select, filter, mutate, case_when, rename],
+    tidyr = tidyr[drop_na],
+    mice = mice[mice, pool],
+    miceadds = miceadds[mids2datlist],
+    #future = future[plan, multisession],
+    #glmnet = glmnet[cv.glmnet],
+    #rstanarm = rstanarm[...],
+    brms = brms[brm, brm_multiple, set_prior, lasso, cumulative, prior],
+    #bayesreg = bayesreg[bayesreg],
+    tibble = tibble[tribble],
+    modelsummary = modelsummary[datasummary_skim, modelsummary],
+    parameters = parameters[parameters, standardize_names],
+    ggplot2 = ggplot2[ggplot, aes, position_dodge, theme_minimal, geom_vline, labs, theme, element_text, ggsave],
+    tidybayes = tidybayes[geom_pointinterval]
+)
+    #* create list object
+data = list()
+    #* Make sure the data you have loaded is the clean data
+        #** 2012
+data[['anes_2012']] = read.csv('data/anes-2012/anes-2012-updated.csv') |>
+    select(wid, female, rboff, pboff, ideo, epast, efuture, unpast, unnext, pid, age, edu, findjob, losejob, offwork, fairjblacks, immtakejobs, raceResent, ecfamily, linked, tradbreak, toofar, govbias, inflwhite, inflblack, whitediscrim)
+        #** 2016
+data[['anes_2016']] = read.csv('data/anes-2016/anes-2016-updated.csv') |>
+    select(rboff, pboff, epast, efuture, unpast, ecfamily, ecjob, tradbreak, pocare, ee, inflwhite, inflblack, govbias, whitediscrim, wid, raceResent, getahead, employment, pid, female, ideo, age, edu)
+        #** 2020
+data[['anes_2020']] = read.csv('data/anes-2020/anes-2020-updated.csv') |>
[…]

### 2023-09-27 src/models/pca_models.R

@@ -1,0 +1,29 @@
+# Title: Exploratory PCA Model creation
+
+# Notes:
+  #* Description:
+    #** Checking out whether a PCA approach to handling the correlates with WID work
+  #* Updated:
+    #** 2023-09-27
+    #** dcr
+
+# Setup
+  #* Set seed
+set.seed(121022)
+  #* Source cleaning scripts
+source('./eda/dcr_anes_2012_cleaning.R')
+source('./eda/dcr_anes_2016_cleaning.R')
+source('./eda/dcr_anes_2020_cleaning.R')
+  #* Modularly load needed functions
+box::use(
+  FactoMineR[...]
+  , factoextra[...]
+  , data.table[...]
+)
+
+# PCA models
+data[['anes_2012_sub']] <- data[['anes_2012']][
+  , -c("wid", "pid", "female")
+]
+pca_2012 <- PCA(data[['anes_2012_sub']], ncp=2)
+fviz_cos2(pca_2012, choice = "var", addlabels = TRUE)

### 2023-09-27 src/models/pca_models.R
adding a quick breadcrumb

@@ -25,5 +25,5 @@ box::use(
 data[['anes_2012_sub']] <- data[['anes_2012']][
   , -c("wid", "pid", "female")
 ]
-pca_2012 <- PCA(data[['anes_2012_sub']], ncp=2)
+pca_2012 <- PCA(data[['anes_2012_sub']], ncp=2) # at most, there are like 5 dimensions and they all have weird components
 fviz_cos2(pca_2012, choice = "var", addlabels = TRUE)

### 2023-09-28 src/manuscript.qmd
final edits to draft before sending out again

@@
 title: |
-  Economic concerns appear to be weak predictors of white political identity
+  Economic concerns appear to be weak predictors of White political identity
 short-title: Economic concerns
 published: "Manuscript draft: please do not share without author permission."
-code-repo: "For replication, go to: <https:://github.com/DamonCharlesRoberts/white-identity-sources>."
+code-repo: "For replication, go to: <https:://github.com/DamonCharlesRoberts/White-identity-sources>."
 author:
@@
 abstract: |
-    Do economic or political threats explain reported white identity?
-Overall, the social identity literature would suggest that white identity would increase in response to economic threats.
-However, a number of those that study white identity, specifically, argue that it results from concern about political influence.
-Considering what whiteness means historically and contemporaneously, I argue that we should expect that political threats reflect stronger associations with white identity.
+    Do economic or political threats explain reported White Political Identity?
+Overall, the social identity literature would suggest that White Political Identity would increase in response to economic threats.
+However, a number of those that study White Political Identity, specifically, argue that it results from concern about political influence.
+Considering what Whiteness means historically and contemporaneously, I argue that we should expect that political threats reflect stronger associations with White Political Identity.
 Using data from the 2012, 2016, and 2020 American National Election Study, I consider a single model using penalized regression containing proxies of economic and political threats.
-I find evidence suggesting that in the post-Trump era, white political identity is strongly associated with reported feelings of Whites’ loss of political influence as opposed to economic threats, as some suggest and may expect.
+I find evidence suggesting that in the post-Trump era, White political identity is strongly associated with reported feelings of Whites’ loss of political influence as opposed to economic threats, as some suggest and may expect.
 thanks: |
@@
 keywords:
-  - white identity
+  - White Political Identity
   - economic threat
@@
 
-There is a growing literature suggesting that White Political Identity predicts a number of outcomes.
-For example evidence suggests that White Political Identity predicts support for Trump in the 2016 Presidential Election [@jardina_disavowing_2021].
-The existing literature is less clear about the predictors of White Political Identity, however.
+There is a growing literature suggesting that White Political Identity predicts a number of important political outcomes.
+For example, some evidence suggests that White Political Identity predicts support for Trump in the 2016 Presidential Election [@jardina_disavowing_2021].
+Nevertheless, the existing literature is a bit less clear about the predictors of White Political Identity.
 I advance an argument as to why we might consider the expression of a White Political Identity to be a reflection of concerns around cultural threat as opposed to economic threat.
@@
 As the American National Election Study is a long-running survey of those residing in the United States, it provides some of the most consistent measures on common political attitudes and behaviors.
-Using data from the 2012, 2016, and 2020 administrations of the ANES, I fit an omnibus model with penalized regression to examine the degree to which additive scales and individual measures of economic and cultural influence explain an individual's uptake of White identity as well as to examine the degree to which changes in the social context may increase or decrease the explanatory power of these measures.
+Using data from the 2012, 2016, and 2020 administrations of the ANES, I fit an omnibus model with penalized regression to examine the degree to which individual measures of economic and cultural status and perceptions of influence explain an individual's uptake of White Political Identity as well as to examine the degree to which changes in the social context may increase or decrease the explanatory power of these measures.
 From this test, I find that measures reporting concerns about Whites' cultural influence are stronger explanatory factors than proxies of economic influence in the post-Trump era.
@@
 Others have experimental evidence supporting this perspective that White Political Identity is motivated by material concerns [@perez_politics_2022].
-Further, @jardina_white_2019 who provides one of the most comprehensive treatments of conceptualizing, measuring, and examining the outcomes of White Political Identity includes indicators of economic status and power in the primary model.
+Further, @jardina_White_2019 who provides one of the most comprehensive treatments of conceptualizing, measuring, and examining the outcomes of White Political Identity includes indicators of economic status and power in the primary model.
 
@@
 A common touchstone of Social Dominance Theory suggests that members of a racial group with the highest status are likely to find ways to maintain the racial hierarchy [@lange_social_2012].
-From this perspective, many argue that Whites in the United States have and continue focus their efforts on maintaining influence in our institutions so that they may continue to wield political influence [@rile_hayward_how_2013;@lopez_white_2006].
+From this perspective, many argue that Whites in the United States have and continue focus their efforts on maintaining influence in our institutions so that they may continue to wield political influence [@rile_hayward_how_2013;@lopez_White_2006].
 For example, @mutz_status_2018 finds little support for the claim that the 2016 Presidential Election -- which support for Trump is associated with an expression of racial identity among Whites [@jardina_disavowing_2021] -- was a response to the economic effects of a globalizing economy and instead argues that it was a response to concerns about the cultural changes brought about by globalization and the racial-diversity it promotes.
@@
 
-I am using a popular conceptualization of racial identity among Whites: that expression of it reflects homophily toward's one's own racial group [@ellemers_self_2002;@jardina_white_2019].
+In this note, I am using a popular conceptualization of racial identity among Whites: that expression of it reflects homophily toward's one's own racial group [@ellemers_self_2002;@jardina_White_2019].
 As identity conceptualized this way may predict negative intergroup attitudes, it reflects a desire to find ways to find similarities rather than differences between one's self and other in-group members [@ellemers_self_2002].
 Identity is largely an elusive concept to operationalize [see @leach_group_2008].
-Though there are implicit measures of White identity [see @knowles_white_2005], the common and measure of white identity is a response to the question, "How important is being White to your identity?" This measure captures a significant feature of an identity: the centrality or the importance of that identity to an individual [@jardina_white_2019].
-Though a rather rough measure of something as complex as identity, it has been shown to produce similar results to more complex measures of white identity and holds a significant amount of face validity [@jardina_white_2019].
+Though there are implicit measures of White Political Identity [see @knowles_White_2005], the common and measure of White Political Identity is a response to the question, "How important is being White to your identity?" This measure captures a significant feature of an identity: the centrality or the importance of that identity to an individual [@jardina_White_2019].
+Though a rather rough measure of something as complex as identity, it has been shown to produce similar results to more complex measures of White Political Identity and holds a significant amount of face validity [@jardina_White_2019].
 
@@
 Status and power.
-Status refers to the respectability of the group whereas power refers to the influence that the group has [@craig_group-based_2023].
-Threat here refers to either a manifestation of economic or political challenges to the current racial hierarchy.
+Status refers to the respectability of the group whereas power refers to the influence of the group [@craig_group-based_2023].
+Threat refers to either a manifestation of economic or political challenges to the current racial hierarchy.
 I do not distinguish between these two sources of threat as common survey measures are too crude to distinguish between how the threat manifests.
-Here, I am focusing on what types of non-material threats activate White identity in the context of politics and will, as a shorthand refer to political, cultural, and societal threats as political threats.
+Here, I am focusing on what types of non-material threats activate White Political Identity in the context of politics and will, as a shorthand refer to political, cultural, and societal threats as political threats.
 
@@
 Therefore, measures of employment, finances, family finances, finances of those they know, and educational attainment all seem likely predictors of racial identity for Whites.
-This fits with measures that @dawson_behind_1994 used to measure the influence of economics in Black American's identity.
+This fits with measures that @dawson_behind_1994 used to examine the influence of economics in Black Americans' identity.
 These measures also fit with @lindsay_relationship_2023 who found that many of these items predicted racial resentment among Whites.
 
-Here I conceptualize of political threat as perceptions that politics and political institutions no longer serves White people.
-Existing work demonstrates that Whites who express a desire to maintain the racial status quo with Whites on top are more likely to express that Whites are discriminated against at similar levels as Blacks [@marshburn_dominance_2022].
+Here I conceptualize of political threat as perceptions that politics and political institutions no longer primarily serves White people.
+Existing work demonstrates that Whites who express a desire to maintain the racial status quo with Whites on top are more likely to express the belief that Whites are discriminated against at similar levels as Blacks [@marshburn_dominance_2022].
 As Whiteness is often a central component of Whites' conceptualization of American society [see @green_migration_2013], changes to "traditional American values" may reflect concern about changes to the influences of White values.
-Finally, a common finding is that expression of White identity is heavily wrapped up in the perception that racial stratifications are by meritocratic means [@lowery_paying_2012].
+Finally, a common finding is that expression of White Political Identity is heavily wrapped up in the perception that racial stratifications are by meritocratic means [@lowery_paying_2012].
 As members of a dominant social group, when given information that either threatens one's belief that their merits are earned or that their group operates as a way to maintain inequality, we expect lower levels of expression in identification with that group.
 The classic items of racial resentment construct capture these ideas in that they revolve around the idea that Blacks are not disadvantaged as the result of systematic oppression by Whites.
-Though many have often used racial resentment as a measure of prejudice which would be a possible outcome of White identity, some make the compelling argument that it is distinct as a concept and operationalization [@davis_racial_2022].
+Though many have often used racial resentment as a measure of prejudice which would be a possible outcome of White Political Identity, some make the compelling argument that it is distinct as a concept and operationalization [@davis_racial_2022].
 
@@
 contain a number of these features.
-As a result, a burgeoning literature in political science that has made this connection takes the racial identity of whites more seriously as one that informs their political attitudes [see @jardina_white_2019].
+As a result, a burgeoning literature in political science that has made this connection takes the racial identity of Whites more seriously as one that informs their political attitudes [see @jardina_White_2019].
 Race in America is subjectively defined [see @rile-hayward_2013], and these definitions are deeply embedded and reinforced by our institutions [@lopez_2006].
-Whiteness and one's willingness to identify with that group is important [@wong_cho_2005] as it reflects a desire for an association with the history and affordances of being white [@mcdermott_samson_2005].
+Whiteness and one's willingness to identify with that group is important [@wong_cho_2005] as it reflects a desire for an association with the history and affordances of being White [@mcdermott_samson_2005].
 ->
-Recent work makes the case that we should seriously consider whites' racial identity as an identity [@jardina_white_2019].
-As an identity, much of the literature view these attitudes resulting from whites' racial identity as stemming from economic threat.
-Some studying white political identity find economic predictors as substantively and statistically significant [see @perez_et-al_2022_pp;@lindsay_relationship_2023 on their racial attitudes writ large].
+Recent work makes the case that we should seriously consider Whites' racial identity as an identity [@jardina_White_2019].
+As an identity, much of the literature view these attitudes resulting from Whites' racial identity as stemming from economic threat.
+Some studying White political identity find economic predictors as substantively and statistically significant [see @perez_et-al_2022_pp;@lindsay_relationship_2023 on their racial attitudes writ large].
 Observational evidence supports this claim as well.
-For example, Trump's messaging about job market competition with immigrants appeared to be successful at attracting white voters [@bunyasi_2019].
+For example, Trump's messaging about job market competition with immigrants appeared to be successful at attracting White voters [@bunyasi_2019].
 After all, the identity literature, more broadly, often focuses on economic threats.
@@
 Many political scientists use this as an explanation for negative attitudes directed toward immigrants among those who feel insecure about their job prospects [@pardos-prado_xena_2019].
-Others argue that deindustrialization and job loss activate whites' concerns about the racial hierarchy and predict their support for Republican challengers that express support for keeping the racial status quo [@baccini_weymouth_2021_apsr].
-
-As much of the literature on whites' racial attitudes rely heavily on social identity theory as a paradigm, a significant portion of it presents the source of these group-based views as responses to economic threat.
-However, some disagree with this position [see @jardina_white_2019] and view these attitudes as the result of changes to the racial hierarchy.
-These other perspectives provide explanations for whites' attitudes and conceptualize the definition of whiteness and white identity as a desire to retain political power.
-
-# Why might white identity be a backlash to loss of institutional privelege?
+Others argue that deindustrialization and job loss activate Whites' concerns about the racial hierarchy and predict their support for Republican challengers that express support for keeping the racial status quo [@baccini_weymouth_2021_apsr].
+
+As much of the literature on Whites' racial attitudes rely heavily on social identity theory as a paradigm, a significant portion of it presents the source of these group-based views as responses to economic threat.
+However, some disagree with this position [see @jardina_White_2019] and view these attitudes as the result of changes to the racial hierarchy.
+These other perspectives provide explanations for Whites' attitudes and conceptualize the definition of Whiteness and White Political Identity as a desire to retain political power.
+
+# Why might White Political Identity be a backlash to loss of institutional privelege?
 
 While a critical component of identity is a feeling of your fate as linked with others in your group [@leach_et-al_2008], this linked fate is the connection between attitudes and material forms of threat but that identity, in and of itself, is better conceptualized as a result of cultural or ideational threat [@donnelly_2020].
-Aligning with the contextual account of racial identity for whites, many conceptualize white identity as an expressed desire for a recognition of privilege and status by institutions [@lopez_2006] and a connection with the history that connects whites to their status [@mcdermott_samson_2005].
-The prominent account of white political identity and its implications in contemporary American politics argues for a strong connection between white political identity and a desire to protect this racial hierarchy and status.
-This conceptualization reflects the tendency to report in-group-oriented, not necessarily out-group-oriented, attitudes [@jardina_white_2019].
-
-Whites' institutional privelege and threats to it are the more potent threats that whites may be less used to and as a result, coalesce as a group to defend.
-@winter_2008 demonstrates that whites' negative attitudes toward social welfare programs are motivated by perceptions that beneficiaries are Black.
-In this book length examination of the question, it is not that the people are poor and could use government assistance, it is that those receiving help are not white.
-Further, the increasing racial polarization is not something that is occuring among whites who are most economically threatened, but the Republican party is whiter with both wealthy and less wealthy whites -- political considerations are causally prior to racial attitudes [@engelhardt_2021_bjps; see also @schaffner_2022_bjps].
-
-An example of a contemporary institutional threat to whites is the election of Barack Obama.
+Aligning with the contextual account of racial identity for Whites, many conceptualize White Political Identity as an expressed desire for a recognition of privilege and status by institutions [@lopez_2006] and a connection with the history that connects Whites to their status [@mcdermott_samson_2005].
+The prominent account of White political identity and its implications in contemporary American politics argues for a strong connection between White political identity and a desire to protect this racial hierarchy and status.
+This conceptualization reflects the tendency to report in-group-oriented, not necessarily out-group-oriented, attitudes [@jardina_White_2019].
+
+Whites' institutional privelege and threats to it are the more potent threats that Whites may be less used to and as a result, coalesce as a group to defend.
+@winter_2008 demonstrates that Whites' negative attitudes toward social welfare programs are motivated by perceptions that beneficiaries are Black.
+In this book length examination of the question, it is not that the people are poor and could use government assistance, it is that those receiving help are not White.
+Further, the increasing racial polarization is not something that is occuring among Whites who are most economically threatened, but the Republican party is Whiter with both wealthy and less wealthy Whites -- political considerations are causally prior to racial attitudes [@engelhardt_2021_bjps; see also @schaffner_2022_bjps].
+
+An example of a contemporary institutional threat to Whites is the election of Barack Obama.
 Evidence suggests that America did not shift into a post-racial era as some political commentators suspected.
-In fact, race became an extremely salient issue for whites [@tesler_2016].
+In fact, race became an extremely salient issue for Whites [@tesler_2016].
 White Republicans exhibited higher levels of old fashioned racism [@tesler_2013].
 Issues were racialized during Obama's presidency and many Whites exhibited lower levels of support for those policies in response [@tesler_2012;@tesler_2016].
-As a result of whites' intense concern with race, non-white political candidates, such as Obama, often had to distance themselves from their racial groups in speeches before white audiences as a way to not activate concerns about being to beholden to their racial group [@stephens-dougan_2020].
-The reason for such potent responses to indicators of change for what groups have influence in political instittuions is simply that whites are not used to it.
-They have spent the entirety of this country's existence defending the idea that whites have higher status in American society regardless of wealth, education, or other forms of status.
+As a result of Whites' intense concern with race, non-White political candidates, such as Obama, often had to distance themselves from their racial groups in speeches before White audiences as a way to not activate concerns about being to beholden to their racial group [@stephens-dougan_2020].
+The reason for such potent responses to indicators of change for what groups have influence in political instittuions is simply that Whites are not used to it.
+They have spent the entirety of this country's existence defending the idea that Whites have higher status in American society regardless of wealth, education, or other forms of status.
 
 My argument is not a new one.
-Historical accounts outlining the complicated history of what it meant to be "white" from a legal standpoint boils down to: did other whites see you as being good enough to be called white?
-There are many examples of groups that are now considered white but have not always been so.
-For example groups from eastern europe were often considered as non-white.
-While these definitions often correlated with economic status (e.g., German immigrants were often considered white but not Irish immigrants), they often changed in response to concerns about weakening white political influence with other groups growing in political influence such as the Irish in Boston.
-During these shifts in definitions of whiteness were no longer about class.
+Historical accounts outlining the complicated history of what it meant to be "White" from a legal standpoint boils down to: did other Whites see you as being good enough to be called White?
+There are many examples of groups that are now considered White but have not always been so.
+For example groups from eastern europe were often considered as non-White.
+While these definitions often correlated with economic status (e.g., German immigrants were often considered White but not Irish immigrants), they often changed in response to concerns about weakening White political influence with other groups growing in political influence such as the Irish in Boston.
+During these shifts in definitions of Whiteness were no longer about class.
 Instead, it became about what it meant for access and privileges afforded by institutions [@rile-hayward_2013].
 In the $19^{th}$ century, privilege came from one's freedom and wealth.
-Those who were not enslaved but were poor were not considered white and were of a lower strata than whites who owned land.
+Those who were not enslaved but were poor were not considered White and were of a lower strata than Whites who owned land.
 With the abolition of slavery, the expressed (explicit) concern became not about wealth but now about race.
 Whites could no longer use the classification of "free" and "land-owning" as an excuse to exclude Blacks from full access to America's institutions.
-This motivated a shift in how the institutions and how whites defined themselves [@rile-hayward_2013].
-It became the focus for whites to be a higher status group, regardless of whether or not they were poorer than formerly enslaved Black Americans [@rile-hayward_2013].
-In other words, the white American tradition is to hold one's whiteness as a status symbol regardless of other forms of status such as class.
+This motivated a shift in how the institutions and how Whites defined themselves [@rile-hayward_2013].
+It became the focus for Whites to be a higher status group, regardless of whether or not they were poorer than formerly enslaved Black Americans [@rile-hayward_2013].
+In other words, the White American tradition is to hold one's Whiteness as a status symbol regardless of other forms of status such as class.
 
@@
 In an effort to address this question in a bite-sized chunk, I first start with examining whether or not correlational evidence aligns with my claim that individual-level economic predictors are weaker than perceptions of political influence.
-In the next section, I examine whether attitudes expressing concern with economic or political threats are more robust correlates of the degree to which someone feels that being white is important to their identity.
-The goal is not to test particular causal mechanisms but to examine whether my belief and others' that white identity reflects reactions to political threats more than it does economic.
+In the next section, I examine whether attitudes expressing concern with economic or political threats are more robust correlates of the degree to which someone feels that being White is important to their identity.
+The goal is not to test particular causal mechanisms but to examine whether my belief and others' that White Political Identity reflects reactions to political threats more than it does economic.
 -->
@@
 
-To examine whether whites' attitudes expressing concern about political threats are more robust correlates than their perceptions of the economy, I use the 2012, 2016, and 2020 American National Election Studies (ANES).
-The studies are nationally representative questions that capture several questions that lend themselves to testing the relationship between economic and political concerns while also asking white respondents to report how important being white is to their identity.
-
-Though I am using the more crude measure of White identity and am limited in the predictors I can choose from, a larger concern is that the degree to which these threats may vary as a result of changes to the political landscape.
+To examine whether Whites' attitudes expressing concern about political threats are more robust correlates than their perceptions of the economy, I use the 2012, 2016, and 2020 American National Election Studies (ANES).
+The studies are nationally representative questions that capture several questions that lend themselves to testing the relationship between economic and political concerns while also asking White respondents to report how important being White is to their identity.
+
+Though I am using the more crude measure of White Political Identity and am limited in the predictors I can choose from, a larger concern is that the degree to which these threats may vary as a result of changes to the political landscape.
 Specifically, in 2008 the United States experienced a massive recession that lead hit the housing market quite hard.
@@
 
-The primary contribution of this paper is to examine which, of several competing measures, of different types of threat may predict white identity for individuals.
+The primary contribution of this paper is to examine which, of several competing measures, of different types of threat may predict White Political Identity for individuals.
 As this requires an omnibus test, I elect to use a penalized regression.
@@
 
-My measure of white identity is ordinal.
+My measure of White Political Identity is ordinal.
 Therefore, I apply a ordered logistic link function.
@@
 @fig-2012-2016-2020-ordinal-lasso presents the estimates of my fitted ordinal models with the LASSO shrinkage prior.^[The full tables of results are included in the Supplementary Information.] These models are fit on 6 chains with 2000 iterations each.
-To examine the evidence of whether a particular predictor "matters" for predicting white political identity, I examine the estimated credible intervals, which report the probability that the true estimate falls within the estimated range.
+To examine the evidence of whether a particular predictor "matters" for predicting White political identity, I examine the estimated credible intervals, which report the probability that the true estimate falls within the estimated range.
 I construct these credible intervals by reporting estimates between the 2.5% and 97.5% quantiles.
@@
       , term == 'b_tradbreak' ~ 'New lifestyles breaking society'
-      , term == 'b_govbias' ~ 'Gov. biased against whites'
-      , term == 'b_inflwhite' ~ 'whites influence politics'
-      , term == 'b_whitediscrim' ~ 'whites are discriminated against'
+      , term == 'b_govbias' ~ 'Gov. biased against Whites'
+      , term == 'b_inflwhite' ~ 'Whites influence politics'
+      , term == 'b_whitediscrim' ~ 'Whites are discriminated against'
       , term == 'b_raceResent' ~ 'Racial resentment'
@@
         , 'New lifestyles breaking society'
-        , 'Gov. biased against whites'
-        , 'whites influence politics'
-        , 'whites are discriminated against'
+        , 'Gov. biased against Whites'
+        , 'Whites influence politics'
+        , 'Whites are discriminated against'
         , 'Racial resentment'
@@
   filter(
-    term == "New lifestyles breaking society" | term == "Gov. biased against whites" |
-    term == "whites influence politics" | term == "whites are discriminated against" |
+    term == "New lifestyles breaking society" | term == "Gov. biased against Whites" |
+    term == "Whites influence politics" | term == "Whites are discriminated against" |
     term == "Racial resentment"
@@
 Examining the figure, we see that none of the proxies for economic threat consistently correlate with White political identity; however there is at least one proxy of political threat that does.
-For economic threats, we see that when the economy was recovering from the recession and Obama was in Office, those who perceived that the economy was better off retrospectively and with a lower reported level of education were correlated with expressing a higher White identity.
-However, when the economy was better and Trump was in office, we see that those less worried about losing a job were reported a higher white identity.
-
-For political threats, we see that in both instances where Obama was in the White House that those who reported being concerned with "new lifestyles" as well as those with a higher score on the racial resentment measures were associated with having higher white political identities.
+For economic threats, we see that when the economy was recovering from the recession and Obama was in Office, those who perceived that the economy was better off retrospectively and with a lower reported level of education were correlated with expressing a higher White Political Identity.
+However, when the economy was better and Trump was in office, we see that those less worried about losing a job were reported a higher White Political Identity.
+
+For political threats, we see that in both instances where Obama was in the White House that those who reported being concerned with "new lifestyles" as well as those with a higher score on the racial resentment measures were associated with having higher White political identities.
 When Trump was in the White House, we see that those who felt that the government was biased against Whites, that Whites were discriminated against, and who scored higher on the racial resentment measures all reported higher levels of White political identity.
@@
 An additional question worth examining is that economic or political threat may appear differently for partisans and for female respondents.
-Recent evidence suggests that racial attitudes are guided by partisanship [@engelhardt_racial_2021].
-Specifically, some evidence suggests that White Democrats are much more sensitive to their racial identity than White Republicans due to the higher amount of racial heterogeneity within the Democratic party.
+Evidence suggests that racial attitudes are guided by partisanship [@engelhardt_racial_2021].
+Specifically, some evidence suggests that White Democrats are much more sensitive to their racial identity than White Republicans due to the higher amount of racial heterogeneity within the Democratic party [@perez_politics_2022].
 This may mean for White Democrats that they have more exposure to intergroup competition for power within the party (such as the choice to support Hillary Clinton versus Barack Obama) and may have higher White identification as a result.
@@
 We therefore may expect that female individuals are more sensitive to economic threats than their male counterparts.
-In addition to the omnibus model, I also include two additional models where I split the samples along the dimension of partisanship (including leaners in each) and on the dimension of sex.
+In addition to the omnibus model, I also include two additional models where I split the samples along the dimension of partisanship (including "leaners" in each) and on the dimension of sex.
 
@@
 #| label: fig-2012-2016-2020-female-male-lasso
-#| fig-cap: "Economic factors are weak predictors of white identity"
+#| fig-cap: "Economic factors are weak predictors of White Political Identity"
 #| fig-subcap:
@@
 #| layout-ncol: 2
-#| fig-cap: "Economic factors are weak predictors of white identity"
+#| fig-cap: "Economic factors are weak predictors of White Political Identity"
 #| fig-subcap:
@@
 
-Overall, these results suggest that common proxies of economic threat are at least inconsistent measures of White identity.
-Even in a time where the economy is recovering from a collapse, more proxies of political threat seemed to predict White identity.
+Overall, these results suggest that common proxies of economic threat are at least inconsistent measures of White Political Identity.
+Even in a time where the economy is recovering from a collapse, more proxies of political threat seemed to predict White Political Identity.
 When a White president well-known for his penchance for racial dog-whistles takes office, we see support for the idea that White grievance politics was at a high as evidenced by those who felt that political institutions and that the direction of society was turning away from the racial status quo were those that reported their racial identity as more central to who they are.
@@
 
-These findings provide a direction for scholars to take when examining the factors driving Whites to take on a white political identity.
-While this particular analysis is not causal, it provides a basis for those wondering whether, when put together, economic or political threats are more fruitful to engender a rise in white political identity.
+These findings provide a direction for scholars to take when examining the factors driving Whites to take on a White political identity.
+While this particular analysis is not causal, it provides a basis for those wondering whether, when put together, economic or political threats are more fruitful to engender a rise in White political identity.
 
@@
 +==============================================+============================================================+==============================================================================================================================================================+===========================================================================================================================================================================+==================+
-| white identity                               | How important is being white to your identity              | 1 = Extremely important; 2 = Very important; 3 = Moderately important; 4 = A little important; 5 = Not at all important; \<1 Missing, not asked, etc         | 1 = Not at all important; 2 = A little important; 3 = Moderately important; 4 = Very important; 5 = Extremely important; NA = Missing, not asked, etc                     | 2012, 2016, 2020 |
+| White Political Identity                               | How important is being White to your identity              | 1 = Extremely important; 2 = Very important; 3 = Moderately important; 4 = A little important; 5 = Not at all important; \<1 Missing, not asked, etc         | 1 = Not at all important; 2 = A little important; 3 = Moderately important; 4 = Very important; 5 = Extremely important; NA = Missing, not asked, etc                     | 2012, 2016, 2020 |
 +----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
@@
 +----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
-| Government biased against whites             | Does the Administration favor Blacks or whites (2012 only) | 1 = Favors whites; 2 = Favors Blacks; 3 = Treats both the same; \< 1 Missing, not asked, etc                                                                 | -1 = Favors whites; 0 = Treats both the same; 1 = Favors Blacks; NA = Missing, not asked, etc                                                                             | 2012, 2016, 2020 |
+| Government biased against Whites             | Does the Administration favor Blacks or Whites (2012 only) | 1 = Favors Whites; 2 = Favors Blacks; 3 = Treats both the same; \< 1 Missing, not asked, etc                                                                 | -1 = Favors Whites; 0 = Treats both the same; 1 = Favors Blacks; NA = Missing, not asked, etc                                                                             | 2012, 2016, 2020 |
 |                                              |                                                            |                                                                                                                                                              |                                                                                                                                                                           |                  |
-|                                              | Does the Federal Gov treat Blacks or whites Better         |                                                                                                                                                              |                                                                                                                                                                           |                  |
-+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
-| whites influence politics                    | How much influence do whites have on U.S.
+|                                              | Does the Federal Gov treat Blacks or Whites Better         |                                                                                                                                                              |                                                                                                                                                                           |                  |
++----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
+| Whites influence politics                    | How much influence do Whites have on U.S.
 politics         | 1 = Too much; 2 = Just about right; 3 = Too little; \< 1 Missing, not asked, etc                                                                             | -1 = Too little; 0 = Just about right; 1 = Too much; NA = Missing, not asked, etc                                                                                         | 2012, 2016, 2020 |
 +----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
-| whites are discriminated against             | Discrimination in U.S.
-against whites                      | 1 = A great deal; 2 = A lot; 3 = A moderate amount; 4 = A little; 5 = None at all; \< 1 Missing, not asked, etc                                              | 1 = None at all; 2 = A little; 3 = A moderate amount; 4 = A lot; 5 = A great deal; NA = Missing, not asked, etc                                                           | 2012, 2016, 2020 |
+| Whites are discriminated against             | Discrimination in U.S.
+against Whites                      | 1 = A great deal; 2 = A lot; 3 = A moderate amount; 4 = A little; 5 = None at all; \< 1 Missing, not asked, etc                                              | 1 = None at all; 2 = A little; 3 = A moderate amount; 4 = A lot; 5 = A great deal; NA = Missing, not asked, etc                                                           | 2012, 2016, 2020 |
 +----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
@@
   , new = c(
-    "White identity", "Retrospective better off", "Economy better", "Unemployment better", "Education", "Income", "Worried about losing job"
-    , "Worried about family finances", "Know someone lost job", "New lifestyles breaking tradition", "Government biased against whites", "Whites influence politics"
+    "White Political Identity", "Retrospective better off", "Economy better", "Unemployment better", "Education", "Income", "Worried about losing job"
+    , "Worried about family finances", "Know someone lost job", "New lifestyles breaking tradition", "Government biased against Whites", "Whites influence politics"
     , "Whites discrimminated against", "Racial resentment", "Party ID", "Female"
@@
   , new = c(
-    "White identity", "Retrospective better off", "Economy better", "Unemployment better", "Education", "Income", "Worried about losing job"
-    , "Worried about family finances", "Know someone lost job", "Government biased against whites", "Whites influence politics"
+    "White Political Identity", "Retrospective better off", "Economy better", "Unemployment better", "Education", "Income", "Worried about losing job"
+    , "Worried about family finances", "Know someone lost job", "Government biased against Whites", "Whites influence politics"
     , "Whites discrimminated against", "Racial resentment", "Party ID", "Female"
@@
 
-The literature surveyed above suggests that there are a small number of variables that do predict white political identity, and there are many others that have no effect.
+The literature surveyed above suggests that there are a small number of variables that do predict White political identity, and there are many others that have no effect.
 Imagining an assumed hypothesis about the distribution of coefficients, this implies that the L2 regularization parameter equals 0.
@@
 Rather than use the classical LASSO with Leave One Out Cross Validation (LOOCV), I take advantage of a type of Markov Chain Monte Carlos (MCMC) called a Hamiltonian Monte Carlo for my LASSO due to a number of benefits outlined by @van_erp_shrinkage_2019.
-As the literature leads me to hold *apriori* expectations about the distribution of predictors that meaningfully effect white political identity, I specify a Laplace distribution as my prior density function.
+As the literature leads me to hold *apriori* expectations about the distribution of predictors that meaningfully effect White political identity, I specify a Laplace distribution as my prior density function.
 I suspect that there are a large number of included predictors that have null effects and a relatively smaller proportion of predictors that *do* have an effect but am relatively unsure about how large their coefficients might be.
@@
 cm <- c('b_rboff' = 'Retrospective better off', 'b_epast' = 'Economy better', 'b_unpast' = 'Unemployment better', 'b_edu' = 'Education', 'b_income' = 'Income', 'b_losejob' = 'Worried about losing job', 'b_ecfamily' = 'Worried about family finances', 'b_ecjob' = 'Know someone lost job', 'b_tradbreak' = 'New lifestyles breaking society', 'b_govbias' = 'Gov.
-biased against whites', 'b_inflwhite' = 'whites influence politics', 'b_whitediscrim' = 'whites are discriminated against', 'b_raceResent' = 'Racial resentment', 'b_pid' = 'Party ID', 'b_female' = 'Female', 'b_Intercept[1]' = 'Threshold 1', 'b_Intercept[2]' = 'Threshold 2', 'b_Intercept[3]' = 'Threshold 3', 'b_Intercept[4]' = 'Threshold 4')
+biased against Whites', 'b_inflwhite' = 'Whites influence politics', 'b_whitediscrim' = 'Whites are discriminated against', 'b_raceResent' = 'Racial resentment', 'b_pid' = 'Party ID', 'b_female' = 'Female', 'b_Intercept[1]' = 'Threshold 1', 'b_Intercept[2]' = 'Threshold 2', 'b_Intercept[3]' = 'Threshold 3', 'b_Intercept[4]' = 'Threshold 4')
 
 cm_gender <- c('b_rboff' = 'Retrospective better off', 'b_epast' = 'Economy better', 'b_unpast' = 'Unemployment better', 'b_edu' = 'Education', 'b_income' = 'Income', 'b_losejob' = 'Worried about losing job', 'b_ecfamily' = 'Worried about family finances', 'b_ecjob' = 'Know someone lost job', 'b_tradbreak' = 'New lifestyles breaking society', 'b_govbias' = 'Gov.
-biased against whites', 'b_inflwhite' = 'whites influence politics', 'b_whitediscrim' = 'whites are discriminated against', 'b_raceResent' = 'Racial resentment', 'b_pid' = 'Party ID', 'b_Intercept[1]' = 'Threshold 1', 'b_Intercept[2]' = 'Threshold 2', 'b_Intercept[3]' = 'Threshold 3', 'b_Intercept[4]' = 'Threshold 4')
+biased against Whites', 'b_inflwhite' = 'Whites influence politics', 'b_whitediscrim' = 'Whites are discriminated against', 'b_raceResent' = 'Racial resentment', 'b_pid' = 'Party ID', 'b_Intercept[1]' = 'Threshold 1', 'b_Intercept[2]' = 'Threshold 2', 'b_Intercept[3]' = 'Threshold 3', 'b_Intercept[4]' = 'Threshold 4')
 
 cm_pid <- c('b_rboff' = 'Retrospective better off', 'b_epast' = 'Economy better', 'b_unpast' = 'Unemployment better', 'b_edu' = 'Education', 'b_income' = 'Income', 'b_losejob' = 'Worried about losing job', 'b_ecfamily' = 'Worried about family finances', 'b_ecjob' = 'Know someone lost job', 'b_tradbreak' = 'New lifestyles breaking society', 'b_govbias' = 'Gov.
-biased against whites', 'b_inflwhite' = 'whites influence politics', 'b_whitediscrim' = 'whites are discriminated against', 'b_raceResent' = 'Racial resentment', 'b_female' = 'Female', 'b_Intercept[1]' = 'Threshold 1', 'b_Intercept[2]' = 'Threshold 2', 'b_Intercept[3]' = 'Threshold 3', 'b_Intercept[4]' = 'Threshold 4')
+biased against Whites', 'b_inflwhite' = 'Whites influence politics', 'b_whitediscrim' = 'Whites are discriminated against', 'b_raceResent' = 'Racial resentment', 'b_female' = 'Female', 'b_Intercept[1]' = 'Threshold 1', 'b_Intercept[2]' = 'Threshold 2', 'b_Intercept[3]' = 'Threshold 3', 'b_Intercept[4]' = 'Threshold 4')
 ```
@@
 #| label: tbl-2020-lasso-simple
-#| tbl-cap: "Predictors of White identity in 2020"
+#| tbl-cap: "Predictors of White Political Identity in 2020"
 model_2020 <- list(

### 2023-09-28 src/manuscript.qmd
fix some citations

@@
 thanks: |
-    Word count (including references): 3898.
+    Word count (including references): 3233 (4049).
 I would like to thank Jennifer Wolak for her feedback on the many drafts of this manuscript.
@@
         \newcommand{\elandscape}{\end{landscape}}
+    #citeproc: false
+    #filters: 
+    #  - "_extensions/andrewheiss/wordcount/citeproc.lua"
+    #  - "_extensions/andrewheiss/wordcount/wordcount.lua"
 linestretch: 2
@@
 Others have experimental evidence supporting this perspective that White Political Identity is motivated by material concerns [@perez_politics_2022].
-Further, @jardina_White_2019 who provides one of the most comprehensive treatments of conceptualizing, measuring, and examining the outcomes of White Political Identity includes indicators of economic status and power in the primary model.
+Further, @jardina_white_2019 who provides one of the most comprehensive treatments of conceptualizing, measuring, and examining the outcomes of White Political Identity includes indicators of economic status and power in the primary model.
 
@@
 A common touchstone of Social Dominance Theory suggests that members of a racial group with the highest status are likely to find ways to maintain the racial hierarchy [@lange_social_2012].
-From this perspective, many argue that Whites in the United States have and continue focus their efforts on maintaining influence in our institutions so that they may continue to wield political influence [@rile_hayward_how_2013;@lopez_White_2006].
+From this perspective, many argue that Whites in the United States have and continue focus their efforts on maintaining influence in our institutions so that they may continue to wield political influence [@rile_hayward_how_2013;@lopez_white_2006].
 For example, @mutz_status_2018 finds little support for the claim that the 2016 Presidential Election -- which support for Trump is associated with an expression of racial identity among Whites [@jardina_disavowing_2021] -- was a response to the economic effects of a globalizing economy and instead argues that it was a response to concerns about the cultural changes brought about by globalization and the racial-diversity it promotes.
 Others demonstrate that the behavioral manifestations of a White constituency in the United States was a reflection of the saliency of race in politics [@agadjanian_changing_2021].
+Additional evidence suggests that rural Americans -- who acutely feel economically left-behind by their governments [@cramer_politics_2016] -- express more racial animus as opposed to White Political Identity as a story of economic threat may suggest [@dawkins_place_2023].
+Rather non-urban Whites who express a White Political Identity are more likely to express an empathy for the rural White American who has been left-behind [@dawkins_place_2023].
 
@@
 
-In this note, I am using a popular conceptualization of racial identity among Whites: that expression of it reflects homophily toward's one's own racial group [@ellemers_self_2002;@jardina_White_2019].
+In this note, I am using a popular conceptualization of racial identity among Whites: that expression of it reflects homophily toward's one's own racial group [@ellemers_self_2002;@jardina_white_2019].
 As identity conceptualized this way may predict negative intergroup attitudes, it reflects a desire to find ways to find similarities rather than differences between one's self and other in-group members [@ellemers_self_2002].
 Identity is largely an elusive concept to operationalize [see @leach_group_2008].
-Though there are implicit measures of White Political Identity [see @knowles_White_2005], the common and measure of White Political Identity is a response to the question, "How important is being White to your identity?" This measure captures a significant feature of an identity: the centrality or the importance of that identity to an individual [@jardina_White_2019].
-Though a rather rough measure of something as complex as identity, it has been shown to produce similar results to more complex measures of White Political Identity and holds a significant amount of face validity [@jardina_White_2019].
+Though there are implicit measures of White Political Identity [see @knowles_white_2005], the common and measure of White Political Identity is a response to the question, "How important is being White to your identity?" This measure captures a significant feature of an identity: the centrality or the importance of that identity to an individual [@jardina_white_2019].
+Though a rather rough measure of something as complex as identity, it has been shown to produce similar results to more complex measures of White Political Identity and holds a significant amount of face validity [@jardina_white_2019].
 
@@
 The classic items of racial resentment construct capture these ideas in that they revolve around the idea that Blacks are not disadvantaged as the result of systematic oppression by Whites.
-Though many have often used racial resentment as a measure of prejudice which would be a possible outcome of White Political Identity, some make the compelling argument that it is distinct as a concept and operationalization [@davis_racial_2022].
+Though many have often used racial resentment as a measure of prejudice which would be a possible outcome of White Political Identity, some make the compelling argument that it is distinct as a concept and operationalization [@darren_w_davis_racial_2022].
 
@@
 contain a number of these features.
-As a result, a burgeoning literature in political science that has made this connection takes the racial identity of Whites more seriously as one that informs their political attitudes [see @jardina_White_2019].
+As a result, a burgeoning literature in political science that has made this connection takes the racial identity of Whites more seriously as one that informs their political attitudes [see @jardina_white_2019].
 Race in America is subjectively defined [see @rile-hayward_2013], and these definitions are deeply embedded and reinforced by our institutions [@lopez_2006].
@@
 ->
-Recent work makes the case that we should seriously consider Whites' racial identity as an identity [@jardina_White_2019].
+Recent work makes the case that we should seriously consider Whites' racial identity as an identity [@jardina_white_2019].
 As an identity, much of the literature view these attitudes resulting from Whites' racial identity as stemming from economic threat.
@@
 As much of the literature on Whites' racial attitudes rely heavily on social identity theory as a paradigm, a significant portion of it presents the source of these group-based views as responses to economic threat.
-However, some disagree with this position [see @jardina_White_2019] and view these attitudes as the result of changes to the racial hierarchy.
+However, some disagree with this position [see @jardina_white_2019] and view these attitudes as the result of changes to the racial hierarchy.
 These other perspectives provide explanations for Whites' attitudes and conceptualize the definition of Whiteness and White Political Identity as a desire to retain political power.
@@
 The prominent account of White political identity and its implications in contemporary American politics argues for a strong connection between White political identity and a desire to protect this racial hierarchy and status.
-This conceptualization reflects the tendency to report in-group-oriented, not necessarily out-group-oriented, attitudes [@jardina_White_2019].
+This conceptualization reflects the tendency to report in-group-oriented, not necessarily out-group-oriented, attitudes [@jardina_white_2019].
 
@@
 
-To account for potential problems stemming from missing data, I impute my data by using multiple imputation with chained equations (MICE) where the particular model I use is a random forest model to allow for more flexibility and to reduce model dependence [@buuren_mice_2011; @marbach_choosing_2022].^[I impute 10 datasets and simultaneously fit my Bayesian LASSO on each dataset and pool the results.
+To account for potential problems stemming from missing data, I impute my data by using multiple imputation with chained equations (MICE) where the particular model I use is a random forest model to allow for more flexibility and to reduce model dependence [@buuren_mice_2011; @moritz_marbach_choosing_2022].^[I impute 10 datasets and simultaneously fit my Bayesian LASSO on each dataset and pool the results.
 Uncertainty reflected in my reported credible intervals not only reflects the uncertainty from one model, but my reported uncertainty also reflects the uncertainty generated from the imputation procedure.

### 2023-09-28 src/manuscript.qmd
fixing a couple bad citation keys

@@
 title: |
-  Economic concerns appear to be weak predictors of White political identity
+  Economic concerns appear to be weak predictors of White Political Identity
 short-title: Economic concerns
@@
 The classic items of racial resentment construct capture these ideas in that they revolve around the idea that Blacks are not disadvantaged as the result of systematic oppression by Whites.
-Though many have often used racial resentment as a measure of prejudice which would be a possible outcome of White Political Identity, some make the compelling argument that it is distinct as a concept and operationalization [@darren_w_davis_racial_2022].
+Though many have often used racial resentment as a measure of prejudice which would be a possible outcome of White Political Identity, some make the compelling argument that it is distinct as a concept and operationalization [@davis_racial_2022].
 
@@
 
-To account for potential problems stemming from missing data, I impute my data by using multiple imputation with chained equations (MICE) where the particular model I use is a random forest model to allow for more flexibility and to reduce model dependence [@buuren_mice_2011; @moritz_marbach_choosing_2022].^[I impute 10 datasets and simultaneously fit my Bayesian LASSO on each dataset and pool the results.
+To account for potential problems stemming from missing data, I impute my data by using multiple imputation with chained equations (MICE) where the particular model I use is a random forest model to allow for more flexibility and to reduce model dependence [@buuren_mice_2011; @marbach_choosing_2022].^[I impute 10 datasets and simultaneously fit my Bayesian LASSO on each dataset and pool the results.
 Uncertainty reflected in my reported credible intervals not only reflects the uncertainty from one model, but my reported uncertainty also reflects the uncertainty generated from the imputation procedure.
